### 2023-10-16 Code/R/RMarkdown/EntropyPlots.Rmd
RMarkdown exercise

---
title: "Exploratory Plots"
author: "Alejandro C. Frery"
date: "2023-10-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggthemes)
library(reshape2)
```

# Gamma-SAR entropy

We will see how the Shannon entropy of the Gamma-SAR model varies.
It is given by
\begin{align}
H_{\Gamma_{\text{SAR}}}(L_0, \mu) &=   L_0 -\ln(L_0/\mu)+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\\
&= \big[L_0 -\ln L_0+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\big] + \ln \mu.
\end{align}
where $L_0\geq 1$ is known, and $\mu>0$ is the mean.
We see that, given $L_0$, the entropy of a random variable following the Gamma-SAR model depends on the logarithm of the mean $\mu$.

```{r FunctionsDefinitions}
entropy_gamma_sar <- function(L, mu) {

  return(L + log(gamma(L)) + (1 - L) * digamma(L) - log(L / mu))

}
```

```{r PlotGammaSAR, echo=FALSE}
mu <- seq(.1, 10, length.out=500)

EntropyL1 <- entropy_gamma_sar(L=1, mu)
EntropyL3 <- entropy_gamma_sar(L=3, mu)
EntropyL8 <- entropy_gamma_sar(L=8, mu)
EntropyL12 <- entropy_gamma_sar(L=12, mu)
EntropyL100 <- entropy_gamma_sar(L=100, mu)

muEntropy <- data.frame(
  mu, EntropyL1, EntropyL3, EntropyL8, EntropyL12, EntropyL100
)

muEntropy.molten <- melt(muEntropy, measure.vars = 2:6, value.name = "Entropy", variable.name = "Looks")

ggplot(muEntropy.molten, aes(x=mu, y=Entropy, col=Looks)) +
  geom_line() +
  theme_pander()

```

# GI0-SAR Entropy

The Shannon entropy of the GI0-SAR model is given by
\begin{multline}
H_{\mathcal{G}_I^0}(\mu, \alpha, L_0) = -\ln(\alpha)+\ln(\mu)+\ln(1+\alpha)
 -(1-\alpha)\psi^{(0)}(-\alpha) +\ln(-\alpha)-\ln L_0
+ (L_0-\alpha)\psi^{(0)}(L_0-\alpha) + \\
\ln\mathcal{B}(L_0,-\alpha)
  + (1-L_0)\psi^{(0)}(L_0), \label{E:E-GIO}
\end{multline}