### 2023-10-18 Code/R/RMarkdown/EntropyPlots.Rmd
R code of entropies

---
title: "Exploratory Plots"
author: "Alejandro C. Frery"
date: "2023-10-16"
output: pdf_document
#html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(plotly)
library(knitr)
library(rgl)
```

# Gamma-SAR entropy

We will see how the Shannon entropy of the Gamma-SAR model varies.
It is given by
\begin{align}
H_{\Gamma_{\text{SAR}}}(L_0, \mu) &=   L_0 -\ln(L_0/\mu)+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\\
&= \big[L_0 -\ln L_0+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\big] + \ln \mu.
\end{align}
where $L_0\geq 1$ is known, and $\mu>0$ is the mean.
We see that, given $L_0$, the entropy of a random variable following the Gamma-SAR model depends on the logarithm of the mean $\mu$.

```{r FunctionsDefinitions}
entropy_gamma_sar <- function(L, mu) {

    return(L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu))

}

entropy_gI0 <- function(mu, alpha, L) {

  term1 <- L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu)   
  term2 <- -L - log(gamma(L-alpha)) + (L-alpha)*(digamma(L - alpha))- (1-alpha)*digamma(- alpha)+log(-1 - alpha)+log(gamma(-alpha))

  entropy <- term1 + term2 
  return(entropy)
}
```

```{r PlotGammaSAR, echo=FALSE}
mu <- seq(.1, 10, length.out=500)

EntropyL1 <- entropy_gamma_sar(L=1, mu)
EntropyL3 <- entropy_gamma_sar(L=3, mu)
EntropyL8 <- entropy_gamma_sar(L=8, mu)
EntropyL12 <- entropy_gamma_sar(L=12, mu)
EntropyL100 <- entropy_gamma_sar(L=100, mu)

muEntropy <- data.frame(
  mu, EntropyL1, EntropyL3, EntropyL8, EntropyL12, EntropyL100
)

muEntropy.molten <- melt(muEntropy, measure.vars = 2:6, value.name = "Entropy", variable.name = "Looks")

ggplot(muEntropy.molten, aes(x=mu, y=Entropy, col=Looks)) +
  geom_line() +
  theme_pander()

```

# GI0-SAR Entropy

The Shannon entropy of the GI0-SAR model is given by

\begin{multline}
\label{E:E-GIO}
H_{\mathcal{G}_I^0}(\mu, \alpha, L_0) =\underbrace{L_0 -\ln L_0+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0) +\ln \mu}_{H_{\Gamma_{\text{SAR}}}} 
-\ln\Gamma(L_0-\alpha)+ (L_0-\alpha) \psi^{(0)}(L_0-\alpha)\\
-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L_0
\end{multline}

```{r PlotGI0, echo=FALSE}
mu <- seq(1, 10, length.out = 500)
alpha <- seq(-20, -2, length.out = 500)

entropy <- outer(mu, alpha, FUN = function(mu, alpha) entropy_gI0(mu, alpha, L = 1))

#
persp(x = mu, y = alpha, z = entropy, col = "lightblue", border = "blue")

#open3d()

#persp3d(x = mu, y = alpha,z = entropy, col = "lightblue", border = "blue")

#-----
# mu <- seq(1, 10, length.out=200)
# alpha <- seq(-2, -20, length.out=200)
# 
# 
# entropyL1 <- outer(mu, alpha, function(mu, alpha) entropy_gI0(mu, alpha, L=1))
# entropyL3 <- outer(mu, alpha,  function(mu, alpha) entropy_gI0(mu, alpha, L=3))
# entropyL8 <- outer(mu, alpha,  function(mu, alpha) entropy_gI0(mu, alpha, L=8))
# entropyL12 <- outer(mu, alpha, function(mu, alpha) entropy_gI0(mu, alpha, L=12))
# entropyL100 <- outer(mu, alpha, function(mu, alpha) entropy_gI0(mu, alpha, L=100))
# 
# 
# fig <- plot_ly( x = alpha, y = mu,    showscale = FALSE)
# 
# fig <- fig %>% add_surface(z = ~entropyL1, colorscale = list(c(0,1),c("rgb(255,107,184)","rgb(128,0,64)")),  name = "L=1")
# fig <- fig %>% add_surface(z = ~entropyL3, colorscale = list(c(0,1),c("rgb(107,184,255)","rgb(90,90,124)")),  name = "L=3")
# fig <- fig %>% add_surface(z = ~entropyL8,  colorscale = list(c(0,1),c("rgb(247,226,157)","rgb(255,192,7)")),  name = "L=8")
# fig <- fig %>% add_surface(z = ~entropyL12, colorscale = list(c(0,1),c("rgb(107,255,184)","rgb(0,124,90)")),  name = "L=12")
# fig <- fig %>% add_surface(z = ~entropyL100, colorscale = list(c(0,1),c("rgb(182,142,242)","rgb(104,3,255)")),  name = "L=100")
# 
# 
# fig <- fig %>% layout(
#   scene = list(
#     xaxis = list(title = "alpha"),
#     yaxis = list(title = "mu"),
#     zaxis = list(title = "Entropy")
#   )
# )
# 
# fig

```