### 2023-10-18 Code/R/RMarkdown/The Entropy as a Proxy for Fully Developed Speckle.Rmd
Working document in RMarkdown

---
title: Identifying Departures from the Fully Developed Speckle Hypothesis in Intensity SAR Data with Non-Parametric Estimation of the Entropy
affiliation:
  ## Author mode :  use one only of the following --- +

  ## one column per author - using only the institution field
  # author-columnar: true # uncomment this line to use

  ## one column per institution - using only the institution field
  # institution-columnar: true # uncomment this line to use

  ## one column wide author/affiliation fields - using author or author_multiline and institution for mark
  wide: true # uncomment this line to use

  ## Author mode: End ----

  ## A custom author ordering field that breaks institution grouping (used with `wide: true`)
  ## Authors cited in institution field are ignored when this author field exists
  ## Comment / Uncomment below
  author:   
   - name: Eldon Tyrell
     mark: [2, 3]
     email: eldon@starfleet-academy.star
   - name: Michael Shell
     mark: 1
   - name: Roy Batty
     mark: 4
     email: roy@replicant.offworld

  ## Put authors in a given order, with multiline possibility (used with `wide: true` if `authors` above unset)
  ## Authors cited in institution are ignored if exists
  ## Comment / Uncomment below
  #author_multiline: 
  #  - line:         ## Create a new author line
  #    - name: Michael Shell
  #      mark: 1
  #    - name: Homer Simpson
  #      mark: 2
  #      email: homer@thesimsons.com
  #    - name: Roy Batty
  #      mark: 4
  #      email: roy@replicant.offworld
  #  - line:         ## Create a second author line
  #    - name: Montgomery Scott
  #      mark: 3
  #    - name: Eldon Tyrell
  #      mark: 4

  ## Define institution, and also authors used when `author-columnar: true` or `institution-columnar: true`
  institution:
    - name: Georgia Institute of Technology
      department: School of Electrical and Computer Engineering
      location: Atlanta, Georgia 30332--0250
      email: ece@datech.edu # if unset, authors' email will be used
      mark: 1
      author:
        - name: Michael Shell
    - name: Twentieth Century Fox
      location: Springfield, USA
      mark: 2
      author:
        - name: Homer Simpson
          email: homer@thesimpsons.com
    - name: Starfleet Academy
      location: San Francisco, California 96678-2391
      other: "Telephone: (800) 555--1212, Fax: (888) 555--1212"
      mark: 3
      author:
        - name: Montgomery Scott
    - name: Tyrell Inc.
      location: 123 Replicant Street, Los Angeles, USA 90210--4321
      mark: 4
      author:
        - name: Eldon Tyrell
          email: eldon@starfleet-academy.star
        - name: Roy Batty
          email: roy@replicant.offworld

keywords: ["SAR", "entropy estimation", "non-parametric analysis", "order statistics"]
abstract: |
  SAR Data are affected by speckle in a non-additive and non-Gaussian way.
  The type of distribution these data follow is paramount for their processing and analysis.
  Good statistical models provide flexibility and accuracy, often at the cost of using several parameters.
  The $\mathcal{G}^0$ distribution is one of the most successful models for SAR data. 
  It includes the Gamma law as particular case which arises in the presence of fully developed speckle.
  Although the latter is a limit distribution of the former, using the same estimation technique for the more general model is numerically unfeasible.
  We propose a two-stage estimation procedure: first, we verify the hypothesis that the data are fully-developed speckle and, if this assumption is rejected, we proceed to estimate the parameters that index the $\mathcal G^0$ distribution.
  Given the uncertainty of the underlying distribution, and the negative impact that using an inadequate model has on maximum likelihood estimation, we employ a non-parametric approach to estimate entropy under the fully-developed speckle hypothesis.

# use some specific Tex packages if needed. 
# with_ifpdf: true
# with_cite: true
# amsmath need to be true to use with bookdown for referencing equations.
with_amsmath: true
# with_algorithmic: true
# with_array: true
# with_dblfloatfix: true

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   - \usepackage{booktabs}
   - \usepackage{color}

bibliography: ["mybibfile.bib"]
output: rticles::ieee_article
#citation_sorting: none   ## used as sorting option of the biblatex package (if selected)
---

```{r Setup, echo=FALSE}
library(ggplot2)
library(ggthemes)

theme_set(theme_pander() +
            theme(text=element_text(family="serif"),
                  legend.position = "top")
          )
```

# Introduction

Remote sensing data from sensors, in particular synthetic aperture radar (SAR), play a key role in understanding the dynamics of the environment, enabling predictive capabilities and facilitating the early detection of disasters. Unlike optical spectrum observations, microwave remote sensing sensors, operating in the microwave band, offer significant advantages, including reduced sensitivity to adverse atmospheric conditions and their self-illuminating properties.

Entropy is a fundamental concept in information theory with broad applications in fields like classification, pattern recognition, statistical physics, stochastic dynamics, and statistics.
Shannon introduced it for a random variables in 1948~\cite{Shannon1948} as a measure of information and uncertainty. 
In statistics, Shannon entropy is a crucial descriptive parameter, particularly for assessing data dispersion and conducting tests for normality, exponentiality, and uniformity~\cite{Wieczorkowski1999}.

Estimating entropy presents practical challenges, especially when the true probability density function (pdf) is unknown.
In such cases, nonparametric methods are employed for entropy estimation. 
In the parametric approach, the form of the pdf is assumed to be known, and its parameters are inferred from the available samples. 
Conversely, nonparametric methods, like histogram or kernel density estimators, do not make such assumptions. 
They estimate the pdf using these techniques and then calculate entropy through numerical or Monte Carlo integration. 
Other nonparametric approaches involve spacing methods, which allow the estimation of the entropy of a random variable with an unknown distribution function when independent and identically distributed (i.i.d.) observations of the random variable are available.

%In this paper we consider an estimator based on spacings.

This study is dedicated to the analysis of SAR intensity data, which frequently contends with the presence of speckle noise. 
Speckle noise can significantly complicate the analysis of these data, necessitating the use of specific models and techniques for information extraction. 
We adopt the $\mathcal G^0$ distribution family as a suitable model for describing SAR intensity data, as it effectively characterizes areas with varying degrees of texture.

The article is structured as follows: 
Section~\ref{sec_02} describes

# Background

## Statistical modeling of Intensity SAR data

The primary models used for intensity SAR data include the Gamma and $\mathcal{G}_I^0$  distributions~\cite{Frery1997}. 
The first is suitable for fully developed speckle and can also be considered a limiting case within the $\mathcal{G}_I^0$ distribution.
Whereas the latter is appealing due to its versatility in accurately representing regions with various roughness characteristics~\cite{Cassetti2022}.
We denote $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ and $Z \sim G_I^0(\alpha, \gamma, L)$ to indicate that $Z$ follows the distributions characterized by the respective probability density functions:
\begin{align}
	f_Z(z;L, \mu)&=\frac{L^L}{\Gamma(L)\mu^L}z^{L-1}\exp\left\{-Lz/\mu\right\},\label{E:gamma1}\\
	f_Z(z; \alpha, \gamma, L)&=\frac{L^L\Gamma(L-\alpha)}{\gamma^{\alpha}\Gamma(-\alpha)\Gamma(L)}\cdot\frac{z^{L-1}}{(\gamma+Lz)^{L-\alpha}},\label{E:gi01}
\end{align}
where, in~\eqref{E:gamma1} $\mu > 0$ is the mean; in~\eqref{E:gi01}  $\gamma > 0$ is the scale, $\alpha < -1$ measures the roughness,  $L \geq 1$ is the number of looks, and $\Gamma(\cdot)$ is the gamma function.

From \eqref{E:gi01}, the $r$th moment of $Z$ is expressed as:
\begin{align}
	E_{G_I^0}\left(Z^r\right)=\left(\frac{\gamma}{L}\right)^r\frac{\Gamma(-\alpha-r)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+r)}{\gamma(L)}, \quad \alpha <-r. 
	\label{E:rmom}
\end{align}

Even though the $\mathcal{G}_I^0$  distribution is defined by the parameters $\alpha$ and $\gamma$, SAR literature commonly utilizes the texture $\alpha$ and the mean $\mu$~\cite{Nascimento2010}.
In this way, we compute the expected value $\mu$ using the expression in~\eqref{E:rmom}, and we reparametrize~\eqref{E:gi01} using $\mu$, $\alpha$, and $L$. Then
\begin{align*}
	\mu=\left(\frac{\gamma}{L}\right)\frac{\Gamma(-\alpha-1)}{\Gamma(-\alpha)}\cdot\frac{\Gamma(L+1)}{\gamma(L)}=-\frac{\gamma}{\alpha+1}.
\end{align*}
Thus, the probability density functions that characterize the $G_I^0(\mu, \alpha, L)$ is
\begin{multline}
		f_Z(z; \mu, \alpha, L)=\frac{L^L\Gamma(L-\alpha)}{\big(-\mu(\alpha+1)\big)^{\alpha}\Gamma(-\alpha)\Gamma(L)}\\ \frac{z^{L-1}}{\big(-\mu(\alpha+1)+Lz\big)^{L-\alpha}}.\label{E:gi02}
\end{multline}

## Parametric \& Non-Parametric Shannon Entropy

The parametric representation of Shannon entropy for a system described by a continuous random variable is as follows:
\begin{equation}
  \label{E:entropy2}
  H_(Z)=-\int_{-\infty }^\infty \ f(z)\log(f(z))\, \mathrm{d}z,
\end{equation}
here, $f(\cdot)$ is the probability density function that characterizes the distribution of the real-valued random variable $Z$.

Using~\eqref{E:entropy2}, we can express the Shannon entropy of $\Gamma_{\text{SAR}}$ in~\eqref{E:gamma1} and $\mathcal{G}_I^0$ in~\eqref{E:gi02} based on~\cite{Cassetti2022, A.Ferreira2020}:
\begin{multline}
\label{E:E-gamma}
H_{\Gamma_{\text{SAR}}}(L, \mu) =   L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) + \ln \mu, 
\end{multline}
\begin{multline}
\label{E:E-GIO}
H_{\mathcal{G}_I^0}(\mu, \alpha, L) =L -\ln L+\ln\Gamma(L)+(1-L)\psi^{(0)}(L) +\ln \mu \\
-\ln\Gamma(L-\alpha)+ (L-\alpha) \psi^{(0)}(L-\alpha)\\
-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L
\end{multline}
where $\psi^{(0)}(\cdot)$ and $\mathcal{B}(\cdot)$ are the digamma and beta functions, respectively.

The problem of non-parametric estimating of $H(Z)$ has been considered by many authors including~\cite{vasicek1976test, Bert1992, Wieczorkowski1999, correa1995new}, who proposed estimators based on spacings.

Vasicek \cite{vasicek1976test}, used $f(z)=p$  to express $ H(Z)$ as
\begin{equation*}
	H(Z)= \int_0^1 \log\frac{\mathrm{d}}{\mathrm{d}p}Q(p)\mathrm{d}p,
\end{equation*}
where $Q(p)=F^{-1}(p)=\inf\left\{z: F(z)\leq p\right\}$ is the quantile function. The derivative of $F^{-1}(p)$ is  estimated by a function of the order statistics~\cite{AlOmari2019}.

Assuming that  $\bm{Z}=(Z_1, Z_2, \ldots,Z_n)$ is a random sample from the distribution $F(z)$, the estimator is defined as:
\begin{equation*}
%\label{E:VanEs}
	\widehat{H}_{V}(\bm{Z})=\frac{1}{n}\sum_{i=1}^{n}\log\left[\frac{n}{2m}\left(Z_{(i+m)}-Z_{(i-m)}\right)\right],
	\end{equation*}
where $m<n/2$ is a positive integer, $Z_{(i+m)}-Z_{(i-m)}$ is the $m$-spacing and $Z_{(1)}\leq Z_{(2)}\leq\ldots\leq Z_{(n)}$ are the order statistics and $Z_{(i)}= Z_{(1)}$ if $i<1$, $Z_{(i)}= Z_{(n)}$ if $i>n$.

Multiple authors have presented  adaptations to Vasicek's estimator, including Van Es~\cite{Bert1992}, who proposed a new estimator of entropy given by:
\begin{multline}
\label{E:VanEs}
	\widehat{H}_{\text{VE}}(\bm{Z})=\frac{1}{n-m}\sum_{i=1}^{n-m}\log\left[\frac{n+1}{m}\left(Z_{(i+m)}-Z_{(i)}\right)\right]\\
	+\sum_{k=m}^n\frac{1}{k}+\log\frac{m}{n+1}.
\end{multline}

Van Es proved that, under general conditions,~\eqref{E:VanEs} converges almost surely to $H(Z)$ when $m, n\rightarrow\infty$,  $m/ \log(n)\rightarrow\infty$, and $m/n \rightarrow0$.
The author also proved the asymptotic normality of the estimator when $m, n\rightarrow\infty$ and $m=o(n^{1/2})$.

# Hypothesis tests

# Results

This is an example of how to produce a plot in an RMarkdown reproducible paper.

First, you define the functions, or you load them from a library.

```{r Functions, echo=FALSE}
myfunction <- function(x){
  return(x^2)
}
```

Then, you produce and display the plot.
```{r MyFirstPlot, echo=FALSE, out.width="100%", fig.cap="My first function in a reproducible RMardown document"}
x <- seq(-3, 3, length.out=500)
y <- myfunction(x)

df <- data.frame(x, y)

ggplot(df, aes(x=x, y=y)) +
  geom_line() +
  xlab(expression(italic(x))) +
  ylab(expression(italic(y))) 
```

# Conclusions

<!-- An example of a floating figure using the graphicx package.
Note that \label must occur AFTER (or within) \caption.
For figures, \caption should occur after the \includegraphics.
Note that IEEEtran v1.7 and later has special internal code that
is designed to preserve the operation of \label within \caption
even when the captionsoff option is in effect. However, because
of issues like this, it may be the safest practice to put all your
\label just after \caption rather than within \caption{}.

Reminder: the "draftcls" or "draftclsnofoot", not "draft", class
option should be used if it is desired that the figures are to be
displayed while in draft mode.

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{myfigure}
where an .eps filename suffix will be assumed under latex, 
and a .pdf suffix will be assumed for pdflatex; or what has been declared
via \DeclareGraphicsExtensions.
\caption{Simulation results for the network.}
\label{fig_sim}
\end{figure}

Note that the IEEE typically puts floats only at the top, even when this
results in a large percentage of a column being occupied by floats.

An example of a double column floating figure using two subfigures.
(The subfig.sty package must be loaded for this to work.)
The subfigure \label commands are set within each subfloat command,
and the \label for the overall figure must come after \caption.
\hfil is used as a separator to get equal spacing.
Watch out that the combined width of all the subfigures on a 
line do not exceed the text width or a line break will occur.

\begin{figure*}[!t]
\centering
\subfloat[Case I]{\includegraphics[width=2.5in]{box}%
\label{fig_first_case}}
\hfil
\subfloat[Case II]{\includegraphics[width=2.5in]{box}%
\label{fig_second_case}}
\caption{Simulation results for the network.}
\label{fig_sim}
\end{figure*}

Note that often IEEE papers with subfigures do not employ subfigure
captions (using the optional argument to \subfloat[]), but instead will
reference/describe all of them (a), (b), etc., within the main caption.
Be aware that for subfig.sty to generate the (a), (b), etc., subfigure
labels, the optional argument to \subfloat must be present. If a
subcaption is not desired, just leave its contents blank,
e.g., \subfloat[].

An example of a floating table. Note that, for IEEE style tables, the
\caption command should come BEFORE the table and, given that table
captions serve much like titles, are usually capitalized except for words
such as a, an, and, as, at, but, by, for, in, nor, of, on, or, the, to
and up, which are usually not capitalized unless they are the first or
last word of the caption. Table text will default to \footnotesize as
the IEEE normally uses this smaller font for tables.
The \label must come after \caption as always.

\begin{table}[!t]
% increase table row spacing, adjust to taste
\renewcommand{\arraystretch}{1.3}
if using array.sty, it might be a good idea to tweak the value of
\extrarowheight as needed to properly center the text within the cells
\caption{An Example of a Table}
\label{table_example}
\centering
% Some packages, such as MDW tools, offer better commands for making tables
% than the plain LaTeX2e tabular which is used here.
\begin{tabular}{|c||c|}
\hline
One & Two\\
\hline
Three & Four\\
\hline
\end{tabular}
\end{table}

Note that the IEEE does not put floats in the very first column
- or typically anywhere on the first page for that matter. Also,
in-text middle ("here") positioning is typically not used, but it
is allowed and encouraged for Computer Society conferences (but
not Computer Society journals). Most IEEE journals/conferences use
top floats exclusively. 
Note that, LaTeX2e, unlike IEEE journals/conferences, places
footnotes above bottom floats. This can be corrected via the
\fnbelowfloat command of the stfloats package. -->

# Conclusion

The conclusion goes here.

<!-- conference papers do not normally have an appendix -->

# Acknowledgment {#acknowledgment}

The authors would like to thank...

# Bibliography styles

Here are two sample references: @Feynman1963118 [@Dirac1953888].

\newpage
# References {#references .numbered}