### 2023-10-25 Code/R/RMarkdown/EntropyPlots.Rmd
Update R code bias

---
title: "Exploratory Plots"
author: "Alejandro C. Frery"
date: "2023-10-16"
output: pdf_document
#html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(ggthemes)
library(reshape2)
library(plotly)
library(knitr)
#library(rgl)
theme_set(theme_pander() +
            theme(text=element_text(family="serif"),
                  legend.position = "top")
          )
```

# Gamma-SAR entropy

We will see how the Shannon entropy of the Gamma-SAR model varies.
It is given by
\begin{align}
H_{\Gamma_{\text{SAR}}}(L_0, \mu) &=   L_0 -\ln(L_0/\mu)+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\\
&= \big[L_0 -\ln L_0+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0)\big] + \ln \mu.
\end{align}
where $L_0\geq 1$ is known, and $\mu>0$ is the mean.
We see that, given $L_0$, the entropy of a random variable following the Gamma-SAR model depends on the logarithm of the mean $\mu$.

```{r FunctionsDefinitions}
entropy_gamma_sar <- function(L, mu) {

    return(L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu))

}

entropy_gI0 <- function(mu, alpha, L) {

  term1 <- L - log(L) + log(gamma(L)) + (1 - L) * digamma(L) + log(mu)   
  term2 <- -L - log(gamma(L-alpha)) + (L-alpha)*(digamma(L - alpha))- (1-alpha)*digamma(- alpha)+log(-1 - alpha)+log(gamma(-alpha))

  entropy <- term1 + term2 
  return(entropy)
}

gamma_sar_sample <- function(L, mu, n) {
  samples <- rgamma(n, shape = L, rate = L / mu)
  return(samples)
}

van_es_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)  # m-spacing
  data_sorted <- sort(data)
  sum_term1 <- 0
  sum_term2 <- 0

  for (i in 1:(n - m)) {
    sum_term1 <- sum_term1 + log(((n + 1) / m) * (data_sorted[i + m] - data_sorted[i]))
  }

  for (k in m:n) {
    sum_term2 <- sum_term2 + 1 / k
  }

  sum_term3 <- log(m / (n + 1))

  return((sum_term1 / (n - m)) + sum_term2+ sum_term3)

}
vasicek_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)  # m <- floor(n / 2)
  data_sorted <- sort(data)
  sum_term <- 0

  for (i in 1:n) {
    if (i <= m) {
      diff_term <- data_sorted[i + m] - data_sorted[1]
    } else if (i >= n - m + 1) {
      diff_term <- data_sorted[n] - data_sorted[i - m]
    } else {
      diff_term <- data_sorted[i + m] - data_sorted[i - m]
    }

    sum_term <- sum_term + log((n / (2 * m)) * diff_term)
  }

  return(sum_term / n)
}
correa_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)
  X_sorted <- sort(data)
  C_mn <- 0

  for (i in 1:n) {
    i_minus_m <- max(1, i - m)
    i_plus_m <- min(n, i + m)

    X_i <- X_sorted[i]
    X_window <- X_sorted[i_minus_m:i_plus_m]
    X_bar_i <- mean(X_window)

    num <- sum((X_window - X_bar_i) * (i_minus_m:i_plus_m - i))
    den <- n * sum((X_window - X_bar_i)^2)

    if (den == 0) {
      b_i <- 0
    } else {
      b_i <- num / den
    }

    if (b_i > 0) {
      C_mn <- C_mn - log(b_i)
    }
  }

  C_mn <- C_mn / n

  return(C_mn)
}
ebrahimi_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)  # m-spacing
  data_sorted <- sort(data)
  sum_term <- 0

  for (i in 1:n) {
    if (i <= m) {
      ci <- 1 + (i - 1) / m
      diff_term <- data_sorted[i + m] - data_sorted[1]
    } else if (i >= n - m + 1) {
      ci <- 1 + (n - i) / m
      diff_term <- data_sorted[n] - data_sorted[i - m]
    } else {
      ci <- 2
      diff_term <- data_sorted[i + m] - data_sorted[i - m]
    }

    sum_term <- sum_term + log((n / (ci * m)) * diff_term)
  }

  return(sum_term / n)
}
noughabi_arghami_estimator <- function(data) {
    n <- length(data)
    m <- round(sqrt(n) + 0.5)  # m-spacing
    data_sorted <- sort(data)
    sum_term <- 0

    for (i in 1:n) {
      if (i <= m) {
        ai <- 1
        diff_term <- data_sorted[i + m] - data_sorted[1]
      } else if (i >= n - m + 1) {
        ai <- 1
        diff_term <- data_sorted[n] - data_sorted[i - m]
      } else {
        ai <- 2
        diff_term <- data_sorted[i + m] - data_sorted[i - m]
      }

      sum_term <- sum_term + log((n / (ai * m)) * diff_term)
    }

    return(sum_term / n)
}
al_omari_1_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)  # m-spacing
  data_sorted <- sort(data)
  sum_term <- 0

  for (i in 1:n) {
    if (i <= m) {
      omega_i <- 3/2
      diff_term <- data_sorted[i + m] - data_sorted[1]
    } else if (i >= n - m + 1) {
      omega_i <- 3/2
      diff_term <- data_sorted[n] - data_sorted[i - m]
    } else {
      omega_i <- 2
      diff_term <- data_sorted[i + m] - data_sorted[i - m]
    }

    sum_term <- sum_term + log((n / (omega_i * m)) * diff_term)
  }

  return(sum_term / n)
}
al_omari_2_estimator <- function(data) {
  n <- length(data)
  m <- round(sqrt(n) + 0.5)  # m-spacing
  data_sorted <- sort(data)
  sum_term <- 0

  for (i in 1:n) {
    if (i <= m) {
      vi <- 1 + (i - 1) / m
      diff_term <- data_sorted[i + m] - data_sorted[1]
    } else if (i >= n - m + 1) {
      vi <- 1 + (n - i) / (2 * m)
      diff_term <- data_sorted[n] - data_sorted[i - m]
    } else {
      vi <- 2
      diff_term <- data_sorted[i + m] - data_sorted[i - m]
    }

    sum_term <- sum_term + log((n / (vi * m)) * diff_term)
  }

  return(sum_term / n)
}
```

```{r PlotGammaSAR, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Simulation of entropy for Gamma SAR"}
L <- c(1, 3, 8, 12, 100)

L_labels <- c("L=1", "L=3", "L=8", "L=12", "L=100")

mu <- seq(0.1, 10, length.out = 500)

entropies <- sapply(L, function(L) entropy_gamma_sar(L, mu))

muEntropy <- data.frame(mu, entropies)

muEntropy.molten <- melt(muEntropy, id.vars = "mu", variable.name = "Looks", value.name = "Entropy")

ggplot(muEntropy.molten, aes(x = mu, y = Entropy, col = Looks)) +
  geom_line() +
  theme_minimal() +
  scale_color_manual(values = rainbow(length(L)), labels = L_labels) +
  labs(col = "Looks") +
  xlab(expression(paste(mu)))

```

# GI0-SAR Entropy

The Shannon entropy of the GI0-SAR model is given by

\begin{multline}
\label{E:E-GIO}
H_{\mathcal{G}_I^0}(\mu, \alpha, L_0) =\underbrace{L_0 -\ln L_0+\ln\Gamma(L_0)+(1-L_0)\psi^{(0)}(L_0) +\ln \mu}_{H_{\Gamma_{\text{SAR}}}} 
-\ln\Gamma(L_0-\alpha)+ (L_0-\alpha) \psi^{(0)}(L_0-\alpha)\\
-(1-\alpha)\psi^{(0)}(-\alpha)+\ln (-1-\alpha)+\ln\Gamma(-\alpha)-L_0
\end{multline}

```{r 3d_GIO, fig.align="center",  out.width = "120%", fig.cap="Simulation of entropy for ${G}_I^0$. ", fig.show="hold",  fig.pos="hbt"}
knitr::include_graphics("../../../Figures/PDF/entropy_plot_3d.pdf")
```

# Non-Parametric Entropy Estimators

In 1995 Correa suggested a modification of Vasicek's estimator. In estimation the density $f$ of $F$ in the interval $\left(X_{(i-m)}, X_{(i+m)}\right)$ he used a local linear model based on $2 m+1$ points: $F\left(X_{(j)}\right)=\alpha+\beta X_{(j)}+\varepsilon, j=m-i, \ldots, m+i$. This yields a following estimator
\begin{equation}
\widehat{H}_C(\mathbf{X})=-\frac{1}{n} \sum_{i=1}^n \ln \left(b_i\right),
\end{equation}
where
$$
\begin{gathered}
b_i=\frac{\sum_{j=i-m}^{i+m}\left(X_{(j)}-\bar{X}_{(i)}\right)(j-i)}{n \cdot \sum_{j=i-m}^{i+m}\left(X_{(j)}-\bar{X}_{(i)}\right)^2}, \\
\bar{X}_{(i)}=\frac{1}{2 m+1} \sum_{j=i-m}^{i+m} X_{(j)},
\end{gathered}
$$
$m$ is a positive integer smaller than $\frac{n}{2}, X_{(i)}=X_{(1)}$ for $i<1$ and $X_{(i)}=X_{(n)}$ for $i>n$.

Ebrahimi et al. (1994), adjusted the weights of Vasicek's estimator, in order to take into account the fact that the differences are truncated around the smallest and the largest data points. i.e. $X_{(i+m)}-X_{(i-m)}$ is replaced by $X_{(i+m)}-X_{(1)}$ when $i \leq m$ and $X_{(i+m)}-X_{(i-m)}$ is replaced by $X_{(n)}-X_{(i-m)}$ when $i \geq n-m+1$. Their estimator is given by
\begin{equation}
\widehat{H}_E(\mathbf{X})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{c_i m}\left(X_{(i+m)}-X_{(i-m)}\right)\right],
\end{equation}
where
$$
c_i=\left\{\begin{array}{lr}
1+\frac{i-1}{m}, & 1 \leq i \leq m \\
2, & m+1 \leq i \leq n-m, \\
1+\frac{n-i}{m}, & n-m+1 \leq i \leq n
\end{array}\right.
$$

Noughabi and Arghami suggested an entropy estimator given by

\begin{equation}
\widehat{H}_{NA}(\mathbf{X})=\frac{1}{n}(\mathbf{X}) \sum_{i=1}^n \log \left[\frac{n}{a_i m}\left(X_{(i+m)}-X_{(i-m)}\right)\right],
\end{equation}
where
$$
a_i=\left\{\begin{array}{lc}
1, & 1 \leq i \leq m, \\
2, & m+1 \leq i \leq n-m, \\
1, & n-m+1 \leq i \leq n,
\end{array}\right.
$$
and $X_{(i-m)}=X_{(1)}$ for $i \leq m$ and $X_{(i+m)}=X_{(n)}$ for $i \geq n-m$.

Al-Omari suggested the following estimators of entropy:

\begin{equation}
\widehat{H}_{\mathrm{AO}_1}(\mathbf{X})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{\omega_i m}\left(X_{(i+m)}-X_{(i-m)}\right)\right]
\end{equation}
where
$$
\omega_i= \begin{cases}3 / 2 & \text { if } 1 \leq i \leq m \\ 2 & \text { if } m+1 \leq i \leq n-m, \\ 3 / 2 & \text { if } n-m+1 \leq i \leq n,\end{cases}
$$
in which $X_{(i-m)}=X_{(1)}$ for $i \leq m$, and $X_{(i+m)}=X_{(n)}$ for $i \geq n-m$.
And
\begin{equation}
\widehat{H}_{\mathrm{AO}_2}(\mathbf{X})=\frac{1}{n} \sum_{i=1}^n \log \left[\frac{n}{v_i m}\left(X_{(i+m)}-X_{(i-m)}\right)\right]
\end{equation}
where
$$
v_i= \begin{cases}1+(i-1) / m & \text { if } 1 \leq i \leq m \\ 2 & \text { if } m+1 \leq i \leq n-m, \\ 1+(n-i) / 2 m & \text { if } n-m+1 \leq i \leq n\end{cases}
$$
in which $X_{(i-m)}=X_{(1)}$ for $i \leq m$, and $X_{(i+m)}=X_{(n)}$ for $i \geq n-m$.

Fig. \ref{fig:PlotBias}, illustrate  the bias  of each estimator as a function of sample size. We observe that as the sample size n increases, the bias tends to zero, for some estimators, this means that the estimators are becoming more unbiased as the amount of data available increases, because they are getting closer to the value true of entropy.

```{r PlotBias, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Bias, for $L= 1$, $\\mu=1$."}
sample_sizes <- c(9, 25, 49, 81, 121, 225)
R <- 1000
mu <- 1
L <- 1

#  True entropy value for the gamma SAR distribution
true_entropy <- entropy_gamma_sar(L, mu)

output <- data.frame(SampleSize = integer(0), Estimator = character(0), Bias = numeric(0))

#List of estimator
estimators <- list(
  list(func = van_es_estimator, label = "Van Es"),
  list(func = correa_estimator, label = "Correa"),
  list(func = ebrahimi_estimator, label = "Ebrahimi"),
  list(func = noughabi_arghami_estimator, label = "Noughabi Arghami"),
  list(func = vasicek_estimator, label = "Vasicek"),
  list(func = al_omari_1_estimator, label = "Al Omari 1"),
  list(func = al_omari_2_estimator, label = "Al Omari 2")
)

for (estimator in estimators) {
  for (ssize in sample_sizes) {
    v.nonparametric.entropy <- numeric(R)
    for (r in 1:R) {
      sample <- gamma_sar_sample(L, mu, ssize)
      v.nonparametric.entropy[r] <- estimator$func(sample)
    }

    # Calculate the bias 
    bias <- mean(v.nonparametric.entropy) - true_entropy

    output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator$label, Bias = bias))
  }
}

# Plot the bias as a function of the sample size 
ggplot(output, aes(x = SampleSize, y = Bias, color = Estimator)) +
  geom_line() +
  geom_point() +
  xlab("Sample Size") +
  ylab(expression(paste("Bias ")))

```

Figure \ref{fig:PlotMSE}, shows the Mean Square Error of each estimator as a function of sample size. We see that the lower the MSE, the better the performance of the estimator, as it indicates that the estimates are closer to the true values.
```{r PlotMSE, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="MSE, for $L= 1$, $\\mu=1$."}
sample_sizes <- c(9, 25, 49, 81, 121, 225)
R <- 1000
mu <- 1
L <- 2

# True entropy value for the gamma SAR distribution
true_entropy <- entropy_gamma_sar(L, mu)

output_mse <- data.frame(SampleSize = integer(0), Estimator = character(0), MSE = numeric(0))
#List of estimator
estimators <- list(
  list(func = van_es_estimator, label = "Van Es"),
  list(func = correa_estimator, label = "Correa"),
  list(func = ebrahimi_estimator, label = "Ebrahimi"),
  list(func = noughabi_arghami_estimator, label = "Noughabi Arghami"),
  list(func = vasicek_estimator, label = "Vasicek"),
  list(func = al_omari_1_estimator, label = "Al Omari 1"),
  list(func = al_omari_2_estimator, label = "Al Omari 2")
)

for (estimator in estimators) {
  for (ssize in sample_sizes) {
    squared_errors <- numeric(R)
    for (r in 1:R) {
      sample <- gamma_sar_sample(L, mu, ssize)
      estimated_entropy <- estimator$func(sample)
      squared_errors[r] <- (estimated_entropy - true_entropy)^2
    }

    # Calculate the MSE
    mse <- mean(squared_errors)

    output_mse <- rbind(output_mse, data.frame(SampleSize = ssize, Estimator = estimator$label, MSE = mse))
  }
}

# Plot the MSE as a function of the sample size
ggplot(output_mse, aes(x = SampleSize, y = MSE, color = Estimator)) +
  geom_line() +
  geom_point() +
  xlab("Sample Size") +
  ylab("MSE")

```