### 2023-12-09 Code/R/RMarkdown/report2_plots_means.Rmd
adding improvements for code and paper

---
title: "Report: Exploring Non-Parametric Entropy Estimators and Bootstrap Enhancements"
author: ""
date: ""
output: pdf_document

# output: 
#   pdf_document:
#     toc: true
#     toc_depth: 2
#html_document
header-includes:
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{amsmath}
  - \usepackage{mathabx}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
#library(RColorBrewer)
library(ggsci)
#library(wesanderson)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

source("../MainFunctions/gamma_sar_sample.r")
source("../MainFunctions/entropy_gamma_sar.r")
source("../MainFunctions/entropy_gI0.r")
source("../MainFunctions/gi0_sample.r")

source("../MainFunctions/van_es_estimator.r")
source("../MainFunctions/correa_estimator.r")
source("../MainFunctions/adaptive_correa_estimator.r")
source("../MainFunctions/ebrahimi_estimator.r")
source("../MainFunctions/noughabi_arghami_estimator.r")
source("../MainFunctions/vasicek_estimator.r")
source("../MainFunctions/al_omari_1_estimator.r")
source("../MainFunctions/al_omari_estimator_m.r")
source("../MainFunctions/al_omari_2_estimator.r")
source("../MainFunctions/yee_estimator.r")

source("../MainFunctions/bootstrap_van_es_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator_opt.r")
source("../MainFunctions/bootstrap_ebrahimi_estimator.r")
source("../MainFunctions/bootstrap_noughabi_arghami_estimator.r")
source("../MainFunctions/bootstrap_vasicek_estimator.r")
source("../MainFunctions/bootstrap_al_omari_1_estimator.r")
source("../MainFunctions/bootstrap_al_omari_2_estimator.r")
source("../MainFunctions/bootstrap_yee_estimator.r")
source("../MainFunctions/bootstrap_yee_estimator_optimized.r")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Programs/functions_sample_bias_mse.R")#

set.seed(1234567890, kind = "Mersenne-Twister")

```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

\newcommand{\bias}{\operatorname{Bias}}

# Mean Entropy for Estimators

```{r Plot_mean1, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for Gamma SAR $L= 1$ and $\\mu=1$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121, 225)
R <- 100
mu <- 1
L <- 1

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gamma_sar_sample(L, mu, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gamma_sar(L, mu)
calculate_entropy <- function(sample_sizes, R, mu, L) {
 # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
     "Correa" = correa_estimator,
     "Ebrahimi" = ebrahimi_estimator,
     "Al Omari 1" = al_omari_1_estimator,
      "Yee" = yee_estimator,
      "Correa Bootstrap" = bootstrap_correa_estimator,
     "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
     "Al Omari 1 Bootstrap" = bootstrap_al_omari_1_estimator
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
       annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

```{r Plot_mean2, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy Gamma SAR for $L= 2$ and $\\mu=5$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 100
mu <- 5
L <- 2

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gamma_sar_sample(L, mu, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gamma_sar(L, mu)
calculate_entropy <- function(sample_sizes, R, mu, L) {
 # true_entropy <- entropy_gamma_sar(L, mu)

# Define a list of estimators
  estimators <- list(
   # "Van Es" = van_es_estimator,
   # "Correa" = correa_estimator,
   # "Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
   # "Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
   #"Al OmariA " = al_omari_estimator_m,
    #"Al Omari 2" = al_omari_2_estimator,
    "Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator,
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
    #"Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
       annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

## Results for  Nonparametric Estimators with  $\mathcal{G}_I^0$

```{r Plot_GI0_mean, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for GI0, with $L= 2$, $\\mu=5$ and $\\alpha=-300$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 100
mu <- 5
L <- 2
alpha <- -300

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gI0(mu, alpha, L)
calculate_entropy <- function(sample_sizes, R, mu, alpha, L) {
  # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
    #"Van Es" = van_es_estimator,
    #"Correa" = correa_estimator,
    #"Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
    #"Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
    #"Al OmariA " = al_omari_estimator_m,
    #"Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator, 
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
   # "Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al O Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, alpha, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, alpha, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
  geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
  annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```
```{r Plot_GI0_mean_p, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for GI0, with $L= 2$, $\\mu=5$ and $\\alpha=-1000$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 100
mu <- 5
L <- 2
alpha <- -1000

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gI0(mu, alpha, L)
calculate_entropy <- function(sample_sizes, R, mu, alpha, L) {
  # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
    #"Van Es" = van_es_estimator,
    #"Correa" = correa_estimator,
    #"Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
    #"Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
    #"Al OmariA " = al_omari_estimator_m,
    #"Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator, 
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
   # "Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al O Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, alpha, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, alpha, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
  geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
  annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

```{r Plot_GI0_mean_y, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for GI0, with $L= 5$, $\\mu=10$ and $\\alpha=-400$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c( 25, 49, 81, 121)
R <- 100
mu <- 10
L <- 5
alpha <- -400

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gI0(mu, alpha, L)
calculate_entropy <- function(sample_sizes, R, mu, alpha, L) {
  # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
    #"Van Es" = van_es_estimator,
    #"Correa" = correa_estimator,
    #"Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
    #"Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
    #"Al OmariA " = al_omari_estimator_m,
    #"Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator, 
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
   # "Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al O Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, alpha, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, alpha, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
  geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
  annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

In the figure we can observe that when the parameter $\alpha$ goes to $-\infty$, the entropy of $\mathcal{G}_I^0$ is close to the entropy of $\Gamma_{\text{SAR}}$.

```{r Plot_GI0_mean_1, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for GI0, with $L= 1$,  $\\mu=1$ and $\\alpha=-300$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 100
mu <- 1
L <- 1
alpha <- -300

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gI0(mu, alpha, L)
calculate_entropy <- function(sample_sizes, R, mu, alpha, L) {
  # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
    #"Van Es" = van_es_estimator,
    #"Correa" = correa_estimator,
    #"Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
    #"Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
    #"Al OmariA " = al_omari_estimator_m,
    "Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator, 
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
   # "Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al O Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, alpha, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 100)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, alpha, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
  geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
  annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

```{r Simulated_data_gi0, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 100

# Number of bootstrap replications
B <- 100
mu_values <- c(1, 3, 5, 20)
alpha <- -20
L <- 1

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Yee" = yee_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/results_gi0.Rdata")

```

```{r Plot_bias_mse_gi0, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=1$, $\\alpha=-20$.", fig.width=9, fig.height=5}

load("./Data/results_gi0.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari","Yee",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{H}_{C}$"),#parse(text=TeX("$H_{2}$")),
                             "Correa Bootstrap" = expression("$\\widetilde{H}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{H}_{E}$"),
                            "Al Omari" = expression("$\\widehat{H}_{AO}$"),
                            "Yee" = expression("$\\widehat{H}_{Y}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{H}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{H}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 2, nrow = 2)

print(combined_plot_gi0)

```

```{r Simulated_data_gi0_2, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 100

# Number of bootstrap replications
B <- 100
mu_values <- c(1, 3, 5, 20)
alpha <- -20
L <- 2

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Yee" = yee_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/results_gi0_2.Rdata")

```

```{r Plot_bias_mse_gi0_2, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=2$, $\\alpha=-20$.", fig.width=9, fig.height=5}

load("./Data/results_gi0_2.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari","Yee",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{H}_{C}$"),#parse(text=TeX("$H_{2}$")),
                             "Correa Bootstrap" = expression("$\\widetilde{H}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{H}_{E}$"),
                            "Al Omari" = expression("$\\widehat{H}_{AO}$"),
                            "Yee" = expression("$\\widehat{H}_{Y}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{H}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{H}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 2, nrow = 2)

print(combined_plot_gi0)

```

```{r Simulated_data_gi0_3, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c(9, 25, 49, 81, 121)

# Number of replications
R <- 100

# Number of bootstrap replications
B <- 100
mu_values <- c(1, 3, 5, 20)
alpha <- -20
L <- 8

estimators <- list(
  "Correa" = correa_estimator,
  "Ebrahimi" = ebrahimi_estimator,
  "Al Omari" = al_omari_1_estimator,
  "Yee" = yee_estimator,
  "Correa Bootstrap" = bootstrap_correa_estimator,
  "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
  "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
)

calculate_results_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_bias_mse_gi0(sample_sizes, R, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0 <- calculate_results_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)

save(results_gi0, file = "./Data/results_gi0_3.Rdata")

```

```{r Plot_bias_mse_gi0_3, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Bias and MSE of entropy estimators for  $G_I^0$, $L=8$, $\\alpha=-20$.", fig.width=9, fig.height=5}

load("./Data/results_gi0_2.Rdata")

estimators_to_plot <- c("Correa", "Ebrahimi", "Al Omari","Yee",  "Correa Bootstrap", "Ebrahimi Bootstrap", "Al Omari Bootstrap" )
  latex_estimator_names <- c("Correa" = expression("$\\widehat{H}_{C}$"),#parse(text=TeX("$H_{2}$")),
                             "Correa Bootstrap" = expression("$\\widetilde{H}_{C}$"),
                            "Ebrahimi" = expression("$\\widehat{H}_{E}$"),
                            "Al Omari" = expression("$\\widehat{H}_{AO}$"),
                            "Yee" = expression("$\\widehat{H}_{Y}$"),
                            "Ebrahimi Bootstrap" = expression("$\\widetilde{H}_{E}$"),
                            "Al Omari Bootstrap" = expression("$\\widetilde{H}_{AO}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_gi0 <- generate_plot_gi0_esp(results_gi0, mu_values, selected_estimators_latex, ncol = 2, nrow = 2)

print(combined_plot_gi0)

```