### 2024-01-05 Code/R/RMarkdown/plots_mean.Rmd
codes test

---
title: "Report: Exploring Non-Parametric Entropy Estimators and Bootstrap Enhancements"
author: ""
date: ""
output: pdf_document

# output: 
#   pdf_document:
#     toc: true
#     toc_depth: 2
#html_document
header-includes:
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \usepackage{amsmath}
  - \usepackage{mathabx}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
#library(RColorBrewer)
library(ggsci)
#library(wesanderson)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

source("../MainFunctions/gamma_sar_sample.r")
source("../MainFunctions/entropy_gamma_sar.r")
source("../MainFunctions/entropy_gI0.r")
source("../MainFunctions/gi0_sample.r")

source("../MainFunctions/van_es_estimator.r")
source("../MainFunctions/correa_estimator.r")
source("../MainFunctions/adaptive_correa_estimator.r")
source("../MainFunctions/ebrahimi_estimator.r")
source("../MainFunctions/noughabi_arghami_estimator.r")
source("../MainFunctions/vasicek_estimator.r")
source("../MainFunctions/al_omari_1_estimator.r")
source("../MainFunctions/al_omari_estimator_m.r")
source("../MainFunctions/al_omari_2_estimator.r")
source("../MainFunctions/yee_estimator.r")

source("../MainFunctions/bootstrap_van_es_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator_opt.r")
source("../MainFunctions/bootstrap_ebrahimi_estimator.r")
source("../MainFunctions/bootstrap_noughabi_arghami_estimator.r")
source("../MainFunctions/bootstrap_vasicek_estimator.r")
source("../MainFunctions/bootstrap_al_omari_1_estimator.r")
source("../MainFunctions/bootstrap_al_omari_2_estimator.r")
source("../MainFunctions/bootstrap_yee_estimator.r")
source("../MainFunctions/bootstrap_yee_estimator_optimized.r")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Programs/functions_sample_bias_mse.R")#

set.seed(1234567890, kind = "Mersenne-Twister")

```

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.pos = 'H')
```

\newcommand{\bias}{\operatorname{Bias}}

# Mean Entropy for Estimators Gamma SAR

```{r Plot_mean1, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for Gamma SAR $L= 1$ and $\\mu=1$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121, 225)
R <- 10
mu <- 1
L <- 1

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gamma_sar_sample(L, mu, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gamma_sar(L, mu)
calculate_entropy <- function(sample_sizes, R, mu, L) {
 # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
     "Correa" = correa_estimator,
     "Ebrahimi" = ebrahimi_estimator,
     "Al Omari 1" = al_omari_1_estimator,
      "Yee" = yee_estimator,
      "Correa Bootstrap" = bootstrap_correa_estimator,
     "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
     "Al Omari 1 Bootstrap" = bootstrap_al_omari_1_estimator
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 10)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
       annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

```{r Plot_mean2, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy Gamma SAR for $L= 2$ and $\\mu=5$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 10
mu <- 5
L <- 2

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gamma_sar_sample(L, mu, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gamma_sar(L, mu)
calculate_entropy <- function(sample_sizes, R, mu, L) {
 # true_entropy <- entropy_gamma_sar(L, mu)

# Define a list of estimators
  estimators <- list(
   # "Van Es" = van_es_estimator,
   # "Correa" = correa_estimator,
   # "Adp Correa" = adaptive_correa_estimator,
    "Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
   # "Vasicek" = vasicek_estimator,
    "Al Omari 1" = al_omari_1_estimator,
   #"Al OmariA " = al_omari_estimator_m,
    #"Al Omari 2" = al_omari_2_estimator,
    "Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator,
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    "Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
    #"Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 10)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, L)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
       annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

## Results for  Nonparametric Estimators with  $\mathcal{G}_I^0$

```{r Plot_GI0_mean_p, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Mean entropy for GI0, with $L= 2$, $\\mu=5$ and $\\alpha=-1000$.", fig.width=6,fig.height=5}
set.seed(1234567890, kind = "Mersenne-Twister")

sample_sizes <- c(9, 25, 49, 81, 121)
R <- 10
mu <- 10
L <- 3
alpha <- -1000

# Function to generate samples for a given sample size and replication
generate_samples <- function(sample_size, replication, mu, alpha, L) {
  samples <- vector("list", replication)
  for (r in 1:replication) {
    samples[[r]] <- gi0_sample(mu, alpha, L, sample_size)
  }
  return(samples)
}

# Function to calculate Variance for both non-parametric and bootstrap estimators
true_entropy <- entropy_gamma_sar(L, mu)
calculate_entropy <- function(sample_sizes, R, mu, alpha, L) {
  # true_entropy <- entropy_gamma_sar(L, mu)

  # Define a list of estimators
  estimators <- list(
    #"Van Es" = van_es_estimator,
    #"Correa" = correa_estimator,
    #"Adp Correa" = adaptive_correa_estimator,
    #"Ebrahimi" = ebrahimi_estimator,
    #"Noughabi Arghami" = noughabi_arghami_estimator,
    #"Vasicek" = vasicek_estimator,
    #"Al Omari 1" = al_omari_1_estimator,
    #"Al OmariA " = al_omari_estimator_m,
    #"Yee" = yee_estimator,
    #"Van Es Bootstrap" = bootstrap_van_es_estimator,
    "Correa Bootstrap" = bootstrap_correa_estimator,
    #"Correa_opt Bootstrap" = bootstrap_correa_estimator_opt, 
    #"Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator,
    #"Noughabi Arghami Bootstrap" = bootstrap_noughabi_arghami_estimator,
    #"Vasicek Bootstrap" = bootstrap_vasicek_estimator,
    "Al Omari Bootstrap" = bootstrap_al_omari_1_estimator
    #"Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator
    #"Yee Bootstrap" = bootstrap_yee_estimator,
    #"Yee_opt Bootstrap" = bootstrap_yee_estimator_optimized
  )

  output <- data.frame(SampleSize = integer(0), Estimator = character(0), MeanEntropy = numeric(0))

  for (ssize in sample_sizes) {
    samples <- generate_samples(ssize, R, mu, alpha, L)

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      v.entropy <- numeric(R)

      for (r in 1:R) {
        sample <- samples[[r]]

        if (grepl("Bootstrap", estimator_name)) {
          v.entropy[r] <- estimator(sample, B = 10)
        } else {
          v.entropy[r] <- estimator(sample)
        }
      }

      mean_entropy <- mean(v.entropy)
      output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name, MeanEntropy = round(mean_entropy, 5)))

    }
  }

  return(output)
}

entropy_data <-calculate_entropy(sample_sizes, R, mu, alpha, L)
#cat("entropy", entropy_data)

ggplot(entropy_data, aes(x = SampleSize, y = MeanEntropy, color = Estimator)) +
  geom_point(size = 2) +
  geom_line(aes(group = Estimator), linetype = "solid", linewidth = 0.5) +
  geom_line(aes(y = true_entropy), linetype = "solid", color = "black") + 
  labs(title = "",
       x = "Sample Size", y = "Mean Entropy") +
  annotate("text", x = Inf, y = Inf, label = parse(text = sprintf("mu == %s", mu)), hjust = 1.08, vjust = 1.3, size = 3)

```

<!-- ```{r Plot_CI1, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.show="hold", fig.align="center",  fig.cap="Confidence Intervals for $L= 2$ and $\\mu=5$.", fig.width=6,fig.height=5} -->
<!-- set.seed(1234567890, kind = "Mersenne-Twister") -->

<!-- sample_sizes <- c(9, 25, 49, 81, 121, 500) -->
<!-- R <- 100 -->
<!-- mu <- 5 -->
<!-- L <- 2 -->
<!-- alpha <- -300 -->

<!-- # Function to generate samples for a given sample size and replication -->
<!-- generate_samples_gi0 <- function(sample_size, replication, mu, alpha, L) { -->
<!--   samples <- vector("list", replication) -->
<!--   for (r in 1:replication) { -->
<!--     samples[[r]] <- gi0_sample(mu, alpha, L, sample_size) -->
<!--   } -->
<!--   return(samples) -->
<!-- } -->

<!-- true_entropy <- entropy_gamma_sar(L, mu) -->
<!-- # Function to calculate Confidence Intervals for both non-parametric and bootstrap estimators -->
<!-- calculate_confidence_intervals <- function(sample_sizes, R, mu,alpha, L, confidence_level = 0.95) { -->
<!--   true_entropy <- entropy_gamma_sar(L, mu) -->

<!--   # Define a list of estimators -->
<!--   estimators <- list( -->
<!--       "Correa Bootstrap" = bootstrap_correa_estimator -->
<!--      #"Ebrahimi Bootstrap" = bootstrap_ebrahimi_estimator -->
<!--     # "Al Omari 1 Bootstrap" = bootstrap_al_omari_1_estimator -->
<!--     # "Al Omari 2 Bootstrap" = bootstrap_al_omari_2_estimator -->
<!--   ) -->

<!--   output <- data.frame(SampleSize = integer(0), Estimator = character(0),  -->
<!--                        LowerCI = numeric(0), UpperCI = numeric(0)) -->

<!--   for (ssize in sample_sizes) { -->

<!--     samples <- generate_samples_gi0(ssize, R, mu, alpha, L) -->

<!--     for (estimator_name in names(estimators)) { -->
<!--       estimator <- estimators[[estimator_name]] -->
<!--       v.entropy <- numeric(R) -->

<!--       for (r in 1:R) { -->
<!--         sample <- samples[[r]] -->

<!--         if (grepl("Bootstrap", estimator_name)) { -->
<!--           v.entropy[r] <- estimator(sample, B = 10) -->
<!--         } else { -->
<!--           v.entropy[r] <- estimator(sample) -->
<!--         } -->

<!--       } -->

<!--       # Calculate Confidence Interval -->
<!--       se <- sd(v.entropy) -->
<!--       margin_of_error <- qt((1 + confidence_level) / 2, R - 1) * se / sqrt(R) -->
<!--       ci_lower <- mean(v.entropy) - margin_of_error -->
<!--       ci_upper <- mean(v.entropy) + margin_of_error -->

<!--       output <- rbind(output, data.frame(SampleSize = ssize, Estimator = estimator_name,  -->
<!--                                          LowerCI = round(ci_lower, 5), UpperCI = round(ci_upper, 5))) -->
<!--     } -->
<!--   } -->

<!--   return(output) -->
<!-- } -->

<!-- confidence_intervals_data <- calculate_confidence_intervals(sample_sizes, R, mu,alpha, L,) -->

<!-- combined_plot <- ggplot(confidence_intervals_data, aes(x = SampleSize, y = UpperCI, ymin = LowerCI, ymax = UpperCI, color = Estimator)) + -->
<!--   geom_line(aes(group = Estimator), linetype = "dashed") + -->
<!--   geom_ribbon(alpha = 0.2) + -->
<!--   geom_point(aes(y = (LowerCI + UpperCI) / 2), color = "#808080", size = 2) +  -->
<!--   geom_line(aes(x = SampleSize, y = (LowerCI + UpperCI) / 2, color = Estimator), linetype = "solid", size = 0.5) +   -->
<!--   geom_line(aes(y = true_entropy), linetype = "solid", color = "black") +   -->
<!--   labs(title = "", -->
<!--        x = "Sample Size", y = "Entropy")  -->

<!-- print(combined_plot) -->
<!-- ``` -->

```{r Simulated_data_test_gi0, echo=FALSE, message=FALSE}

set.seed(1234567890, kind = "Mersenne-Twister")
sample_sizes <- c( 25, 49, 81, 121, 200)
N<- 200
# Number of replications
R <- 100

# Number of bootstrap replications
B <- 2
mu_values <- c(1)
alpha <- -600
L <- 5

  estimators <- list(

    "$\\widetilde{H}_{C}$ " = bootstrap_correa_estimator,
    #"$\\widetilde{H}_{E}$ " = bootstrap_ebrahimi_estimator,
    "$\\widetilde{H}_{AO}$ " = bootstrap_al_omari_1_estimator
  )
 #true_entropy <- entropy_gamma_sar(L, mu) 
calculate_entropy_and_test_montecarlo <- function(sample_sizes, R, N, B, mu, alpha, L, estimators, true_entropy) {
  true_entropy <- entropy_gamma_sar(L, mu)
  output <- data.frame(
    SampleSize = integer(0),
    Estimator = character(0),
    TypeIErrorRate = numeric(0)
  )

  for (ssize in sample_sizes) {

    for (estimator_name in names(estimators)) {
      estimator <- estimators[[estimator_name]]
      reject_H0_count <- 0

      for (n_replication in 1:N) {

        #samples <- gi0_sample(mu, alpha, L, ssize)
        v.entropy <- numeric(R)
        for (r in 1:R) {
        # Generar muestras aleatorias para cada repetición y cada estimador
        samples <- gi0_sample(mu, alpha, L, ssize)

        if (grepl(" ", estimator_name)) {
          v.entropy[r] <- estimator(samples, B = B)
        } else {
          v.entropy[r] <- estimator(samples)
        }
      }

        # if (grepl(" ", estimator_name)) {
        #   entropy_values <- replicate(R, estimator(samples, B = B))
        # } else {
        #   entropy_values <- replicate(R, estimator(samples))
        # }
        mean_entropy <- mean(v.entropy)
      z_statistic <- sqrt(R) * (mean_entropy - true_entropy) / sd(v.entropy)
      p_value <- 2 * (1 - pnorm(abs(z_statistic)))
      alpha_t = 0.05

        # Calcular p-values para cada replicación
       # p_values <- 2 * (1 - pnorm(abs((mean(entropy_values) - true_entropy) / sd(entropy_values))))

        # Contar los casos donde se rechaza la hipótesis nula
        reject_H0_count <- reject_H0_count + sum(p_value < 0.05)
      }

      typeI_error_rate <- reject_H0_count / (N)

      output <- rbind(
        output,
        data.frame(
          SampleSize = ssize,
          Estimator = estimator_name,
          TypeIErrorRate = round(typeI_error_rate, 5)
        )
      )
    }
  }

  colnames(output) <- c("$n$", "Estimator", "Type I Error Rate")

  return(output)
}

calculate_results_test_gi0_montecarlo <- function(sample_sizes, R, N, B, mu_values, alpha, L, estimators) {
  results_list <- list()

  for (mu_val in mu_values) {

    results <- calculate_entropy_and_test_montecarlo(sample_sizes, R, N, B, mu_val, alpha, L, estimators)
    df <- as.data.frame(results)

    results_list[[as.character(mu_val)]] <- df
  }

  return(results_list)
}

results_gi0_1_montecarlo <- calculate_results_test_gi0_montecarlo(sample_sizes, R, N, B, mu_values, alpha, L, estimators)
print(results_gi0_1_montecarlo)
save(results_gi0_1_montecarlo, file = "./Data/results_gi0_1_montecarlo.Rdata")

# calculate_results_test_gi0 <- function(sample_sizes, R, B, mu_values, alpha, L, estimators) {
#   results_list <- list()
# 
#   for (mu_val in mu_values) {
#     
#     results <- calculate_entropy_and_test(sample_sizes, R, B, mu_val, alpha, L, estimators)
#     df <- as.data.frame(results)
# 
#     
#     results_list[[as.character(mu_val)]] <- df
#   }
# 
#   return(results_list)
# }
# 
# 
# results_gi0_1 <- calculate_results_test_gi0(sample_sizes, R, B, mu_values, alpha, L, estimators)
# 
# 
# save(results_gi0_1, file = "./Data/results_gi0_1.Rdata")

```

```{r Table_test_gi0_montecarlo, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
load("./Data/results_gi0_1_montecarlo.Rdata")

# Interpretación de rechazo o no rechazo
for (mu_val in mu_values) {
  results_data <- results_gi0_1_montecarlo[[as.character(mu_val)]]

  table_result <- kbl(
    results_data,
   # caption = paste("Hypothesis Testing for $G_I^0$, $\\mu =", mu_val, "$, $L=2$, $\\pha1=-1000$ , #$H_{\\Gamma_{\\text{SAR}}}=2.493$."),
    format = "latex",
    booktabs = TRUE,
    align = "c",
    escape = FALSE,
    digits = 4,
    label = "table_hipotesis"
  ) %>%
    collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")

  #cat("### Results for mu =", mu_val, "\n")
  print(table_result)

}
# # Imprimir los nombres de las columnas en los resultados
# print(names(results_gi0_1_montecarlo[[1]]))
# 
# # Verificar las columnas disponibles en los resultados
# available_columns <- unique(unlist(lapply(results_gi0_1_montecarlo, names)))
# 
# # Seleccionar solo las columnas deseadas si están presentes
# desired_columns <- c("SampleSize", "Estimator", "TypeIErrorRate")
# selected_columns <- intersect(available_columns, desired_columns)
# 
# results_data <- lapply(results_gi0_1_montecarlo, function(df) {
#   df[, selected_columns, drop = FALSE]
# })
# 
# # Combinar los resultados en un único dataframe
# results_combined <- do.call(rbind, results_data)
# 
# # Mostrar la tabla
# table_result <- kbl(
#   results_combined,
#   format = "latex",
#   booktabs = TRUE,
#   align = "c",
#   escape = FALSE,
#   digits = 4,
#   label = "table_hipotesis_montecarlo"
# ) %>%
#   collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle")
# 
# print(table_result)

```
<!-- ```{r Table_test_gi0, echo=FALSE, message=FALSE, warning=FALSE, results='asis'} -->
<!-- load("./Data/results_gi0_1.Rdata") -->

<!-- # Imprimir los nombres de las columnas en los resultados -->
<!-- print(names(results_gi0_1[[1]])) -->

<!-- # Seleccionar solo las columnas deseadas -->
<!-- results_data <- lapply(results_gi0_1, function(df) { -->
<!--   df[, c("SampleSize", "Estimator", "TypeIErrorRate")] -->
<!-- }) -->

<!-- # Combinar los resultados en un único dataframe -->
<!-- results_combined <- do.call(rbind, results_data) -->

<!-- # Mostrar la tabla -->
<!-- table_result <- kbl( -->
<!--   results_combined, -->
<!--   format = "latex", -->
<!--   booktabs = TRUE, -->
<!--   align = "c", -->
<!--   escape = FALSE, -->
<!--   digits = 4, -->
<!--   label = "table_hipotesis" -->
<!-- ) %>% -->
<!--   collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") -->

<!-- print(table_result) -->

<!-- ``` -->

<!-- ```{r Table_test_gi0, echo=FALSE, message=FALSE, warning=FALSE, results='asis'} -->

<!-- load("./Data/results_gi0_1.Rdata") -->

<!-- # Interpretación de rechazo o no rechazo -->
<!-- for (mu_val in mu_values) { -->
<!--   results_data <- results_gi0_1[[as.character(mu_val)]] -->

<!--   table_result <- kbl( -->
<!--     results_data, -->
<!--    # caption = paste("Hypothesis Testing for $G_I^0$, $\\mu =", mu_val, "$, $L=2$, $\\pha1=-1000$ , #$H_{\\Gamma_{\\text{SAR}}}=2.493$."), -->
<!--     format = "latex", -->
<!--     booktabs = TRUE, -->
<!--     align = "c", -->
<!--     escape = FALSE, -->
<!--     digits = 4, -->
<!--     label = "table_hipotesis" -->
<!--   ) %>% -->
<!--     collapse_rows(columns = 1:3, latex_hline = "major", valign = "middle") -->

<!--   cat("### Results for mu =", mu_val, "\n") -->
<!--   print(table_result) -->

<!--   # Interpretación de rechazo o no rechazo -->
<!--   cat(ifelse(nrow(results_data) > 0, -->
<!--              paste("For mu =", mu_val, ":\n"), -->
<!--              paste("No observations for mu =", mu_val, ":\n"))) -->

<!--   cat(ifelse(nrow(results_data) > 0, -->
<!--              paste( -->
<!--                "Sample Size ", results_data$`$n$`, " Estimator ", results_data$Estimator, -->
<!--                ifelse(results_data$RejectNull, -->
<!--                       " the null hypothesis is rejected.", -->
<!--                       " the null hypothesis is not rejected." -->
<!--                ), "\n" -->
<!--              ), -->
<!--              "\n")) -->
<!-- } -->

<!-- ``` -->