### 2024-02-20 Code/R/RMarkdown/Report_goodness-of-fit.Rmd
new update paper and code

---
title: "Goodness-of-fit test "
author: ""
date: ""
output: pdf_document

# output: 
#   pdf_document:
#     toc: true
#     toc_depth: 2
#html_document
header-includes:
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache=TRUE)
#knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)

library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(invgamma)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(ggsci)
library(wesanderson)
library(ggpubr)
library(patchwork)
#options(kableExtra.latex.load_packages = FALSE)
library(devtools)
#devtools::install_github("haozhu233/kableExtra")
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(geomtextpath)
library(latex2exp)
library(univariateML)
#install.packages('univariateML')
#univariateML::univariateML_models

library(fitdistrplus)
library(logspline)
theme_set(theme_bw()  +
            theme(text=element_text(family="serif"),
                  legend.position = "top")# Gridtop , right , bottom , or left#, panel.grid = element_blank()
)
# Helpful for latex tables
# library(xtable)
# options(xtable.caption.placement='top',
#         xtable.table.placement='!t',
#         xtable.include.rownames=F,
#         xtable.comment=F)
# #library(rgl)

source("../MainFunctions/gamma_sar_sample.r")
source("../MainFunctions/entropy_gamma_sar.r")
source("../MainFunctions/entropy_gI0.r")

source("../MainFunctions/van_es_estimator.r")
source("../MainFunctions/correa_estimator.r")
source("../MainFunctions/ebrahimi_estimator.r")
source("../MainFunctions/noughabi_arghami_estimator.r")
source("../MainFunctions/vasicek_estimator.r")
source("../MainFunctions/al_omari_1_estimator.r")
source("../MainFunctions/al_omari_2_estimator.r")

source("../MainFunctions/bootstrap_van_es_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator.r")
source("../MainFunctions/bootstrap_ebrahimi_estimator.r")
source("../MainFunctions/bootstrap_noughabi_arghami_estimator.r")
source("../MainFunctions/bootstrap_vasicek_estimator.r")
source("../MainFunctions/bootstrap_al_omari_1_estimator.r")
source("../MainFunctions/bootstrap_al_omari_2_estimator.r")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Programs/functions_sample_bias_mse.R")#

set.seed(1234567890, kind = "Mersenne-Twister")

```

The p-values of a Kolmogorov-Smirnov-Test (KS-Test) with estimated parameters can be quite wrong because the p-value does not take the uncertainty of the estimation into account. So unfortunately, you can't just fit a distribution and then use the estimated parameters in a Kolmogorov-Smirnov-Test to test your sample. There is a normality test called Lilliefors test which is a modified version of the KS-Test that allows for estimated parameters.

```{r Plot1, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt"}

load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "skyblue", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
```

```{r 2, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Simulation of entropy for Gamma SAR"}
load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")
x1<-as.vector(cv_values)

descdist(x1, discrete = FALSE)

fit.weibull <- fitdist(x1, "weibull")
plot(fit.weibull)
fit.weibull$aic
fit.norm <- fitdist(x1, "norm")
fit.norm$aic
plot(fit.norm)
```

```{r}
load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")
x1<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x1)
mydata_dist
```

<!-- ```{r} -->
<!-- load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata") -->
<!-- x1<-as.vector(cv_values) -->
<!-- n.sims <- 100 -->

<!-- stats <- replicate(n.sims, { -->
<!--   r <- rweibull(n = length(x1) -->
<!--                 , shape= fit.weibull$estimate["shape"] -->
<!--                 , scale = fit.weibull$estimate["scale"] -->
<!--   ) -->
<!--   estfit.weibull <- fitdist(r, "weibull") # added to account for the estimated parameters -->
<!--   as.numeric(ks.test(r -->
<!--                      , "pweibull" -->
<!--                      , shape= estfit.weibull$estimate["shape"] -->
<!--                      , scale = estfit.weibull$estimate["scale"])$statistic -->
<!--   ) -->
<!-- }) -->

<!-- plot(ecdf(stats), las = 1, main = "KS-test statistic simulation (CDF)", col = "darkorange", lwd = 1.7) -->
<!-- grid() -->

<!-- ``` -->

```{r}
# load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")
# x1<-as.vector(cv_values)
# n.sims <- 5
# 
# stats_gamma <- replicate(n.sims, {
#   r <- rgamma(n = length(x1),
#               shape = fit.gamma$estimate["shape"],
#               scale = fit.gamma$estimate["scale"]
#   )
#   estfit_gamma <- fitdist(r, "gamma") # Añadido para tener en cuenta los parámetros estimados
#   as.numeric(ks.test(r,
#                      "pgamma",
#                      shape = estfit_gamma$estimate["shape"],
#                      scale = estfit_gamma$estimate["scale"])$statistic
#   )
# })
# # Suponiendo que has ejecutado la simulación con la distribución gamma y guardado los resultados en stats_gamma
# 
# # Gráfico de la ECDF de las estadísticas de la prueba KS simuladas
# plot(ecdf(stats_gamma), las = 1, main = "KS-test statistic simulation (CDF) - Gamma Distribution", col = "darkorange", lwd = 1.7)
# grid()

```

<!-- ```{r} -->
<!-- # Suponiendo que ya has cargado los datos y ajustado la distribución Rayleigh -->
<!-- load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata") -->
<!-- x1 <- as.vector(cv_values) -->
<!-- n.sims <- 10 -->

<!-- stats_rayleigh <- replicate(n.sims, { -->
<!--   r <- rrayleigh(n = length(x1), -->
<!--                  scale = fit_rayleigh$estimate["scale"] -->
<!--   ) -->
<!--   estfit_rayleigh <- fitdist(r, "rayleigh") # Añadido para tener en cuenta los parámetros estimados -->
<!--   as.numeric(ks.test(r, -->
<!--                      "prayleigh", -->
<!--                      scale = estfit_rayleigh$estimate["scale"] -->
<!--                )$statistic -->
<!--   ) -->
<!-- }) -->

<!-- plot(ecdf(stats_rayleigh), las = 1, main = "KS-test statistic simulation (CDF) - Rayleigh Distribution", col = "darkorange", lwd = 1.7) -->
<!-- grid() -->

<!-- ``` -->