### 2024-02-24 Code/R/RMarkdown/Report_goodness-of-fit.Rmd
final corrections, added heatmap for p-values

---
title: "Goodness-of-fit test "
author: ""
date: ""
output: pdf_document

# output: 
#   pdf_document:
#     toc: true
#     toc_depth: 2
#html_document
header-includes:
  - \usepackage{graphicx}
  - \usepackage{float}
  - \usepackage{bm}
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,cache=TRUE)
#knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)

library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(gtools)
library(stats4)
library(rmutil)
library(invgamma)
library(tidyverse)
library(scales)
library(RColorBrewer)
library(ggsci)
library(wesanderson)
library(ggpubr)
library(patchwork)
#options(kableExtra.latex.load_packages = FALSE)
library(devtools)
#devtools::install_github("haozhu233/kableExtra")
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(geomtextpath)
library(latex2exp)
library(univariateML)
library(viridis)
#install.packages('univariateML')
#univariateML::univariateML_models

library(fitdistrplus)
library(logspline)
theme_set(theme_bw()  +
            theme(text=element_text(family="serif"),
                  legend.position = "top")# Gridtop , right , bottom , or left#, panel.grid = element_blank()
)
# Helpful for latex tables
# library(xtable)
# options(xtable.caption.placement='top',
#         xtable.table.placement='!t',
#         xtable.include.rownames=F,
#         xtable.comment=F)
# #library(rgl)

source("../MainFunctions/gamma_sar_sample.r")
source("../MainFunctions/entropy_gamma_sar.r")
source("../MainFunctions/entropy_gI0.r")

source("../MainFunctions/van_es_estimator.r")
source("../MainFunctions/correa_estimator.r")
source("../MainFunctions/ebrahimi_estimator.r")
source("../MainFunctions/noughabi_arghami_estimator.r")
source("../MainFunctions/vasicek_estimator.r")
source("../MainFunctions/al_omari_1_estimator.r")
source("../MainFunctions/al_omari_2_estimator.r")

source("../MainFunctions/bootstrap_van_es_estimator.r")
source("../MainFunctions/bootstrap_correa_estimator.r")
source("../MainFunctions/bootstrap_ebrahimi_estimator.r")
source("../MainFunctions/bootstrap_noughabi_arghami_estimator.r")
source("../MainFunctions/bootstrap_vasicek_estimator.r")
source("../MainFunctions/bootstrap_al_omari_1_estimator.r")
source("../MainFunctions/bootstrap_al_omari_2_estimator.r")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../Programs/functions_sample_bias_mse.R")# 
source("../imagematrix.R")
source("../imagematrix1.R")
set.seed(1234567890, kind = "Mersenne-Twister")

```

The p-values of a Kolmogorov-Smirnov-Test (KS-Test) with estimated parameters can be quite wrong because the p-value does not take the uncertainty of the estimation into account. So unfortunately, you can't just fit a distribution and then use the estimated parameters in a Kolmogorov-Smirnov-Test to test your sample. There is a normality test called Lilliefors test which is a modified version of the KS-Test that allows for estimated parameters.

```{r Plot1, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt"}

load("../Programs/Data/results_data_Flevoland_cv_300_5_v2.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "skyblue", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
```

```{r 2, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Simulation of entropy for Gamma SAR"}
load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")
x1<-as.vector(cv_values)

descdist(x1, discrete = FALSE)

fit.weibull <- fitdist(x1, "weibull")
plot(fit.weibull)
fit.weibull$aic
# fit.norm <- fitdist(x1, "norm")
# fit.norm$aic
# plot(fit.norm)
```

```{r, eval=FALSE}
load("../Programs/Data/results_data_Flevoland_cv_300_5_v2.Rdata")
x1<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x1)
mydata_dist
```

```{r cv5, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="p values for CV window 5x5, Flevoland"}
load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata")
alpha <- 5.452
beta <- 2.061

pgamma_inv <- function(x, alpha, beta) {
  1 - pgamma(1/x, shape = alpha, rate = beta)
}

p_values_matrix <- matrix(NA, nrow = nrow(cv_values), ncol = ncol(cv_values))

for (i in 1:nrow(cv_values)) {
  for (j in 1:ncol(cv_values)) {
    cv_value <- cv_values[i, j]

    p_values_matrix[i, j] <- pgamma_inv(cv_value, alpha, beta)
  }
}

#save(p_values_matrix, file = "./Data/results_pvalue_Flevoland_cv_300_5.Rdata")

#source("../imagematrix.R")
source("../imagematrix1.R")
#hist(p_values_matrix)
#par(mfrow=c(1,2))
#plot(imagematrix(p_values_matrix ))
plot(imagematrix1(p_values_matrix ))
#plot(imagematrix(p_values_matrix >0.1))
```

```{r Plot2, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Density cv, window 7x7, Flevoland" }

load("../Programs/Data/results_data_Flevoland_cv_300_7.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "blue", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
source("../imagematrix.R")
#hist(cv_values)
```

```{r, eval=FALSE}
load("../Programs/Data/results_data_Flevoland_cv_300_7.Rdata")
x2<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x2)
mydata_dist
```
```{r cv7, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="p values for CV window 7x7, Flevoland"}
load("../Programs/Data/results_data_Flevoland_cv_300_7.Rdata")
alpha <- 5.762
beta <- 2.611

pgamma_inv <- function(x, alpha, beta) {
  1 - pgamma(1/x, shape = alpha, rate = beta)
}

p_values_matrix <- matrix(NA, nrow = nrow(cv_values), ncol = ncol(cv_values))

for (i in 1:nrow(cv_values)) {
  for (j in 1:ncol(cv_values)) {
    cv_value <- cv_values[i, j]

    p_values_matrix[i, j] <- pgamma_inv(cv_value, alpha, beta)
  }
}

#save(p_values_matrix, file = "./Data/results_pvalue_Flevoland_cv_300_5.Rdata")

source("../imagematrix.R")
#hist(p_values_matrix)
par(mfrow=c(1,2))
#plot(imagematrix(equalize(x)))
plot(imagematrix(p_values_matrix))
plot(imagematrix(p_values_matrix >0.1))
```

```{r, eval=FALSE}
load("../Programs/Data/results_data_Flevoland_cv_300_9.Rdata")
x2<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x2)
mydata_dist
```

```{r cv9, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="p values for CV window 9x9, Flevoland"}

#Los parametros fueron estimados con phitter.io
load("../Programs/Data/results_data_Flevoland_cv_300_9.Rdata")
alpha <- 3.58
beta <- 1.32

pgamma_inv <- function(x, alpha, beta) {
  1 - pgamma(1/x, shape = alpha, rate = beta)
}

p_values_matrix <- matrix(NA, nrow = nrow(cv_values), ncol = ncol(cv_values))

for (i in 1:nrow(cv_values)) {
  for (j in 1:ncol(cv_values)) {
    cv_value <- cv_values[i, j]

    p_values_matrix[i, j] <- pgamma_inv(cv_value, alpha, beta)
  }
}

#save(p_values_matrix, file = "./Data/results_pvalue_Flevoland_cv_300_5.Rdata")

source("../imagematrix.R")
#hist(p_values_matrix)
par(mfrow=c(1,2))
#plot(imagematrix(equalize(x)))
plot(imagematrix(p_values_matrix))
plot(imagematrix(p_values_matrix >0.1))
```

```{r Plot3, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Density cv, window 5x5, Ottawa" }

load("../Programs/Data/results_data_Ottawa_cv_512_5.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "red", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
```

```{r 5ottawa, eval=FALSE}
load("../Programs/Data/results_data_Ottawa_cv_512_5.Rdata")
x2<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x2)
mydata_dist
```

```{r cvOttawa, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="p values for CV window 5x5, Ottawa"}

#Los parametros fueron estimados con phitter.io
load("../Programs/Data/results_data_Ottawa_cv_512_5.Rdata")
alpha <- 5.29
beta <- 2.20

pgamma_inv <- function(x, alpha, beta) {
  1 - pgamma(1/x, shape = alpha, rate = beta)
}

p_values_matrix <- matrix(NA, nrow = nrow(cv_values), ncol = ncol(cv_values))

for (i in 1:nrow(cv_values)) {
  for (j in 1:ncol(cv_values)) {
    cv_value <- cv_values[i, j]

    p_values_matrix[i, j] <- pgamma_inv(cv_value, alpha, beta)
  }
}

#save(p_values_matrix, file = "./Data/results_pvalue_Flevoland_cv_300_5.Rdata")

source("../imagematrix.R")
#hist(p_values_matrix)
par(mfrow=c(1,2))
#plot(imagematrix(equalize(x)))
plot(imagematrix(p_values_matrix))
plot(imagematrix(p_values_matrix >0.1))
```

```{r Plotcqv, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Density cqv, window 5x5, Ottawa" }

load("../Programs/Data/results_data_Ottawa_cqv_512_5.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "green", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
```

```{r 5ottawa_cqv, eval=FALSE}
load("../Programs/Data/results_data_Ottawa_cqv_512_5.Rdata")
x2<-as.vector(cv_values)
mydata_dist <- univariateML::model_select(x2)
mydata_dist
```

```{r cqvOttawa_5, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="p values for CV window 5x5, Ottawa"}

#Los parametros fueron estimados con phitter.io
load("../Programs/Data/results_data_Ottawa_cv_512_5.Rdata")
alpha <- 3.58
beta <- 1.32

pgamma_inv <- function(x, alpha, beta) {
  1 - pgamma(1/x, shape = alpha, rate = beta)
}

p_values_matrix <- matrix(NA, nrow = nrow(cv_values), ncol = ncol(cv_values))

for (i in 1:nrow(cv_values)) {
  for (j in 1:ncol(cv_values)) {
    cv_value <- cv_values[i, j]

    p_values_matrix[i, j] <- pgamma_inv(cv_value, alpha, beta)
  }
}

#save(x, p_values_matrix, file = "./Data/results_pvalue_Flevoland_cv_300_5.Rdata")

source("../imagematrix.R")
hist(p_values_matrix)
par(mfrow=c(1,2))
plot(imagematrix(equalize(x)))
plot(imagematrix(p_values_matrix))
plot(imagematrix(p_values_matrix >0.05))
```
```{r Plotcqvmad, echo=FALSE, message=FALSE, warning=FALSE, out.width="80%", fig.show="hold", fig.align="center", fig.pos="hbt", fig.cap="Density cqv, window 5x5, Ottawa" }

load("../Programs/Data/results_data_Ottawa_cd_mad_matrix_512_5.Rdata")

ggplot() +
  geom_density(aes(x = as.vector(cv_values)), fill = "pink", color = "black") +
  labs(title = " ", x = "cv", y = "Density")
```

<!-- ```{r} -->
<!-- load("../Programs/Data/results_data_Flevoland_cv_300_5.Rdata") -->
<!-- x1<-as.vector(cv_values) -->
<!-- n.sims <- 100 -->

<!-- stats <- replicate(n.sims, { -->
<!--   r <- rweibull(n = length(x1) -->
<!--                 , shape= fit.weibull$estimate["shape"] -->
<!--                 , scale = fit.weibull$estimate["scale"] -->
<!--   ) -->
<!--   estfit.weibull <- fitdist(r, "weibull") # added to account for the estimated parameters -->
<!--   as.numeric(ks.test(r -->
<!--                      , "pweibull" -->
<!--                      , shape= estfit.weibull$estimate["shape"] -->
<!--                      , scale = estfit.weibull$estimate["scale"])$statistic -->
<!--   ) -->
<!-- }) -->

<!-- plot(ecdf(stats), las = 1, main = "KS-test statistic simulation (CDF)", col = "darkorange", lwd = 1.7) -->
<!-- grid() -->

<!-- ``` -->