### 2024-05-08 Text/Thesis/partes/acknowledgements.tex

\begin{acknowledgements}
\addchaptertocentry{\acknowledgementname} % Add the acknowledgements to the table of contents
This PhD project is the outcome of the work of many people: 
\\
Whenever fitting, specific acknowledgements are given inside the text or in footnotes, but here I would especially like to thank all of my colleagues at \\
In particular,

%\newpage
%\section*{Author's note}
%This thesis has been compressed for uploading and is missing hyperlinks.
%The original version, containing full-resolution images and hyperlinks, can be obtained from \url{http://users.ictp.it/~afantini/PhD_thesis/PhD_thesis_final_15feb2019.pdf} or by requesting it to \href{mailto:adr.fantini@gmail.com}{adr.fantini@gmail.com}.

\end{acknowledgements}