### 2025-03-15 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
Update paper with corrected comments

---
title: Rényi Entropy-Based Heterogeneity Detection in SAR Data
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default

author:
  #- id: 
  - name: Janeth Alpala, Abraão D.\ C.\ Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 50670-901
      #- name: Unknown affiliation
   # orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    #attributes:
    #  corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.\ Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 6140
      #- name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems. A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes. In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed. 
  In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task. 
  In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy. 
  In particular, we use a nonparametric estimator for the Rényi entropy and investigate some of its properties. 
  As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories. 
  The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy.
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR,  hypothesis tests]

#funding: 
 # statement: "The `quarto-ieee` "
pageheader:
  left: Journal XXX, Month Year
  right: #'D. Folio:  A Sample Article Using quarto-ieee'

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   #- \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{mathtools}
  # - \usepackage[utf8]{inputenc}
   - \usepackage{hyperref}
   #- \hypersetup{draft} #Desactiva enlaces y referencias cruzadas
   - \usepackage{float}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{hyperref}
   - \hypersetup{
      colorlinks=true,
      linkcolor=black,
      filecolor=black,
      citecolor=black,
      urlcolor=black}

#bibliography: references.bib
bibliography: ../../Common/references.bib

# execute:
#   echo: false
#   eval: true

---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

# Configurar CRAN
options(repos = c(CRAN = "https://cran.rstudio.com/"))

# install and load packages only if they are missing
install_and_load <- function(packages) {
  missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
  if (length(missing_packages)) {
    install.packages(missing_packages, dependencies = TRUE)
  }
  invisible(lapply(packages, library, character.only = TRUE))
}
#
#  packages
required_packages <- c(
 "ggplot2", "reshape2", "knitr", "pandoc", "gridExtra", 
  "gtools", "stats4", "rmutil", "scales", "tidyr", "invgamma", 
  "tidyverse", "RColorBrewer", "ggsci", "carData", "ggpubr",  "patchwork", "dplyr", 
  "kableExtra", "ggthemes", "latex2exp", "e1071", "viridis", "nortest", "bookdown"
)

# Install and load only missing packages
install_and_load(required_packages)

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom"))

# External functions
source("./Code/gamma_sar_sample.R")
source("./Code/gi0_sample.R")
source("./Code/al_omari_1_estimator.R")
source("./Code/bootstrap_al_omari_1_estimator.R")
source("./Code/renyi_entropy_estimator_v1.R")
source("./Code/bootstrap_renyi_entropy_estimator_v1.R")
source("./Code/entropy_renyi_gamma_sar.R")
source("./Code/functions_sample_bias_mse.R")
source("./Code/functions_sample_bias_mse_1.R")
#source("./Code/read_ENVI_images.R")

```

# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications\ [@Yu2023;@Akbarizadeh2012;@Mondini2021]. It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions to facilitate global Earth monitoring\ [@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect\ [@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution has been shown to be an effective law for SAR intensity data, as it captures different levels of roughness. A special case is the Gamma distribution, which arises when speckle is fully developed, indicating homogeneous regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

Entropy measures have gained attention as valuable statistical tools for analyzing SAR data, with applications in edge detection\ [@Nascimento2014], segmentation\ [@Nobre2016], classification\ [@Cassetti2022], and noise reduction\ [@Chan2022].
Traditionally, Shannon entropy\ [@Shannon1948] has been widely used to quantify uncertainty and disorder in data. However, this study explores an alternative information measure: Rényi entropy—a generalization of Shannon’s formulation. 
This more flexible approach provides additional insights for identifying heterogeneity, making it a promising tool for enhancing SAR image analysis.

In this work, we propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy in a given region significantly differs from its expected theoretical value under the assumption of homogeneity.
Compared to our previous approach  using Shannon entropy\ [@Frery2024], the Rényi-based approach improves the detection of heterogeneity in real SAR images.

The remainder of this article is organized as follows.
Section\ \ref{sec:pre} outlines an overview of statistical models for SAR intensity data and introduces Rényi entropy.
Section\ \ref{sec:met} describes the proposed hypothesis test and the formulation of the test statistic. 
Section\ \ref{sec:app} evaluates the performance of the test using real SAR data.
Finally, the concluding remarks are drawn in Section\ \ref{sec:conclusion}.

# PRELIMINARIES {#sec:pre}

## Statistical Models

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness\ [@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{align}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    &= \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z) \label{E:gamma1}
    \intertext{and}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    &= \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}
    \notag \\
    &\quad \times
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{align}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by\ [@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

## Rényi Entropy

Introduced by Alfréd Rényi in 1961\ [@renyi1961measures], this measure generalizes several well-known entropies, including the Shannon entropy\ [@Ribeiro2021]. For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda$, with $\lambda > 0$ and $\lambda \neq 1$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L) \\  + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr]
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{\,1 - \lambda\,}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
The derivation of this result is given in the Appendix.

@fig-plot presents the Rényi entropy of $\mathcal{G}^0_I$ as a function of $\mu$ for different $\alpha$ values. As $\alpha \to -\infty$, it approaches the Rényi entropy of $\Gamma_{\text{SAR}}$, which is aligned with the fact that $\Gamma_{\text{SAR}}$ is a limiting case of the $\mathcal{G}^0_I$ model.

```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(\\mathcal{G}^0_I)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=4.5, fig.height=3.5}

entropy_renyi_gamma_sar <- function(L, mu, lambda) {
  entropy <- (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mu / L)
  return(entropy)
}

entropy_GI0_renyi <- function(alpha, mu, L, lambda) {
  if (lambda <= 0 || lambda == 1) {
    stop("Lambda must be greater than 0 and not equal to 1.")
  }

  gamma <- -mu * (alpha + 1)
  if (any(gamma <= 0)) {
    stop("Gamma must be positive. Check the values of mu and alpha.")
  }

  a <- lambda * (L - 1) + 1
  b <- lambda * (-alpha + 1) - 1
  ab_sum <- lambda * (L - alpha)

  if (any(a <= 0) || any(b <= 0) || any(ab_sum <= 0)) {
    stop("Arguments of the Gamma functions must be positive. Check the values of lambda, L, and alpha.")
  }

  term1 <- log(gamma / L)

  term2 <- lambda * (lgamma(L - alpha) - lgamma(-alpha) - lgamma(L))

  term3 <- lgamma(a)
  term4 <- lgamma(b)
  term5 <- lgamma(ab_sum)

  numerator <- term2 + term3 + term4 - term5

  entropy <- term1 + numerator / (1 - lambda)

  return(entropy)
}

L <- 8
alphas <- c(-3, -8, -20, -1000)
alpha_labels <- c(expression(italic(alpha) == -3), 
                  expression(italic(alpha) == -8), 
                  expression(italic(alpha) == -20), 
                  expression(italic(alpha) == -1000))

mu <- seq(0.1, 10, length.out = 500)
lambda <- 0.8  # Fixed lambda

muEntropy <- data.frame()

for (alpha in alphas) {
  entropies_GI0 <- entropy_GI0_renyi(alpha, mu, L, lambda)
  muEntropy <- rbind(muEntropy, data.frame(mu = mu, Entropy = entropies_GI0, alpha = as.factor(alpha)))
}

muEntropy.molten <- melt(muEntropy, id.vars = c("mu", "alpha"), value.name = "Entropy")

entropies_gamma <- entropy_renyi_gamma_sar(L, mu, lambda)

Entropy_gamma <- data.frame(mu, Entropy_Gamma = entropies_gamma)

Entropy_gamma.molten <- melt(Entropy_gamma, id.vars = "mu", value.name = "Entropy_Gamma")

ggplot() +

  geom_line(data = Entropy_gamma.molten, aes(x = mu, y = Entropy_Gamma), color = "black", 
            linetype = "solid", linewidth = 1.5) +

  geom_line(data = muEntropy.molten, aes(x = mu, y = Entropy, color = alpha), 
            linetype = "longdash", linewidth = 1) +

  annotate("text", x = max(mu) + 0.2, y = max(Entropy_gamma.molten$Entropy_Gamma), 
           label = TeX("${italic(H)}_{\\lambda}(\\Gamma_{\\tiny{SAR}})$"), 
           vjust = 1.6, hjust = 0.8, color = "black",linewidth = 0.2) +

  theme_minimal() +

  scale_color_manual(values = pal_jama()(7)[2:5], labels = alpha_labels) +
 # scale_color_manual(values = brewer.pal(4, "Dark2"), labels = alpha_labels) +
  # Labels and titles
  labs(color = "Roughness", 
       x = expression(mu), 
       y = "Rényi Entropy", 
       linetype = NULL) +

  theme(text = element_text(family = "serif"),
        legend.position = "bottom") +

  coord_cartesian(xlim = c(0, 10), ylim = c(min(muEntropy.molten$Entropy, Entropy_gamma.molten$Entropy_Gamma), 
                                            max(muEntropy.molten$Entropy, Entropy_gamma.molten$Entropy_Gamma) + 0.5))

```

## Nonparametric Estimation of Rényi Entropy

Nonparametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999; @AlOmari2019]. Recently, Al-Labadi et al.\ [@AlLabadi2024] proposed a nonparametric estimator for Rényi entropy following this approach.

Let $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ be an independent and identically distributed random sample of size $n$ from a distribution $F$, and let $Z_{(1)} \leq Z_{(2)} \leq \dots \leq Z_{(n)}$ denote its order statistics.
The $m$-spacing density estimator is defined as:
$$
f_n(Z_{(i)}) = \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}},
$$
where $Z_{(i-m)} = Z_{(1)}$ when $i \leq m$, and $Z_{(i+m)} = Z_{(n)}$ if $i \geq n - m$. 
The coefficient $c_i$ is given by:
$$
c_i = 
\begin{cases}
\frac{m + i - 1}{m}, & \text{if } 1 \leq i \leq m, \\%[6pt]
2, & \text{if } m+1 \leq i \leq n - m, \\%[6pt]
\frac{n + m - i}{m}, & \text{if } n - m + 1 \leq i \leq n.
\end{cases}
$$

Following Vasicek [@vasicek1976test] and Ebrahimi et al. [@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, the Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We choose to use the heuristic formula for spacing, $m=\left[\sqrt{n}+0.5\right]$.

# PROPOSED METHODOLOGY {#sec:met}

## On $\lambda$ optimality for a sample size

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$.
To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$. Lower MSE and bias indicate better performance of the estimator in approximating the true entropy.

Based on the results, we find that the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias. This choice is the same for $L > 1$. However, for $L = 1$, the optimal $\lambda$ tends to be higher (e.g., $\lambda = 3$) to achieve good results. @fig-plotf illustrates the case for $L = 5$.
```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

data <- data.frame(
  Lambda = c(0.9,  0.85, 0.99, 1.1, 1.5),
  Bias = c(0.00158,  -0.00250, 0.02077, 0.03751, 0.06512),
  MSE = c(0.01273,  0.01441, 0.01653, 0.01697, 0.01906)
)

data <- data[order(data$Lambda), ]

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#00AFBB", linewidth = 1.0) +  
  geom_point(color = "#00AFBB", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#E69F00", linewidth = 1.0) +  
  geom_point(color = "#E69F00", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

grid.arrange(bias_plot, mse_plot, nrow = 2)

```

## Bootstrap Correction for Entropy Estimator

Following Refs.\ [@Frery2024;@Alpala2024], we also refine the non-parametric entropy estimator $\widehat{H}_{\lambda}$ in \eqref{eq:est_R} using the bootstrap technique to reduce its bias and apply the improved version, denoted as $\widetilde{H}_{\lambda}$:
$$
\widetilde{H}_{\lambda} = 2\widehat{H}_{\lambda}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_{\lambda}(\bm{Z}^{(b)}),
$$
where $B$ is the number of bootstrap replications, and $\bm{Z}^{(b)}$ is the $b$-th resampled dataset obtained by drawing $n$ observations with replacement from $\bm{Z}$.

A Monte Carlo study with $1,000$ samples from $\Gamma_{\text{SAR}}$ ($\mu = 1$, $L=5$, $\lambda = 0.9$), using $B = 200$ replications, confirms that the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$  outperforms the original $\widehat{H}_{\lambda}$ by reducing both bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.

```{r Simulated_data_bias_B1, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

file_name <- "./Data/results_renyi_B1.Rdata"

if (file.exists(file_name)) {

  load(file_name)
  message("Loaded existing results from results_renyi_B1.Rdata")
} else {
  # Parameters
  sample_sizes <- c(9, 25, 49, 81, 121)
  R <- 500        
  B <- 200        
  mu <- 1        
  L <- 5          
  alpha_values <- c(0.9)

  estimators <- list(
    "Renyi Estimator" = renyi_entropy_estimator_v1,
    "Renyi Estimator Bootstrap" = bootstrap_renyi_entropy_estimator_v1
  )

  results <- calculate_bias_mse_r(sample_sizes, R, B, mu, L, alpha_values, estimators)

  save(results, file = file_name)
  message("Simulations completed and results saved.")
}
```

```{r fig-Plot_bias_msef3, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%", fig.pos="hbt",  fig.cap="Bias and MSE of the Rényi entropy estimators for $\\Gamma_{\\text{SAR}}$.", fig.width=7, fig.height=5}

load("./Data/results_renyi_B1.Rdata")

alpha_values <- 0.9
estimators_to_plot <- c("Renyi Estimator", "Renyi Estimator Bootstrap")
latex_estimator_names <- c("Renyi Estimator" = expression("$\\widehat{italic(H)}_{\\lambda}$"),# 
                           "Renyi Estimator Bootstrap" = expression("$\\widetilde{italic(H)}_{\\lambda}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_renyi <- generate_plot_renyi(results, alpha_values, selected_estimators_latex, ncol = 1, nrow = 1)

print(combined_plot_renyi)

```
For subsequent simulations we use the $\widetilde{H}_{\lambda}$ estimator.

## Hypothesis Testing

We test whether the observed data come from a homogeneous ($\Gamma_{\text{SAR}}$) or a heterogeneous ($\mathcal{G}^0_I$) region, as follows:
\begin{equation}\label{eq:hypothesis_test}
\begin{cases}
\mathcal{H}_0: \mathbb{E}[\widetilde{H}_{\lambda}] = H_{\lambda}(\Gamma_{\text{SAR}}), & \text{(Homogeneous region)} \\[6pt]
\mathcal{H}_1: \mathbb{E}[\widetilde{H}_{\lambda}] = H_{\lambda}(\mathcal{G}^0_I), & \text{(Heterogeneous region)}
\end{cases}
\end{equation}

Under $\mathcal{H}_0$, the expected value of the entropy estimator should match the theoretical $H_{\lambda}(\Gamma_{\text{SAR}})$. If the estimated entropy significantly deviates from its expected value, we reject $\mathcal{H}_0$, indicating the presence of heterogeneity.

Since homogeneity corresponds to $\alpha \to -\infty$ in the $\mathcal{G}^0_I$ model, classical inference methods are inapplicable. Instead, we propose a test statistic based on the Rényi entropy estimator, comparing it with its theoretical expectation under $\mathcal{H}_0$ to distinguish between homogeneous and heterogeneous areas.

## The Proposed Test

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing \eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation \eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\  
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\}.
\end{multline}
where $\ln \widehat{\mu}={\textstyle\frac{1}{n}\sum_{i=1}^n Z_{i}}$ is the natural logarithm of the sample mean.

When the null hypothesis holds, this test statistic is expected to be close to zero.
For illustrative purposes, @fig-density_entropyR shows the empirical density of $S(\bm{Z}; L)$, based on $10,000$ simulations of the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 10000
mu <- 1
B <- 50

lambda <- 0.9

sample.size <- c(49, 81, 121)
L_values <- c(5, 18)

all_summary_stats <- list()
all_TestStatistics <- list()

# 
for (L in L_values) {

  file_name <- paste0("./Data/resultsR_", L, ".Rdata")

  if (file.exists(file_name)) {
    load(file_name)
    message(paste("Loaded existing data for L =", L))
  } else {

    TestStatistics1 <- list()  
    summary_stats <- data.frame(
      LValue = character(),
      SampleSize = numeric(),
      Mean = numeric(),
      SD = numeric()
      #Variance = numeric(),
      #Skewness = numeric(),
     # Kurtosis = numeric(),
      #adpvalue = numeric()
    )

    # Para cada tamaño de muestra
    for (s in sample.size) {
      TestStat1 <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat1[r] <- TestStat <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) -((lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                                                                        (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mean(z))-log(L))
      }

      TestStatistics1[[as.character(s)]] <- data.frame(
        "SampleSize" = rep(s, R), 
        "Test_Statistics" = TestStat1
      )

      mean_val <- mean(TestStat1)
      sd_val <- sd(TestStat1)
      #var_val <- var(TestStat1)
     # skewness_val <- skewness(TestStat1)
     # kurtosis_val <- kurtosis(TestStat1)
     # ad_p_value <- ad.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value

      summary_stats <- rbind(summary_stats, data.frame(
        LValue = as.character(L),
        SampleSize = s,
        Mean = mean_val,
        SD = sd_val
      #  Variance = var_val,
       # Skewness = skewness_val,
       # Kurtosis = kurtosis_val,
       # adpvalue = ad_p_value
      )) 
    }

    all_TestStatistics[[as.character(L)]] <- TestStatistics1
    all_summary_stats[[as.character(L)]] <- summary_stats

    save(all_TestStatistics, all_summary_stats, file = file_name)
    message(paste("Saved new results for L =", L))
  }
}

```

```{r fig-density_entropyR, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="47%", fig.pos="hbt",  fig.cap="Empirical densities of $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z}; L)$ under $\\mathcal{H}_0$.", fig.width=9, fig.height=4.0}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

x_limits <- c(-0.5, 0.5)  # Límites del eje X
x_breaks <- seq(-0.4, 0.4, by = 0.2)  # Ticks del eje X
y_limits <- c(0, 7)       # Límites del eje Y

selected_L_values <- c(5, 18)

all_plots <- list()

for (L in selected_L_values) {
  load(paste0("./Data/resultsR_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  combined_data$CurveOrder <- factor(combined_data$SampleSize, levels = rev(sample.size))

  combined_data$LegendOrder <- factor(combined_data$SampleSize, levels = sample.size)

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.5) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.2, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(
        x = expression("Test Statistic" ~ S[widetilde(italic(H))[lambda]](italic(bold(Z))* ";" **phantom(" ")* italic(L))), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 2)),
          axis.text = element_text(size = 16),     
        axis.title = element_text(size = 16),    
        legend.text = element_text(size = 16),   
        legend.title = element_text(size = 16)   
          )

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 2, nrow = 1) +
  plot_layout(guides = "collect")

print(combined_plot)

```
Under the asymptotic properties of the entropy estimator\ [@vasicek1976test], for sufficiently large samples, $S(\bm{Z}; L)$ follows an asymptotic normal distribution:
\begin{equation*}
S(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2})\,.
\end{equation*}
We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S = \mathbb{E}[S(\bm{Z}; L)]$ and $\hat{\sigma}_S = \sqrt{\text{Var}[S(\bm{Z}; L)]}$.
By the central limit theorem, under the null hypothesis, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,,
\end{equation*}
where $d$ represents convergence in distribution.
@fig-density_entropyR presents the behavior of the empirical distribution of the standardized test statistic.
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
```{r Simulated_densityR_standardized0, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 20000
mu <- 1
B <- 100

lambda <- 0.9

sample.size <- c(49, 81, 121)
L_values <- c(18)

all_summary_stats <- list()
all_TestStatistics <- list()

# 
for (L in L_values) {

  file_name <- paste0("./Data/resultsRE0_", L, ".Rdata")

  if (file.exists(file_name)) {
    load(file_name)
    message(paste("Loaded existing data for L =", L))
  } else {

    TestStatistics1 <- list()  
    summary_stats <- data.frame(
      LValue = character(),
      SampleSize = numeric(),
      Mean = numeric(),
      SD = numeric()
    )

    for (s in sample.size) {
      TestStat1 <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat1[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - ((lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                                                                        (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mean(z))-log(L))
      }

      mean_val <- mean(TestStat1)
      sd_val <- sd(TestStat1)

      TestStat_standardized <- (TestStat1 -  mean_val) / sd_val

      TestStatistics1[[as.character(s)]] <- data.frame(
        "SampleSize" = rep(s, R), 
        "Test_Statistics" = TestStat_standardized
      )

      summary_stats <- rbind(summary_stats, data.frame(
        LValue = as.character(L),
        SampleSize = s,
        Mean = 0,   # 
        SD = 1      # 
      )) 
    }

    all_TestStatistics[[as.character(L)]] <- TestStatistics1
    all_summary_stats[[as.character(L)]] <- summary_stats

    save(all_TestStatistics, all_summary_stats, file = file_name)
    message(paste("Saved new results for L =", L))
  }
}
```

```{r fig-density_entropyR_standardized0, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="35%", fig.pos="hbt", fig.cap="Standardized empirical densities of $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z}; L)$ under $\\mathcal{H}_0$.", fig.width=4, fig.height=2.5}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

x_limits <- c(-4, 4)  # 
x_breaks <- seq(-4, 4, by = 1)  # Ticks 
y_limits <- c(0, 0.5)  #

selected_L_values <- c( 18)

all_plots <- list()

for (L in selected_L_values) {
  load(paste0("./Data/resultsRE0_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  combined_data$CurveOrder <- factor(combined_data$SampleSize, levels = rev(sample.size))
  combined_data$LegendOrder <- factor(combined_data$SampleSize, levels = sample.size)

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.3, alpha = 0.7) +
    scale_color_viridis(discrete = TRUE, option = "G", direction = -1, begin = 0.1, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(#x = expression("Standardized Test Statistic" ~ epsilon),
        x = expression("Standardized Test Statistic"), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 10, margin = margin(b = 2)),
          axis.text = element_text(size = 10),     
          axis.title = element_text(size = 10),    
          legend.text = element_text(size = 10),   
          legend.title = element_text(size = 10)   
    )

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 1, nrow = 1) +
  plot_layout(guides = "collect")

print(combined_plot)

```
In general, hypothesis tests aim to achieve a controlled type I error rate (size) and high test power (sensitivity to departures from $\mathcal{H}_0$).
To assess these properties, we performed a Monte Carlo simulation with $1000$ replications at significance levels  $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution), varying sample size and values of $L$ for $\mu=1$. 
The observed type I error rates aligned with the nominal values, confirming the test validity.

Further, we examined the power under the alternative hypothesis ($\mathcal{G}^0_I$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and number of looks increase, demonstrating the test effectiveness. The results are shown in Table \ref{tab:table_size_power}.
```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
if (!file.exists("./Data/type_I_results.Rdata")) {

  message("AFile type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
    epsilon <- test_statistic / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value < alpha_nominal)  
  }

  R <- 1000
  mu <- 1
  B <- 100
  lambda <- 0.9
  L_values <- c(5, 8, 18)
  sample_sizes <- c(25, 49, 81, 121)
  alpha_nominals <- c(0.01, 0.05, 0.1)
  results <- data.frame()

  for (L in L_values) {
    for (alpha_nominal in alpha_nominals) {
      TestStatistics <- NULL
      mean_entropy <- numeric(length(sample_sizes))
      sd_entropy <- numeric(length(sample_sizes))

      for (s in sample_sizes) {
        TestStat <- numeric(R)

        for (r in 1:R) {
          z <- gamma_sar_sample(L, mu, s)
          TestStat[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - (
            (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) + 
             (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) +
            log(mean(z)) - log(L)
          )
        }

        mean_entropy[sample_sizes == s] <- mean(TestStat)
        sd_entropy[sample_sizes == s] <- sd(TestStat)

        TestStatistics <- rbind(
          TestStatistics,
          data.frame(
            "L" = rep(L, R),
            "Sample_Size" = rep(s, R),
            "Test_Statistics" = TestStat
          )
        )
      }

      mu_W <- mean_entropy
      sigma_W <- sqrt(sd_entropy^2)

      p_values <- apply(TestStatistics, 1, function(row) {
        calculate_p_value(
          row["Test_Statistics"],
          mu_W[sample_sizes == row["Sample_Size"]],
          sigma_W[sample_sizes == row["Sample_Size"]],
          alpha_nominal
        )
      })

      result <- data.frame(
        "L" = TestStatistics$L,
        "Sample_Size" = TestStatistics$Sample_Size,
        "Alpha_Nominal" = alpha_nominal,
        "P_Value" = p_values
      )
      results <- rbind(results, result)
    }
  }

  save(results, file = "./Data/type_I_results.Rdata")

} else {

  message("File type_I_results.Rdata found. Generating results...")
  load("./Data/type_I_results.Rdata")

}
```

```{r Simulated_power, echo=FALSE, message=FALSE}

if (!file.exists("./Data/results_power.Rdata")) {

  message("File type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
    epsilon <- test_statistic  / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value)
  }

  calculate_type_II_error_rate <- function(p_values, alpha_nominal) {
    type_II_error_rate <- sum(p_values >= alpha_nominal) / length(p_values)
    return(type_II_error_rate)
  }

  calculate_power <- function(R, mu, L_values, B, lambda, sample_sizes, alpha_nominals) {
    results <- data.frame()

    for (L in L_values) {
      for (alpha_nominal in alpha_nominals) {
        TestStatistics <- list()
        mean_entropy <- numeric(length(sample_sizes))
        sd_entropy <- numeric(length(sample_sizes))

        for (s in sample_sizes) {
          TestStat <- numeric(R)

          for (r in 1:R) {
            z <- gi0_sample(mu, -2, L, s)
            TestStat[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - (
              (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) + 
               (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + 
              log(mean(z)) - log(L)
            )
          }

          mean_entropy[sample_sizes == s] <- mean(TestStat)
          sd_entropy[sample_sizes == s] <- sd(TestStat)
          TestStatistics[[as.character(s)]] <- TestStat
        }

        mu_W <- mean_entropy  
        sigma_W <- sqrt(sd_entropy^2)

        p_values <- lapply(TestStatistics, function(TestStat) {
          apply(
            data.frame("Test_Statistics" = TestStat),
            1,
            function(row) {
              calculate_p_value(row["Test_Statistics"], mu_W, sigma_W[as.numeric(names(TestStatistics)) == as.numeric(s)])
            }
          )
        })

        type_II_error_rates <- sapply(p_values, function(p_values_for_size) {
          calculate_type_II_error_rate(p_values_for_size, alpha_nominal)
        })

        power <- 1 - type_II_error_rates

        result_row <- data.frame(
          L = L,
          alpha_nominal = alpha_nominal,
          Sample_Size = sample_sizes,
          power = power
        )

        results <- rbind(results, result_row)
      }
    }

    return(results)
  }

  R <- 1000
  mu <- 1
  L_values <- c(5, 8, 18)
  B <- 100
  lambda <- 0.9
  sample_sizes <- c(25, 49, 81, 121)
  alpha_nominals <- c(0.01, 0.05, 0.1)

  results_power <- calculate_power(R, mu, L_values, B, lambda, sample_sizes, alpha_nominals)

  save(results_power, file = "./Data/results_power.Rdata")

} else {

  message("File type_I_results.Rdata found. Generating results...")
  load("./Data/results_power.Rdata")

}
```

```{r Table_size_and_power, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
suppressMessages({
  suppressWarnings({

    load("./Data/type_I_results.Rdata")
    type_I_error_rates <- tapply(
      results$P_Value, 
      INDEX = list(results$L, results$Sample_Size, results$Alpha_Nominal),
      FUN = function(p_values) sum(p_values) / length(p_values)
    )

    type_I_error_results <- as.data.frame(as.table(type_I_error_rates))
    colnames(type_I_error_results) <- c("L", "Sample_Size", "Alpha_Nominal", "Type_I_Error_Rate")

    spread_results <- type_I_error_results %>%
      pivot_wider(names_from = Alpha_Nominal, values_from = Type_I_Error_Rate)

    load("./Data/results_power.Rdata")

    summary_stats <- results_power %>%
      pivot_wider(names_from = alpha_nominal, values_from = power)

    combined_results <- merge(spread_results, summary_stats, by = c("L", "Sample_Size")) %>%
      arrange(L, Sample_Size)

    selected_columns <- c("L", "Sample_Size", grep("^0\\.", colnames(combined_results), value = TRUE))

    if (length(selected_columns) == 8) {
      combined_results <- combined_results[, selected_columns, drop = FALSE]
      colnames(combined_results) <- c("$\\bm{L}$", "$\\bm{n}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$")
    }

    combined_results[] <- lapply(combined_results, function(x) {
      if (is.numeric(x)) {
        if (all(x %% 1 == 0)) {
          sprintf("$%d$", x)  # Para enteros en formato LaTeX
        } else {
          ifelse(
            x < 0, 
            sprintf("$%.3f$", x),
            sprintf("$\\phantom{-}%.3f$", x)
          )
        }
      } else {
        x
      }
    })

  })
})

table_combined_result <- knitr::kable(
  combined_results,
  caption = "Size and Power of the $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z})$ test statistic.",
  format = "latex",
  booktabs = TRUE,
  align = "ccccccccc",
  escape = FALSE,
  digits = 3,
  label = "table_size_power",
  centering = FALSE,
  table.envir = "table", 
  position="htb", 
  linesep = ""
) %>%
  add_header_above(c(" ", " ", "Size" = 3,  "Power" = 3)) %>%
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 7) %>%
  kable_styling(full_width = T)

print(table_combined_result)
```

# Applications to SAR Data {#sec:app}

In this section, we evaluate the proposed test statistic on real SAR data to assess its effectiveness in detecting heterogeneity.
The test was applied to SAR intensity images from several regions: urban, agricultural, and water. In particular, we analyzed three images: the first from London, United Kingdom; the second from the surroundings of Munich, Germany; and the third from  Dublin Port, Ireland, as shown in Figs.\ \ref{fig:london}(a), \ref{fig:munich}(a), and \ref{fig:dublin}(a).
Table\ \ref{tab:table_param} provides detailed acquisition parameters for these images.
\renewcommand{\arraystretch}{3}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

SAR_data <- data.frame(
  Site = c("London", "Munich", "Dublin" ),
  Mission = c("TanDEM-X", "UAVSAR", "TanDEM-X"),
  Band = c("X", "L", "X"),
  Polarization = c("HH", "HH", "HH"),
  Size = c("$2000\\times2000$", "$1024\\times1024$", "$1100\\times1100$"),
  L = c(1, 12, 16),
  Resol = c("$0.99/0.99$", "$4.9/7.2$", "$1.35/1.35$"),
  Date = c("12-11-2021", "16-04-2015", "03-09-2017")
)

colnames(SAR_data) <- c("\\textbf{Site}", "\\textbf{Mission}", "\\textbf{Band}", "\\textbf{Polarization}", "\\textbf{Size (pixels)}",  "$\\bm{L}$", "\\textbf{Resolution [m]} ", "\\textbf{Acquisition Date}")

SAR_data[] <- lapply(SAR_data, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.1f$", x), sprintf("$\\phantom{-}%.1f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

kable(SAR_data, caption = "Parameters of selected SAR images.",
      format = "latex",
      booktabs = TRUE,
      align = "cccccccc",
      escape = FALSE,
      digits = 2,
      label = "table_param",
      centering = FALSE,
      table.envir = "table", position = "hbt") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 21) %>% 
  kable_styling(full_width = TRUE) %>%
  column_spec(1, width = "3.0cm") %>%
  column_spec(2, width = "3.7cm") %>%
  column_spec(3, width = "1.5cm") %>%
  column_spec(4, width = "2.5cm") %>%
  column_spec(5, width = "4cm") %>%
  column_spec(6, width = "1.5cm") %>%
  column_spec(7, width = "3.0cm") %>%
  column_spec(8, width = "3.5cm")

```

To evaluate the performance of our proposal, we applied the test statistics $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$ and  $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$, based on Shannon entropy as described in\ [@Frery2024], using a local sliding window of size $7\times7$. 
The resulting $p$-value maps are presented in Figs.\ \ref{fig:london}(b), \ref{fig:london}(d), \ref{fig:munich}(b), \ref{fig:munich}(d), \ref{fig:dublin}(b), and \ref{fig:dublin}(d). 
These maps employ a color gradient, where darker regions indicate areas with higher roughness, while lighter regions correspond to smoother, less textured surfaces.

In order to better illustrate the decision-making process of both tests, binary maps at a significance level of $0.05$ are shown in Figs.\ \ref{fig:london}(c), \ref{fig:london}(e), \ref{fig:munich}(c), \ref{fig:munich}(e), \ref{fig:dublin}(c), and \ref{fig:dublin}(e).

In these binary representations, white areas correspond to $p$-values greater than $0.05$, indicating that there is no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions where the $p$-values are less than $0.05$, indicating statistically significant heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations compared to the Shannon-based approach, due to the flexibility provided by the parameter $\lambda$. 
When $\lambda \to 1$, Rényi entropy converges to Shannon entropy, resulting in similar heterogeneity detection. 
Our analysis shows that the optimal $\lambda$ value depends on the number of looks: for single-look images like London, $\lambda = 3$ provided the best contrast, while for multi-look images like Munich and Dublin, $\lambda = 0.9$ was more effective.

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/london_2000.png}
        \caption{ SAR image}
        \label{fig:1a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_renyi.png}
        \caption{ $p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:1b}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:1c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_Shannon_c1.png}
        \caption{$p$-values for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:1d}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:1e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.  }
    \label{fig:london}
\end{figure*}

```

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/munich_1024_2.png}
        \caption{SAR image}
        \label{fig:2a}
    \end{subfigure}
    \begin{subfigure}{0.23\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_muni_renyi.png}
        \caption{$p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:2b}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_munich_renyi.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:2c}
    \end{subfigure}
    \begin{subfigure}{0.23\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_muni_Shan22.png}
        \caption{$p$-values for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:2d}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_munich_sh_AO_L12.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:2e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over Munich, Germany: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:munich}
\end{figure*}

```

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/dublin_1100_hh.png}
        \caption{SAR image}
        \label{fig:3a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/dublin_renyi_09_w7_b100.png}
        \caption{$p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:3b}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_dublin_renyi_09_w7_b100.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:3c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/AO_w7_L16_b100.png}
        \caption{$p$-values for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:3d}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_AO_w7_L16_b100.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:3e}
    \end{subfigure}
   \caption{Detection of heterogeneous areas in a SAR image over Dublin Port, Ireland: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}

    \label{fig:dublin}
\end{figure*}

```

# Conclusion {#sec:conclusion}

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The statistics of the proposed test was computed nonparametrically and its performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of type I error while maintaining a high detection performance that improves as the sample size increases.

The performance of the Rényi-based test was evaluated by comparing its results with those of a test based on Shannon entropy.
This comparison was performed using $p$-values and binary maps. 
The results showed that while both tests successfully detected heterogeneous regions, the Rényi-based test performed better due to its flexibility in adjusting the $\lambda$ parameter, which allowed for finer differentiation of texture variations.

These results suggested that the Rényi entropy was a more accurate alternative to the Shannon entropy for detecting heterogeneity in SAR images, especially due to its adaptability.

Future research will focus on a more quantitative evaluation of both approaches, using statistical measures such as mean square error and other validation metrics to assess their accuracy in detecting heterogeneity.

<!-- # Acknowledgment {-} -->

<!-- This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and the Fundação de Amparo à Ciência e Tecnologia de Pernambuco (FACEPE), Brazil. -->

<!-- # Computational Information {-}   -->
<!-- This article was written using Quarto v6.4, RStudio v2024.12.0.467, and R v4.4.2 and is fully reproducible.  -->

<!-- Code is available at: <https://github.com/rjaneth/renyi-entropy>   -->

[]{.appendix options="Derivation of the Rényi Entropy of $\mathcal{G}^0_I$"}

Let $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$, defined in\ [@Frery2024],  we compute $I = \int_0^\infty [f_{\mathcal{G}^0_I}(z)]^\lambda dz$. Using the pdf  
\begin{equation*}
f_{\mathcal{G}^0_I}(z) = C \frac{z^{L-1}}{(\gamma + Lz)^{L-\alpha}},  
\quad C = \frac{L^L \Gamma(L - \alpha)}{\gamma^\alpha \Gamma(-\alpha) \Gamma(L)}, \label{eq:pdf_G0I}
\end{equation*}
and setting $t = Lz/\gamma$, we obtain  
\begin{equation*}
I = C^\lambda \frac{\gamma^{1-\lambda+\lambda\alpha}}{L^{\lambda L + 1 - \lambda}}  
\int_0^\infty \frac{t^{\lambda(L-1)}}{(1+t)^{\lambda(L-\alpha)}} dt.
\end{equation*}
Applying the Beta function identity, $B(a,b) = \int_0^\infty \frac{t^{a-1}}{(1+t)^{a+b}} dt$ with $a = \lambda(L-1) + 1$, $b = \lambda(-\alpha+1) - 1$, we get  
\begin{equation*}
I= \gamma^{\,1 - \lambda}\,
   L^{\,\lambda - 1}
   \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
   \,B(a,b).
\end{equation*}
Using the Rényi entropy definition in \eqref{E:entropy2}: $H_\lambda(Z) = \frac{1}{1-\lambda} \ln I$,
and substituting $\gamma = -\mu(\alpha + 1)$, we obtain the final expression for $\mathcal{G}^0_I(\mu, \alpha, L)$ in \eqref{eq-HGI0}.

Code is available at: <https://github.com/rjaneth/renyi-entropy>

::: {.content-visible when-format="pdf"}
# References {-}
:::

<!-- [^issues-1023]: ["_[longtable not compatible with 2-column LaTeX documents](https://github.com/jgm/pandoc/issues/1023>)_",  -->

<!-- [^issues-2275]: See the issue here <https://github.com/quarto-dev/quarto-cli/issues/2275> -->

<!-- [IEEEXplore<sup>®</sup>]: <https://ieeexplore.ieee.org/> -->