### 2025-03-15 Text/Reports/report-Renyi/Appendix3.Rmd

---
link-citations: true
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    number_sections: true
    toc_depth: 2
    fig_caption: true
    latex_engine: pdflatex
    template: ./svm-latex-ms.tex
bibliography: ../../Common/references.bib
header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   #- \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{position=bottom,
        labelfont={bf, small, stretch=1.17},
        labelsep=space,
        textfont={small, stretch=1.5},
        aboveskip=1pt, 
        belowskip=1pt,
        singlelinecheck=off,
        justification=centering}   
   #- \usepackage{placeins}
   - \usepackage{amsthm}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{example}{Example}
   - \usepackage{appendix}

biblio-style: apsr
title: "Rényi Entropy-Based Heterogeneity Detection in SAR Data"
thanks: ""
author:
- name: ""
  affiliation:

date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: libertine
fontsize: 11pt
# spacing: double
endnote: no
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE)
# knitr::opts_chunk$set(echo = FALSE,cache=TRUE,
#                       message=FALSE, warning=FALSE,
#                       fig.path='figs/',
#                       cache.path = '_cache/',
#                       fig.process = function(x) {
#                       x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
#                       if (file.rename(x, x2)) x2 else x
#                       })

library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# 
# source("../../../Code/R/MainFunctions/gamma_sar_sample.R")
# source("../../../Code/R/MainFunctions/entropy_gamma_sar.R")
# source("../../../Code/R/MainFunctions/entropy_renyi_gamma_sar.R")
# source("../../../Code/R/MainFunctions/entropy_gI0.R")
# source("../../../Code/R/MainFunctions/gi0_sample.R")
# 
# source("../../../Code/R/MainFunctions/van_es_estimator.R")
# source("../../../Code/R/MainFunctions/correa_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_correa_estimator_log_mean.R")
# source("../../../Code/R/MainFunctions/ebrahimi_estimator.R")
# source("../../../Code/R/MainFunctions/noughabi_arghami_estimator.R")
# source("../../../Code/R/MainFunctions/vasicek_estimator.R")
# source("../../../Code/R/MainFunctions/al_omari_1_estimator.R")
# source("../../../Code/R/MainFunctions/renyi_entropy_estimator_v1.R")
# 
# source("../../../Code/R/MainFunctions/bootstrap_van_es_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_correa_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_ebrahimi_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_noughabi_arghami_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_vasicek_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_al_omari_1_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_renyi_entropy_estimator_v1.R")
# #The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
# source("../../../Code/R/Programs/functions_sample_bias_mse.R")# read_ENVI_images
# source("../../../Code/R/Programs/functions_sample_bias_mse_1.R")
# source("../../../Code/R/Programs/read_ENVI_images.R")

```

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

\appendix
# APPENDIX

# Derivation of the Rényi Entropy for the $\Gamma_{\text{SAR}}(L, \mu)$ Distribution.

The Rényi entropy of order $\lambda$ for a continuous random variable $Z$ with density $f_Z(z)$ is given by
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln \Bigl( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \Bigr),
\quad \lambda > 0, \;\lambda \neq 1.
\label{eq:RenyiDefinition}
\end{align}

Let $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ with pdf
\begin{align*}
f_{\Gamma_{\text{SAR}}}(z; L, \mu)
&= \frac{L^L}{\Gamma(L)\,\mu^L}\,z^{\,L - 1} 
  \exp\Bigl(-\tfrac{L z}{\mu}\Bigr)\mathbbm 1_{\mathbbm R_+}(z).
\end{align*}
Define
\begin{align*}
I 
&= \int_{0}^{\infty}\!\bigl[f_{\Gamma_{\text{SAR}}}(z; L,\mu)\bigr]^\lambda \,dz 
 = \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \int_{0}^{\infty} 
   z^{\,\lambda\,(L-1)} \exp\Bigl(-\tfrac{\lambda\,L}{\mu}\,z\Bigr)\,dz.
\end{align*}
Using the Gamma integral 
$\displaystyle
  \int_{0}^{\infty} x^{p-1} e^{-qx}\,dx 
   = \frac{\Gamma(p)}{q^p},$
with $p = \lambda L - \lambda + 1$ and $q = \tfrac{\lambda L}{\mu}$, it follows that
\begin{align*}
I 
&= \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \frac{\Gamma(\lambda L - \lambda + 1)}
        {\Bigl(\tfrac{\lambda\,L}{\mu}\Bigr)^{\lambda L - \lambda + 1}}.
\end{align*}
Taking the natural logarithm,
\begin{align}
\ln I 
&= \lambda\Bigl(L \ln L - L \ln \mu - \ln \Gamma(L)\Bigr)
   \;+\; \ln\Gamma\!\bigl(\lambda L - \lambda + 1\bigr)
   \;-\; \bigl(\lambda L - \lambda + 1\bigr)\,\Bigl(\ln(\lambda L) - \ln\mu\Bigr).
\label{eq:lnI}
\end{align}
By expanding \eqref{eq:lnI} and collecting terms in $\ln L$ and $\ln \mu$, 
\begin{align}
\ln I 
&= (1 - \lambda)\bigl(\ln \mu - \ln L\bigr)
   \;-\;\lambda\,\ln \Gamma(L)
   \;+\;\ln\Gamma\!\bigl(\lambda(L-1)+1\bigr)
   \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln \lambda.
\label{eq:lnIsimplified}
\end{align}
Substituting \eqref{eq:lnIsimplified} into \eqref{eq:RenyiDefinition} and simplifying,
\begin{multline}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L,\mu)\bigr) 
= \ln \mu - \ln L 
  + \frac{1}{\,1-\lambda\,}
  \Bigl[
    -\lambda\,\ln\Gamma(L)
    + \ln\Gamma\!\bigl(\lambda\,(L-1)+1\bigr)
    - \bigl(\lambda\,(L-1)+1\bigr)\,\ln\lambda
  \Bigr].
\label{eq:RenyiFinal}
\end{multline}
This completes the derivation.

# Derivation of the Rényi Entropy for the \texorpdfstring{$\mathcal{G}^0_I$}{G0I} Distribution

\medskip

\noindent
Let $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$ with pdf
\begin{align*}
f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L) 
&= \frac{L^L\,\Gamma(L-\alpha)}{\gamma^{\alpha}\,\Gamma(-\alpha)\,\Gamma(L)}
   \,\frac{z^{\,L-1}}{\bigl(\gamma + L\,z\bigr)^{\,L-\alpha}}\mathbbm 1_{\mathbbm R_+}(z). \label{E:gamma1}
\end{align*}
In particular, this parameterization is consistent with $\gamma = -\mu(\alpha + 1)$, 
so the final expression can be rewritten in terms of $\mu$.

\medskip

\noindent
Define
\begin{align*}
I 
&= \int_{0}^{\infty} \bigl[f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L)\bigr]^\lambda \,dz
= C^\lambda 
  \int_{0}^{\infty} 
    \frac{z^{\,\lambda(L - 1)}}
         {\bigl(\gamma + L\,z\bigr)^{\,\lambda(L - \alpha)}} 
  \,dz,
\end{align*}
where
\[
C 
= \frac{L^L\,\Gamma(L - \alpha)}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}.
\]
Using the change of variables 
$t = \tfrac{Lz}{\gamma}$, $z = \tfrac{\gamma\,t}{L}$, and $dz = \tfrac{\gamma}{L}\,dt$, 
we obtain
\begin{align*}
I 
&= C^\lambda
   \int_{0}^{\infty}
     \Bigl(\tfrac{\gamma\,t}{L}\Bigr)^{\,\lambda(L - 1)}
     \Bigl(\gamma + L\,\tfrac{\gamma\,t}{L}\Bigr)^{-\lambda(L - \alpha)}
     \,\tfrac{\gamma}{L}\,dt
\\
&= C^\lambda 
   \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \int_{0}^{\infty}
     \frac{t^{\,\lambda(L - 1)}}
          {(1 + t)^{\,\lambda(L - \alpha)}}
   \,dt.
\end{align*}
By the Beta-function identity
\[
\int_{0}^{\infty} \frac{t^{\,a - 1}}{(1 + t)^{\,a + b}} \, dt 
= B(a,b),
\]
where 
\[
a = \lambda(L - 1) + 1,
\quad
b = \lambda(-\alpha + 1) - 1,
\]
it follows that
\begin{align*}
I 
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \,B(a,b).
\end{align*}
Next, we note that 
$\gamma^{\,1 + \lambda(\alpha - 1)} = \gamma^{\,1 - \lambda + \lambda\alpha}$ 
and 
$L^{\,1 + \lambda(L - 1)} = L^{\,\lambda L + 1 - \lambda}.$ 
Since
\[
C^\lambda 
= \biggl(\tfrac{L^L}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}\,\Gamma(L - \alpha)\biggr)^{\!\lambda}
= L^{\lambda L}\,\gamma^{-\alpha \lambda}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda,
\]
we obtain
\begin{align*}
I
&= \gamma^{\,1 - \lambda}\,
   L^{\,\lambda - 1}
   \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
   \,B(a,b).
\end{align*}

\medskip

\noindent
By \eqref{eq:RenyiDefinition}, the Rényi entropy, is given by:
\begin{align*}
H_\lambda(Z)
&= \frac{1}{\,1 - \lambda\,} \,\ln I.
\end{align*}
Hence,
\begin{align*}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,}
  \,\ln\!\Bigl[
    \gamma^{\,1 - \lambda}\,
    L^{\,\lambda - 1}\,
    \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
    \,B(a,b)
  \Bigr].
\end{align*}
Thus, for $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$,
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha, \gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{\,L}\Bigr)
  + \frac{1}{\,1 - \lambda\,}
    \Bigl[
      \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
      + \ln B(a,b)
    \Bigr].
\end{align*}
Using the property 
\[
\ln B(a,b) 
= \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a + b),
\]
where $a + b = \lambda(L - \alpha)$, we have
\begin{align}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{L}\Bigr)
 + \frac{1}{\,1 - \lambda\,}
   \Bigl[
     \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
     + \ln \Gamma(a)
     + \ln \Gamma(b)
     - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
   \Bigr].
\label{eq:GI0RenyiInGamma}
\end{align}
Finally, noting that 
\[
\mu = -\tfrac{\gamma}{\alpha + 1}
\quad\Longrightarrow\quad
\gamma = -\mu(\alpha + 1),
\]
and substituting $\gamma$ into \eqref{eq:GI0RenyiInGamma}, we obtain
\begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\mu, L)\bigr)
= \ln \mu  -  \ln L + \ln(- 1-\alpha)
+ \frac{1}{\,1 - \lambda\,}
  \Bigl[
    \lambda\Bigl(\ln \Gamma(L - \alpha) 
        - \ln \Gamma(-\alpha) 
        - \ln \Gamma(L)\Bigr)\\
    + \ln \Gamma\bigl(\lambda(L - 1) + 1\bigr)
    + \ln \Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
    - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
  \Bigr],
\label{eq:RenyiGI0Final}
\end{multline}
which completes the derivation.

# Relation to the \texorpdfstring{$\Gamma_{\mathrm{SAR}}$}{Gamma SAR} Distribution

The Rényi entropy of the 
$\mathcal{G}^0_I(\alpha,\mu,L)$ distribution can be expressed 
in terms of the Rényi entropy of the 
$\Gamma_{\mathrm{SAR}}(L,\mu)$ distribution, plus additional terms 
involving $\alpha$ and the Gamma function. 
Specifically, we can write:
\begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha,\mu,L)\bigr)
= 
\underbrace{\Bigl[
  \ln\mu \;-\;\ln L 
  \;+\; \frac{1}{\,1-\lambda\,}\Bigl(
    -\lambda \,\ln\Gamma(L) 
    \;+\;\ln\Gamma\bigl(\lambda(L-1)+1\bigr)
    \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
  \Bigr)
\Bigr]}_{\displaystyle H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)}
\\[6pt]
+~\ln\bigl(-1-\alpha\bigr)
+~\frac{1}{\,1-\lambda\,} 
 \Bigl[
   \lambda\bigl(\ln\Gamma(L-\alpha) - \ln\Gamma(-\alpha)\bigr)
   \;+\;\ln\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
   \;-\;\ln\Gamma\bigl(\lambda(L-\alpha)\bigr)
   \;+\;\bigl(\lambda(L-1)+1\bigr)\,\ln(\lambda)
 \Bigr].
\label{eq:GI0_in_terms_of_GammaSAR}
\end{multline}
From \eqref{eq:GI0_in_terms_of_GammaSAR}, the bracketed expression 
on the first line matches 
$H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)$,
while the remaining terms account for the parameter $\alpha$ through additional Gamma functions 
and logarithmic corrections. This decomposition highlights the 
close relationship between the Rényi entropies of the 
$\mathcal{G}^0_I$ and $\Gamma_{\mathrm{SAR}}$ distributions.

# Limit Behavior of \texorpdfstring{$H_\lambda(\mathcal{G}^0_I)$}{Hl(GI0)} as \texorpdfstring{$\alpha \to -\infty$}{alpha->-∞}

\begin{proof}
We want to show that 
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu, \alpha, L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu, L).
\]

We can express \eqref{eq:GI0_in_terms_of_GammaSAR} as follows:
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
&=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L)
\;+\;
\ln\!\bigl(-1-\alpha\bigr)
\\
&\quad
+ \frac{1}{1-\lambda}
\ln \Biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\Biggr].
\end{align*}
Set
\begin{align*}
\Delta(\alpha)
&=
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
-
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\end{align*}
Then 
\begin{align}
\Delta(\alpha)
=
\ln(-1-\alpha)
+
\frac{1}{1-\lambda}
\ln \biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\biggr].
\label{eq:remain}
\end{align}

As \(\alpha \to -\infty\), we have \(-1-\alpha \approx |\alpha|\), so 
\[
\ln(-1-\alpha) 
\sim 
\ln|\alpha|.
\]
Note that for large \(|\alpha|\), we can the asymptotic relation 
\(
\Gamma(x+a)/\Gamma(x+b) 
\sim 
x^{\,a-b}
\).
Specifically:
\[
\Gamma(L-\alpha)/\Gamma(-\alpha) \;\sim\; |\alpha|^L,
\quad
\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)/\Gamma\bigl(\lambda(L-\alpha)\bigr)
\;\sim\; 
\bigl(\lambda|\alpha|\bigr)^{\,(\lambda-1)-\lambda L}.
\]
Thus, inside the logarithm in \eqref{eq:remain},
\[
\frac{
  \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
}{
  \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
}
\;\sim\;
|\alpha|^{\lambda L}
\times
|\alpha|^{(\lambda-1)-\lambda L}
=
|\alpha|^{\,\lambda-1}.
\]
Since \(\lambda^{\lambda(L-1)+1}\)  does not depend on \(\alpha\), multiplying by this constant factor does not alter the asymptotic behavior in \(\alpha\). Therefore,
\[
\frac{1}{1-\lambda}\,
\ln\!\Bigl[\dots\Bigr]
\;\sim\;
\frac{1}{1-\lambda}\;\ln\!\bigl(|\alpha|^{\,\lambda-1}\bigr)
=
\frac{\lambda-1}{1-\lambda}\,\ln|\alpha|
=
-\ln|\alpha|.
\]
Hence
\[
\Delta(\alpha)
\;\sim\;
\ln|\alpha|
-
\ln|\alpha|
=
0
\quad
\text{as}\, \alpha\to -\infty.
\]
This shows 
\(\Delta(\alpha)\to 0\), 
and consequently
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\]
\end{proof}

\section{Bias Reduction via Bootstrap for Entropy Estimation}

The non-parametric Rényi entropy estimator \(\widehat{H}_\lambda(\bm{Z})\), as defined in \eqref{eq:est_R}, exhibits bias, meaning its expectation does not coincide with the true entropy \(H_\lambda\):
\begin{equation}
\operatorname{Bias}(\widehat{H}_\lambda) = {E}[\widehat{H}_\lambda] - H_\lambda \neq 0.
\end{equation}

To reduce this bias, we employ a bootstrap correction approach. Let \(\bm{Z} = (Z_1, Z_2, \dots, Z_n)\) be a sample of size \(n\), and let \(\bm{Z}^{(b)} = (Z_1^{(b)}, Z_2^{(b)}, \dots, Z_n^{(b)})\) for \(b = 1, 2, \dots, B\) be a bootstrap sample obtained by resampling with replacement from \(\bm{Z}\). For each bootstrap sample, we compute the entropy estimator \(\widehat{H}_\lambda(\bm{Z}^{(b)})\). Since bootstrap resampling mimics the sampling variability, we estimate the bias as:
\begin{equation}
\widehat{\operatorname{Bias}} = \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}) - \widehat{H}_\lambda(\bm{Z}).
\end{equation}

A bias-corrected estimator can then be defined as:
\begin{align}
\widetilde{H}_\lambda &= \widehat{H}_\lambda - \widehat{\operatorname{Bias}},\\
&= 2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}).
\end{align}
This estimator is motivated by the well-known bias correction principle:
\begin{equation}
\text{Bias-corrected estimate} = \text{Original estimate} - \text{Bias estimate}.
\end{equation}
The expectation of \(\widetilde{H}_\lambda\) is given by:
\begin{align}
\mathbb{E}[\widetilde{H}_\lambda] &= \mathbb{E}[2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)})]\\
&= 2\mathbb{E}[\widehat{H}_\lambda(\bm{Z})] - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})]\\
&= \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] + \big( H_\lambda - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] \big)\\
&= H_\lambda.
\end{align}
Thus, \(\widetilde{H}_\lambda\) is an **approximately unbiased** estimator of \(H_\lambda\). The effectiveness of this correction depends on the choice of \(B\) and the underlying sample distribution, but it systematically reduces bias in entropy estimation, particularly for small sample sizes.