### 2025-03-19 Text/Articles/RENYI_ENTROPY_SAR_2025/R0-Renyi-entropy-sar.qmd
Returned with comments

---
title: Quantifying Roughness in SAR Imagery with the Rényi Entropy
format:
  ieee-pdf:
    pdf-engine: pdflatex # xelatex
    keep-tex: true  
    #conference: true # comment this line to use journal
    #journaltype: conference # comment this line to use journal
    fig-cap-location: bottom # to crossref figure
    link-citations: true
    colorlinks: true
    linkcolor: black # equations
    citecolor: black # cites
    urlcolor: black
  #ieee-html: default

author:
  #- id: 
  - name: Janeth Alpala, Abraão D.&nbsp;C.&nbsp;Nascimento 
    affiliations:
      - name: Universidade Federal de Pernambuco
        department: Departamento de Estatística
        city: Recife 
        country: PE, Brazil
        postal-code: 50670-901
      #- name: Unknown affiliation
   # orcid: 
    email: janeth.alpala@ufpe.br, abraao@de.ufpe.br
    #url: 
    #membership: "Member, IEEE"
    #attributes:
    #  corresponding: true
    #photo: 
    #bio: |
    #  Use `IEEEbiography`  with figure as  option and
    #  the author name as the argument followed by the biography text.
  - name: Alejandro C.&nbsp;Frery
    affiliations:
      - name: Victoria University of Wellington
        department: School of Mathematics and Statistics
        city: Wellington 
        country: New Zealand
        postal-code: 6140
        orcid: "0000-0002-8002-5341"
        membership: "Fellow, IEEE"
      #- name: Unknown affiliation
    #orcid: 
    email: alejandro.frery@vuw.ac.nz
    # bio: |
    #   Use `IEEEbiographynophoto` and the author name
    #   as the argument followed by the biography text.
    # note: "Template created June 23, 2023; revised `r format(Sys.Date(),format='%B %d, %Y')`."
abstract: |
  Quantifying surface roughness in synthetic aperture radar (SAR) data is critical for accurate geophysical interpretation and remote sensing applications. 
  We propose a test statistic based on a non-parametric estimation of Rényi entropy to characterize surface roughness from SAR intensity data. 
  The statistic is refined using bootstrap to improve its stability, size and power. 
  This approach enhances roughness quantification by capturing scale-dependent variations and addressing data-driven uncertainty.
  Experimental results demonstrate the robustness of the proposed method in distinguishing roughness patterns, offering a statistically rigorous tool for SAR-based terrain analysis.
  <!-- Synthetic aperture radar (SAR) systems have already been successfully used to solve remote sensing problems.  -->
  <!-- A disadvantage of SAR images is the presence of speckle, which is fully developed in homogeneous areas and gamma-distributed in these scenes.  -->
  <!-- In heterogeneous areas, the intensity values are $\mathcal{G}^0_I$-distributed.  -->
  <!-- In this way, the identification of roughness SAR areas (as opposed to homogeneous ones) is an important task.  -->
  <!-- In this work, we propose a family of hypothesis tests driven by an order parameter to identify roughness features in SAR intensity data using Rényi entropy.  -->
  <!-- In particular, we use a non-parametric estimator for the Rényi entropy and investigate some of its properties.  -->
  <!-- As a practical evaluation method, we develop $p$-value maps on which one can observe both (i) the heterogeneous evidence change per texture and (ii) the prediction of homogeneous and heterogeneous categories.  -->
  <!-- The results are in favor of the Rényi-based heterogeneity detector compared to the one based on Shannon entropy. -->
keywords: [Rényi entropy, Gamma distribution, heterogeneity, SAR, hypothesis tests, bootstrap]

#funding: 
 # statement: "The `quarto-ieee` "
pageheader:
  left: Journal XXX, Month Year
  right: #'D. Folio:  A Sample Article Using quarto-ieee'

header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   - \usepackage[T1]{fontenc}
   #- \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{mathtools}
  # - \usepackage[utf8]{inputenc}
   - \usepackage{hyperref}
   #- \hypersetup{draft} #Desactiva enlaces y referencias cruzadas
   - \usepackage{float}
   - \usepackage{array}
   - \usepackage{multirow}
   - \usepackage{wrapfig}
   - \usepackage{colortbl}
   - \usepackage{pdflscape}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{hyperref}
   - \hypersetup{
      colorlinks=true,
      linkcolor=black,
      filecolor=black,
      citecolor=black,
      urlcolor=black}

#bibliography: references.bib
bibliography: ../../Common/references.bib

# execute:
#   echo: false
#   eval: true

---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = FALSE, cache = TRUE)

# Configurar CRAN
options(repos = c(CRAN = "https://cran.rstudio.com/"))

# install and load packages only if they are missing
install_and_load <- function(packages) {
  missing_packages <- packages[!(packages %in% installed.packages()[, "Package"])]
  if (length(missing_packages)) {
    install.packages(missing_packages, dependencies = TRUE)
  }
  invisible(lapply(packages, library, character.only = TRUE))
}
#
#  packages
required_packages <- c(
 "ggplot2", "reshape2", "knitr", "pandoc", "gridExtra", 
  "gtools", "stats4", "rmutil", "scales", "tidyr", "invgamma", 
  "tidyverse", "RColorBrewer", "ggsci", "carData", "ggpubr",  "patchwork", "dplyr", 
  "kableExtra", "ggthemes", "latex2exp", "e1071", "viridis", "nortest", "bookdown"
)

# Install and load only missing packages
install_and_load(required_packages)

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom"))

# External functions
source("./Code/gamma_sar_sample.R")
source("./Code/gi0_sample.R")
source("./Code/al_omari_1_estimator.R")
source("./Code/bootstrap_al_omari_1_estimator.R")
source("./Code/renyi_entropy_estimator_v1.R")
source("./Code/bootstrap_renyi_entropy_estimator_v1.R")
source("./Code/entropy_renyi_gamma_sar.R")
source("./Code/functions_sample_bias_mse.R")
source("./Code/functions_sample_bias_mse_1.R")
#source("./Code/read_ENVI_images.R")

```

# Introduction
[S]{.IEEEPARstart}[ynthetic]{}
 Aperture Radar (SAR) technology has become essential for a wide range of applications&nbsp;[@Yu2023;@Akbarizadeh2012;@Mondini2021]. 
 It provides high-resolution data that is independent of sunlight and operates in a variety of weather conditions that facilitate global Earth monitoring&nbsp;[@Zeng2020].
Despite its advantages, the effective use of SAR data depends on understanding its statistical properties because this data is affected by speckle, a noise-like interference effect&nbsp;[@Argenti2013; @Choi2019]. In intensity format, speckle is typically non-Gaussian, and its presence complicates subsequent image analysis tasks.

The $\mathcal{G}^0_I$ distribution effectively characterizes SAR intensity data because it captures different levels of roughness. 
A limiting case is the Gamma distribution, which arises when speckle is fully developed, indicating textureless regions. 
Although these assumptions offer flexibility in describing SAR intensities, selecting the appropriate distribution can be challenging for two main reasons: the small sample sizes used in practical applications and the inherent difficulties associated with parameter estimation. 
These challenges complicate model selection and highlight the need to explore alternative statistical approaches.

Entropy measures have gained attention as valuable statistical tools for analyzing SAR data, with applications in edge detection&nbsp;[@Nascimento2014], 
segmentation&nbsp;[@Nobre2016], 
classification&nbsp;[@Cassetti2022], 
and noise reduction&nbsp;[@Chan2022].
Traditionally, Shannon entropy&nbsp;[@Shannon1948] has been widely used to quantify data uncertainty and disorder. 
We explore a more general information measure: Rényi entropy—a generalization of Shannon’s formulation. 
This more flexible approach provides additional insights for identifying heterogeneity, making it a promising tool for enhancing SAR image analysis.

We propose a statistical test based on a non-parametric estimator of Rényi entropy to identify heterogeneous regions in SAR data.
The test assesses whether the observed Rényi entropy significantly differs from its expected theoretical value under the assumption of homogeneity.
Compared to our previous approach using Shannon entropy&nbsp;[@Frery2024], the Rényi-based approach improves the detection of heterogeneity.

The remainder of this article is organized as follows.
Section&nbsp;\ref{sec:pre} overviews the statistical models for SAR intensity data and introduces Rényi entropy.
Section&nbsp;\ref{sec:met} describes the proposed hypothesis test and the test statistic.
Section&nbsp;\ref{sec:app} evaluates the performance of the test using SAR data.
Finally, the concluding remarks are drawn in Section&nbsp;\ref{sec:conclusion}.

# PRELIMINARIES {#sec:pre}

## Statistical Models

The main distributions considered for SAR intensity data are the $\Gamma_{\text{SAR}}$ distribution, which is suitable for fully developed speckle, and the $\mathcal{G}^0_I$ distribution, which is able to describe roughness&nbsp;[@Frery1997]. These distributions are characterized by the following probability density functions (pdfs):
\begin{equation}
	f_{\Gamma_{\text{SAR}}}\bigl(z;L, \mu \bigr) 
    = \frac{L^L}{\Gamma(L)\,\mu^L} z^{L-1} 
    \exp \biggl(-\frac{Lz}{\mu}\biggr)
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gamma1}
\end{equation}
and
\begin{multline}
    f_{\mathcal{G}^0_I}\bigl(z; \mu, \alpha, L \bigr) 
    = \frac{L^L\,\Gamma(L-\alpha)}
    {\bigl[-\mu(\alpha+1)\bigr]^{\alpha} \Gamma(-\alpha)\,\Gamma(L)}\\
    \frac{z^{L-1}}
    {\bigl[-\mu(\alpha+1)+Lz\bigr]^{L-\alpha}}
    \mathbbm 1_{\mathbbm R_+}(z), \label{E:gi01}
\end{multline}
where $\mu > 0$ is the mean,
$\alpha < 0$ measures the roughness, $L \geq 1$ is the number of
looks, $\Gamma(\cdot)$ is the gamma function, and
$\mathbbm 1_{A}(z)$ is the indicator function of the set $A$. 
As demonstrated by&nbsp;[@Frery1997], the $\Gamma_{\text{SAR}}$  model is a particular case of the $\mathcal{G}^0_I$ distribution. Specifically, for a given $\mu$ fixed,
$$
f_{\mathcal{G}^0_I}\big(z; \mu, \alpha, L\big)
\longrightarrow 
f_{\Gamma_{\text{SAR}}}(z;L, \mu) 
$$
when $\alpha\to-\infty$.

## Rényi Entropy

Introduced by Alfréd Rényi in 1961&nbsp;[@renyi1961measures], this measure generalizes several well-known entropies, including Shannon's&nbsp;[@Ribeiro2021]. 
For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$, is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions:
\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}
and
\begin{multline}\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr)=\ln\mu-\ln L +\ln(-1 - \alpha)\\
+ \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
      -\ln\Gamma(L)
   \bigr)\\
   +\ln\Gamma\bigl(\lambda(L - 1) + 1\bigr)
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)\\
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\end{multline}
These results are proven in the Appendix.
<!-- ACF Have you checked if it is possible to state the difference between \label{eq-HGI0} and \label{eq-HGammaSAR}? At the end of the day, that is what we want to assess. -->

@fig-plot presents the Rényi entropy of the $\mathcal{G}^0_I$ model as a function of $\mu$ for different values of $\alpha$. 
As $\alpha \to -\infty$, it approaches the Rényi entropy of the $\Gamma_{\text{SAR}}$ law, in line with the fact that the latter is a limiting case of the former.

```{r fig-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="45%",  fig.pos="hbt", fig.cap="$H_{\\lambda}(\\mathcal{G}^0_I)$ converges to the $H_{\\lambda}(\\Gamma_{\\text{SAR}})$ when $\\alpha\\to-\\infty$, with $L=8$ and $\\lambda=0.8$.", fig.width=4.5, fig.height=3.5}

entropy_renyi_gamma_sar <- function(L, mu, lambda) {
  entropy <- (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mu / L)
  return(entropy)
}

entropy_GI0_renyi <- function(alpha, mu, L, lambda) {
  if (lambda <= 0 || lambda == 1) {
    stop("Lambda must be greater than 0 and not equal to 1.")
  }

  gamma <- -mu * (alpha + 1)
  if (any(gamma <= 0)) {
    stop("Gamma must be positive. Check the values of mu and alpha.")
  }

  a <- lambda * (L - 1) + 1
  b <- lambda * (-alpha + 1) - 1
  ab_sum <- lambda * (L - alpha)

  if (any(a <= 0) || any(b <= 0) || any(ab_sum <= 0)) {
    stop("Arguments of the Gamma functions must be positive. Check the values of lambda, L, and alpha.")
  }

  term1 <- log(gamma / L)

  term2 <- lambda * (lgamma(L - alpha) - lgamma(-alpha) - lgamma(L))

  term3 <- lgamma(a)
  term4 <- lgamma(b)
  term5 <- lgamma(ab_sum)

  numerator <- term2 + term3 + term4 - term5

  entropy <- term1 + numerator / (1 - lambda)

  return(entropy)
}

L <- 8
alphas <- c(-3, -8, -20, -1000)
alpha_labels <- c(expression(italic(alpha) == -3), 
                  expression(italic(alpha) == -8), 
                  expression(italic(alpha) == -20), 
                  expression(italic(alpha) == -1000))

mu <- seq(0.1, 10, length.out = 500)
lambda <- 0.8  # Fixed lambda

muEntropy <- data.frame()

for (alpha in alphas) {
  entropies_GI0 <- entropy_GI0_renyi(alpha, mu, L, lambda)
  muEntropy <- rbind(muEntropy, data.frame(mu = mu, Entropy = entropies_GI0, alpha = as.factor(alpha)))
}

muEntropy.molten <- melt(muEntropy, id.vars = c("mu", "alpha"), value.name = "Entropy")

entropies_gamma <- entropy_renyi_gamma_sar(L, mu, lambda)

Entropy_gamma <- data.frame(mu, Entropy_Gamma = entropies_gamma)

Entropy_gamma.molten <- melt(Entropy_gamma, id.vars = "mu", value.name = "Entropy_Gamma")

ggplot() +

  geom_line(data = Entropy_gamma.molten, aes(x = mu, y = Entropy_Gamma), color = "black", 
            linetype = "solid", linewidth = 1.5) +

  geom_line(data = muEntropy.molten, aes(x = mu, y = Entropy, color = alpha), 
            linetype = "longdash", linewidth = 1) +

  annotate("text", x = max(mu) + 0.2, y = max(Entropy_gamma.molten$Entropy_Gamma), 
           label = TeX("${italic(H)}_{\\lambda}(\\Gamma_{\\tiny{SAR}})$"), 
           vjust = 1.6, hjust = 0.8, color = "black",linewidth = 0.2) +

  theme_minimal() +

  scale_color_manual(values = pal_jama()(7)[2:5], labels = alpha_labels) +
 # scale_color_manual(values = brewer.pal(4, "Dark2"), labels = alpha_labels) +
  # Labels and titles
  labs(color = "Roughness", 
       x = expression(mu), 
       y = "Rényi Entropy", 
       linetype = NULL) +

  theme(text = element_text(family = "serif"),
        legend.position = "bottom") +

  coord_cartesian(xlim = c(0, 10), ylim = c(min(muEntropy.molten$Entropy, Entropy_gamma.molten$Entropy_Gamma), 
                                            max(muEntropy.molten$Entropy, Entropy_gamma.molten$Entropy_Gamma) + 0.5))

```

## Non-parametric Estimation of Rényi Entropy

In order to estimate&nbsp;\eqref{E:entropy2} from a random sample, one may estimate the parameters that index the probability density function $f$, and integrate.
This approach heavily relies on the model assumptions and on the absence of outliers.
Alternatively, one may estimate directly $f$.
Such non-parametric estimation of $H(Z)$ has been widely studied using spacing-based estimators, which rely on differences between order statistics [@vasicek1976test; @Ebrahimi1994; @Wieczorkowski1999; @AlOmari2019]. 
Recently, Al-Labadi et al.&nbsp;[@AlLabadi2024] proposed a non-parametric estimator for Rényi entropy following this approach.

Let $\bm{Z}=(Z_1, Z_2,\ldots,Z_n)$ be an independent and identically distributed random sample of size $n$ from a distribution $F$, and let $Z_{(1)} \leq Z_{(2)} \leq \dots \leq Z_{(n)}$ denote its order statistics.
The $m$-spacing density estimator is defined as:
$$
f_n(Z_{(i)}) = \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}},
$$
where $Z_{(i-m)} = Z_{(1)}$ when $i \leq m$, and $Z_{(i+m)} = Z_{(n)}$ if $i \geq n - m$. 
The coefficient $c_i$ is given by:
$$
c_i = 
\begin{cases}
\frac{m + i - 1}{m}, & \text{if } 1 \leq i \leq m, \\%[6pt]
2, & \text{if } m+1 \leq i \leq n - m, \\%[6pt]
\frac{n + m - i}{m}, & \text{if } n - m + 1 \leq i \leq n.
\end{cases}
$$

Following Vasicek&nbsp;[@vasicek1976test] and Ebrahimi et al.&nbsp;[@Ebrahimi1994] for Shannon entropy estimation, and using the $m$-spacing density estimator, the Rényi entropy can be estimated as:
\begin{align}
\label{eq:est_R}
\widehat{H}_\lambda(\bm{Z}) = \frac{1}{1 - \lambda} \ln \left[\frac{1}{n} \sum_{i=1}^{n} \left( \frac{c_i m / n}{Z_{(i+m)} - Z_{(i-m)}} \right)^{\lambda - 1} \right].
\end{align}
This estimator is asymptotically consistent, i.e., it converges in
probability to the true value when $m,n\rightarrow\infty$ and
$m/n\rightarrow0$. 
We used the heuristic spacing $m=\left[\sqrt{n}+0.5\right]$.

# PROPOSED METHODOLOGY {#sec:met}

## On $\lambda$ optimality for a sample size

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$.
To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$. Lower MSE and bias indicate better performance of the estimator in approximating the true entropy.

We found that for $L > 1$ the optimal value is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
However, for $L = 1$, the optimal $\lambda$ tends to be higher and we chose $\lambda = 3$ to achieve good results. 
@fig-plotf illustrates the case for $L = 5$.

```{r fig-plotf, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%",  fig.pos="hbt", fig.cap="Bias and MSE as a function of $\\lambda$, with $n=49$, $L=5$.", fig.width=7, fig.height=4.5}

data <- data.frame(
  Lambda = c(0.9,  0.85, 0.99, 1.1, 1.5),
  Bias = c(0.00158,  -0.00250, 0.02077, 0.03751, 0.06512),
  MSE = c(0.01273,  0.01441, 0.01653, 0.01697, 0.01906)
)

data <- data[order(data$Lambda), ]

bias_plot <- ggplot(data, aes(x = Lambda, y = Bias)) +
  geom_line(color = "#00AFBB", linewidth = 1.0) +  
  geom_point(color = "#00AFBB", size = 3) +  
  geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$Bias), 
         #  label = "Bias", color = "#00AFBB", fontface = "italic", hjust = 1.0) +
  scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) +
  labs(x = expression(lambda), y = "Bias") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

mse_plot <- ggplot(data, aes(x = Lambda, y = MSE)) +
  geom_line(color = "#E69F00", linewidth = 1.0) +  
  geom_point(color = "#E69F00", size = 3) +
  #geom_hline(yintercept = 0, color = "gray50", linetype = "dashed", linewidth = 0.5) +
  #annotate("text", x = max(data$Lambda) - 0.1, y = max(data$MSE), 
           #label = "MSE", color = "#E69F00", fontface = "italic", hjust = 1.0) +
   scale_x_continuous(breaks = data$Lambda, labels = data$Lambda) + 
  labs(x = expression(lambda), y = "MSE") +
  theme_minimal() +
  theme(text = element_text(family = "serif"),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 16))

grid.arrange(bias_plot, mse_plot, nrow = 2)

```

## Bootstrap Correction for Entropy Estimator

Following Refs.&nbsp;[@Frery2024;@Alpala2024], we refined the non-parametric entropy estimator $\widehat{H}_{\lambda}$ in&nbsp;\eqref{eq:est_R} reducing its bias with bootstrap, obtaining $\widetilde{H}_{\lambda}$:
$$
\widetilde{H}_{\lambda} = 2\widehat{H}_{\lambda}(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_{\lambda}(\bm{Z}^{(b)}),
$$
where $B$ is the number of bootstrap replications, and $\bm{Z}^{(b)}$ is the $b$-th resampled dataset obtained by drawing $n$ observations with replacement from $\bm{Z}$.

A Monte Carlo study with $1000$ samples from $\Gamma_{\text{SAR}}$, $\mu = 1$ and $L=5$, using $B = 200$ replications, confirms that the bootstrap-corrected estimator $\widetilde{H}_{\lambda}$, with $\lambda = 0.9$, outperforms the original $\widehat{H}_{\lambda}$ by reducing both bias and MSE for small sample sizes, as shown in @fig-Plot_bias_msef3.
<!-- ACF Remove the minor vertical grid (there is nothing between 49 and 81) -->

```{r Simulated_data_bias_B1, echo=FALSE, message=FALSE, cache = TRUE, autodep = TRUE}
set.seed(1234567890, kind = "Mersenne-Twister")

file_name <- "./Data/results_renyi_B1.Rdata"

if (file.exists(file_name)) {

  load(file_name)
  message("Loaded existing results from results_renyi_B1.Rdata")
} else {
  # Parameters
  sample_sizes <- c(9, 25, 49, 81, 121)
  R <- 500        
  B <- 200        
  mu <- 1        
  L <- 5          
  alpha_values <- c(0.9)

  estimators <- list(
    "Renyi Estimator" = renyi_entropy_estimator_v1,
    "Renyi Estimator Bootstrap" = bootstrap_renyi_entropy_estimator_v1
  )

  results <- calculate_bias_mse_r(sample_sizes, R, B, mu, L, alpha_values, estimators)

  save(results, file = file_name)
  message("Simulations completed and results saved.")
}
```

```{r fig-Plot_bias_msef3, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="42%", fig.pos="hbt",  fig.cap="Bias and MSE of the Rényi entropy estimators for $\\Gamma_{\\text{SAR}}$.", fig.width=7, fig.height=5}

load("./Data/results_renyi_B1.Rdata")

alpha_values <- 0.9
estimators_to_plot <- c("Renyi Estimator", "Renyi Estimator Bootstrap")
latex_estimator_names <- c("Renyi Estimator" = expression("$\\widehat{italic(H)}_{\\lambda}$"),# 
                           "Renyi Estimator Bootstrap" = expression("$\\widetilde{italic(H)}_{\\lambda}$"))
selected_estimators_latex <- latex_estimator_names[estimators_to_plot]

combined_plot_renyi <- generate_plot_renyi(results, alpha_values, selected_estimators_latex, ncol = 1, nrow = 1)

print(combined_plot_renyi)

```
For subsequent simulations we use the $\widetilde{H}_{\lambda}$ estimator.

## Hypothesis Testing

We test whether the observed data come from a homogeneous ($\Gamma_{\text{SAR}}$) or a heterogeneous ($\mathcal{G}^0_I$) region, as follows:
\begin{equation}\label{eq:hypothesis_test}
\begin{cases}
\mathcal{H}_0: \mathbb{E}[\widetilde{H}_{\lambda}] = H_{\lambda}(\Gamma_{\text{SAR}}) & \text{(Homogeneous region)}, \\[6pt]
\mathcal{H}_1: \mathbb{E}[\widetilde{H}_{\lambda}] = H_{\lambda}(\mathcal{G}^0_I) & \text{(Heterogeneous region)}.
\end{cases}
\end{equation}

Under $\mathcal{H}_0$, the expected value of the entropy estimator should match the theoretical $H_{\lambda}(\Gamma_{\text{SAR}})$. If the estimated entropy significantly deviates from its expected value, we reject $\mathcal{H}_0$, indicating the presence of heterogeneity.

Since homogeneity corresponds to $\alpha \to -\infty$ in the $\mathcal{G}^0_I$ model, classical inference methods are inapplicable. Instead, we propose a test statistic based on the Rényi entropy estimator, comparing it with its theoretical expectation under $\mathcal{H}_0$ to distinguish between homogeneous and heterogeneous areas.

## The Proposed Test

Assume we have an estimator $\widetilde{H}_{\lambda}$ for the Rényi entropy of an arbitrary model. For testing \eqref{eq:hypothesis_test}, this estimator is expected to be close to the theoretical expression given in Equation \eqref{eq-HGammaSAR}. 
Since $L\geq1$ is known, we define the test statistic as follows:
\begin{multline}
\label{eq-test}
S(\bm{Z}; L) = \widetilde{H}_{\lambda} - \bigl\{\ln \widehat{\mu} - \ln L + \frac{1}{1-\lambda}
\bigl[-\lambda\,\ln\Gamma(L) \\ 
+ \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  
- \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\bigr]\bigr\},
\end{multline}
where $\widehat{\mu}={n}^{-1}\sum_{i=1}^n Z_{i}$ is the sample mean.

When the null hypothesis holds, this test statistic is expected to be close to zero.
@fig-density_entropyR shows the empirical density of $S(\bm{Z}; L)$, based on $10^4$ replications of the $\Gamma_{\text{SAR}}$ model, with $\lambda=0.9$ and $L \in \{5,18\}$.
<!-- ACF 10^4 replications of samples of size?   -->

```{r Simulated_densityR, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 10000
mu <- 1
B <- 50

lambda <- 0.9

sample.size <- c(49, 81, 121)
L_values <- c(5, 18)

all_summary_stats <- list()
all_TestStatistics <- list()

# 
for (L in L_values) {

  file_name <- paste0("./Data/resultsR_", L, ".Rdata")

  if (file.exists(file_name)) {
    load(file_name)
    message(paste("Loaded existing data for L =", L))
  } else {

    TestStatistics1 <- list()  
    summary_stats <- data.frame(
      LValue = character(),
      SampleSize = numeric(),
      Mean = numeric(),
      SD = numeric()
      #Variance = numeric(),
      #Skewness = numeric(),
     # Kurtosis = numeric(),
      #adpvalue = numeric()
    )

    # Para cada tamaño de muestra
    for (s in sample.size) {
      TestStat1 <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat1[r] <- TestStat <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) -((lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                                                                        (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mean(z))-log(L))
      }

      TestStatistics1[[as.character(s)]] <- data.frame(
        "SampleSize" = rep(s, R), 
        "Test_Statistics" = TestStat1
      )

      mean_val <- mean(TestStat1)
      sd_val <- sd(TestStat1)
      #var_val <- var(TestStat1)
     # skewness_val <- skewness(TestStat1)
     # kurtosis_val <- kurtosis(TestStat1)
     # ad_p_value <- ad.test(TestStatistics1[[as.character(s)]]$Test_Statistics)$p.value

      summary_stats <- rbind(summary_stats, data.frame(
        LValue = as.character(L),
        SampleSize = s,
        Mean = mean_val,
        SD = sd_val
      #  Variance = var_val,
       # Skewness = skewness_val,
       # Kurtosis = kurtosis_val,
       # adpvalue = ad_p_value
      )) 
    }

    all_TestStatistics[[as.character(L)]] <- TestStatistics1
    all_summary_stats[[as.character(L)]] <- summary_stats

    save(all_TestStatistics, all_summary_stats, file = file_name)
    message(paste("Saved new results for L =", L))
  }
}

```

```{r fig-density_entropyR, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="47%", fig.pos="hbt",  fig.cap="Empirical densities of $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z}; L)$ under $\\mathcal{H}_0$.", fig.width=9, fig.height=4.0}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

x_limits <- c(-0.5, 0.5)  # Límites del eje X
x_breaks <- seq(-0.4, 0.4, by = 0.2)  # Ticks del eje X
y_limits <- c(0, 7)       # Límites del eje Y

selected_L_values <- c(5, 18)

all_plots <- list()

for (L in selected_L_values) {
  load(paste0("./Data/resultsR_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  combined_data$CurveOrder <- factor(combined_data$SampleSize, levels = rev(sample.size))

  combined_data$LegendOrder <- factor(combined_data$SampleSize, levels = sample.size)

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.5) +
    scale_color_viridis(discrete = TRUE, option = "C", direction = -1, begin = 0.2, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(
        x = expression("Test Statistic" ~ S[widetilde(italic(H))[lambda]](italic(bold(Z))* ";" **phantom(" ")* italic(L))), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 16, margin = margin(b = 2)),
          axis.text = element_text(size = 16),     
        axis.title = element_text(size = 16),    
        legend.text = element_text(size = 16),   
        legend.title = element_text(size = 16)   
          )

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 2, nrow = 1) +
  plot_layout(guides = "collect")

print(combined_plot)

```

Vasicek&nbsp;[@vasicek1976test] proved that, for sufficiently large samples, $S(\bm{Z}; L)$ asymptotically follows a normal distribution:
\begin{equation*}
S(\bm{Z}; L) \overset{d}{\longrightarrow} \mathcal{N}(\mu_S,\,\sigma^{2}_S)\,,
\end{equation*}
where $d$ represents convergence in distribution. Here, $\mu_S  = \mathbb{E}[S(\bm{Z}; L)]$ and $\sigma^{2}_S = \text{Var}[S(\bm{Z}; L)]$ are the theoretical mean and variance of $S(\bm{Z}; L)$.

We define the standardized test statistic:
\begin{equation*}
\varepsilon = \frac{S(\bm{Z}; L) - \hat{\mu}_S}{\hat{\sigma}_S},
\end{equation*}
where $\hat{\mu}_S$ and $\hat{\sigma}_S$ are the empirical mean and standard deviation of $S(\bm{Z}; L)$, obtained by Monte Carlo simulations under the null hypothesis. 
By the central limit theorem, $\varepsilon$ asymptotically follows a standard normal distribution:
\begin{equation*}
\varepsilon \overset{d}{\longrightarrow} \mathcal{N}(0,1)\,.
\end{equation*}
Thus, the $p$-values are calculated as $2\Phi(-|\varepsilon|)$,
where $\Phi(\cdot)$ is the cumulative distribution function of the standard normal distribution.
@fig-density_entropyR_standardized0 shows smoothed histograms of standardized test statistics.

```{r Simulated_densityR_standardized0, echo=FALSE, message=FALSE}
set.seed(1234567890, kind = "Mersenne-Twister")

R <- 20000
mu <- 1
B <- 100

lambda <- 0.9

sample.size <- c(49, 81, 121)
L_values <- c(18)

all_summary_stats <- list()
all_TestStatistics <- list()

# 
for (L in L_values) {

  file_name <- paste0("./Data/resultsRE0_", L, ".Rdata")

  if (file.exists(file_name)) {
    load(file_name)
    message(paste("Loaded existing data for L =", L))
  } else {

    TestStatistics1 <- list()  
    summary_stats <- data.frame(
      LValue = character(),
      SampleSize = numeric(),
      Mean = numeric(),
      SD = numeric()
    )

    for (s in sample.size) {
      TestStat1 <- numeric(R)

      for (r in 1:R) {
        z <- gamma_sar_sample(L, mu, s)
        TestStat1[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - ((lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) +
                                                                        (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + log(mean(z))-log(L))
      }

      mean_val <- mean(TestStat1)
      sd_val <- sd(TestStat1)

      TestStat_standardized <- (TestStat1 -  mean_val) / sd_val

      TestStatistics1[[as.character(s)]] <- data.frame(
        "SampleSize" = rep(s, R), 
        "Test_Statistics" = TestStat_standardized
      )

      summary_stats <- rbind(summary_stats, data.frame(
        LValue = as.character(L),
        SampleSize = s,
        Mean = 0,   # 
        SD = 1      # 
      )) 
    }

    all_TestStatistics[[as.character(L)]] <- TestStatistics1
    all_summary_stats[[as.character(L)]] <- summary_stats

    save(all_TestStatistics, all_summary_stats, file = file_name)
    message(paste("Saved new results for L =", L))
  }
}
```

```{r fig-density_entropyR_standardized0, echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth = TRUE, out.width="35%", fig.pos="hbt", fig.cap="Standardized empirical densities of $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z}; L)$ under $\\mathcal{H}_0$.", fig.width=4, fig.height=2.5}

theme_set(theme_minimal() +
            theme(text = element_text(family = "serif"),
                  legend.position = "bottom",
                  legend.text = element_text(angle = 0, vjust = 0.5)))

x_limits <- c(-4, 4)  # 
x_breaks <- seq(-4, 4, by = 1)  # Ticks 
y_limits <- c(0, 0.5)  #

selected_L_values <- c( 18)

all_plots <- list()

for (L in selected_L_values) {
  load(paste0("./Data/resultsRE0_", L, ".Rdata"))

  combined_data <- do.call(rbind, all_TestStatistics[[as.character(L)]])
  combined_data$L <- L

  combined_data$CurveOrder <- factor(combined_data$SampleSize, levels = rev(sample.size))
  combined_data$LegendOrder <- factor(combined_data$SampleSize, levels = sample.size)

  p <- ggplot(combined_data, aes(x = Test_Statistics, col = LegendOrder, linetype = LegendOrder, group = CurveOrder)) +
    geom_line(stat = "density", linewidth = 1.3, alpha = 0.7) +
    scale_color_viridis(discrete = TRUE, option = "G", direction = -1, begin = 0.1, end = 0.7, name = "Sample Size") +
    scale_linetype_manual(values = rep("solid", length(sample.size)), name = "Sample Size") +
    scale_x_continuous(limits = x_limits, breaks = x_breaks) + 
    scale_y_continuous(limits = y_limits) +  
    labs(#x = expression("Standardized Test Statistic" ~ epsilon),
        x = expression("Standardized Test Statistic"), 
        y = "Density"
    ) +
    ggtitle(bquote(italic(L) == .(L))) +
    theme(plot.title = element_text(hjust = 0.5, size = 10, margin = margin(b = 2)),
          axis.text = element_text(size = 10),     
          axis.title = element_text(size = 10),    
          legend.text = element_text(size = 10),   
          legend.title = element_text(size = 10)   
    )

  all_plots[[as.character(L)]] <- p
}

combined_plot <- wrap_plots(all_plots, ncol = 1, nrow = 1) +
  plot_layout(guides = "collect")

print(combined_plot)

```
In general, hypotheses tests aim to control the Type&nbsp;I error rate (size) with high test power (sensitivity to departures from $\mathcal{H}_0$, low Type&nbsp;II error rate).
To assess these properties, we performed a Monte Carlo simulation with $1000$ replications at significance levels  $1\%$, $5\%$, and $10\%$, evaluating the test under the null hypothesis ($\Gamma_{\text{SAR}}$ distribution), varying sample size and values of $L$ for $\mu=1$. 
<!-- ACF I wish you had used the siunitx package for typing percentages, but not if this is incompatible with the tables -->
The observed Type&nbsp;I error rates align with the nominal values, confirming the test validity.

Further, we examined the power under the alternative hypothesis ($\mathcal{G}^0_I$ distribution) with $\mu=1$, $\lambda=0.9$ and $\alpha=-2$.
As expected, power improves as the sample size and number of looks increase, demonstrating the test effectiveness. The results are shown in Table \ref{tab:table_size_power}.
```{r Simulated_error_type_I, echo=FALSE, message=FALSE}
if (!file.exists("./Data/type_I_results.Rdata")) {

  message("AFile type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W, alpha_nominal) {
    epsilon <- test_statistic / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value < alpha_nominal)  
  }

  R <- 1000
  mu <- 1
  B <- 100
  lambda <- 0.9
  L_values <- c(5, 8, 18)
  sample_sizes <- c(25, 49, 81, 121)
  alpha_nominals <- c(0.01, 0.05, 0.1)
  results <- data.frame()

  for (L in L_values) {
    for (alpha_nominal in alpha_nominals) {
      TestStatistics <- NULL
      mean_entropy <- numeric(length(sample_sizes))
      sd_entropy <- numeric(length(sample_sizes))

      for (s in sample_sizes) {
        TestStat <- numeric(R)

        for (r in 1:R) {
          z <- gamma_sar_sample(L, mu, s)
          TestStat[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - (
            (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) + 
             (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) +
            log(mean(z)) - log(L)
          )
        }

        mean_entropy[sample_sizes == s] <- mean(TestStat)
        sd_entropy[sample_sizes == s] <- sd(TestStat)

        TestStatistics <- rbind(
          TestStatistics,
          data.frame(
            "L" = rep(L, R),
            "Sample_Size" = rep(s, R),
            "Test_Statistics" = TestStat
          )
        )
      }

      mu_W <- mean_entropy
      sigma_W <- sqrt(sd_entropy^2)

      p_values <- apply(TestStatistics, 1, function(row) {
        calculate_p_value(
          row["Test_Statistics"],
          mu_W[sample_sizes == row["Sample_Size"]],
          sigma_W[sample_sizes == row["Sample_Size"]],
          alpha_nominal
        )
      })

      result <- data.frame(
        "L" = TestStatistics$L,
        "Sample_Size" = TestStatistics$Sample_Size,
        "Alpha_Nominal" = alpha_nominal,
        "P_Value" = p_values
      )
      results <- rbind(results, result)
    }
  }

  save(results, file = "./Data/type_I_results.Rdata")

} else {

  message("File type_I_results.Rdata found. Generating results...")
  load("./Data/type_I_results.Rdata")

}
```

```{r Simulated_power, echo=FALSE, message=FALSE}

if (!file.exists("./Data/results_power.Rdata")) {

  message("File type_I_results.Rdata not found. Generating results...")

  set.seed(1234567890, kind = "Mersenne-Twister")

  calculate_p_value <- function(test_statistic, mu_W, sigma_W) {
    epsilon <- test_statistic  / sigma_W
    p_value <- 2 * (1 - pnorm(abs(epsilon)))
    return(p_value)
  }

  calculate_type_II_error_rate <- function(p_values, alpha_nominal) {
    type_II_error_rate <- sum(p_values >= alpha_nominal) / length(p_values)
    return(type_II_error_rate)
  }

  calculate_power <- function(R, mu, L_values, B, lambda, sample_sizes, alpha_nominals) {
    results <- data.frame()

    for (L in L_values) {
      for (alpha_nominal in alpha_nominals) {
        TestStatistics <- list()
        mean_entropy <- numeric(length(sample_sizes))
        sd_entropy <- numeric(length(sample_sizes))

        for (s in sample_sizes) {
          TestStat <- numeric(R)

          for (r in 1:R) {
            z <- gi0_sample(mu, -2, L, s)
            TestStat[r] <- bootstrap_renyi_entropy_estimator_v1(z, B, lambda) - (
              (lambda * lgamma(L) - lgamma(lambda * (L - 1) + 1) + 
               (lambda * (L - 1) + 1) * log(lambda)) / (lambda - 1) + 
              log(mean(z)) - log(L)
            )
          }

          mean_entropy[sample_sizes == s] <- mean(TestStat)
          sd_entropy[sample_sizes == s] <- sd(TestStat)
          TestStatistics[[as.character(s)]] <- TestStat
        }

        mu_W <- mean_entropy  
        sigma_W <- sqrt(sd_entropy^2)

        p_values <- lapply(TestStatistics, function(TestStat) {
          apply(
            data.frame("Test_Statistics" = TestStat),
            1,
            function(row) {
              calculate_p_value(row["Test_Statistics"], mu_W, sigma_W[as.numeric(names(TestStatistics)) == as.numeric(s)])
            }
          )
        })

        type_II_error_rates <- sapply(p_values, function(p_values_for_size) {
          calculate_type_II_error_rate(p_values_for_size, alpha_nominal)
        })

        power <- 1 - type_II_error_rates

        result_row <- data.frame(
          L = L,
          alpha_nominal = alpha_nominal,
          Sample_Size = sample_sizes,
          power = power
        )

        results <- rbind(results, result_row)
      }
    }

    return(results)
  }

  R <- 1000
  mu <- 1
  L_values <- c(5, 8, 18)
  B <- 100
  lambda <- 0.9
  sample_sizes <- c(25, 49, 81, 121)
  alpha_nominals <- c(0.01, 0.05, 0.1)

  results_power <- calculate_power(R, mu, L_values, B, lambda, sample_sizes, alpha_nominals)

  save(results_power, file = "./Data/results_power.Rdata")

} else {

  message("File type_I_results.Rdata found. Generating results...")
  load("./Data/results_power.Rdata")

}
```

```{r Table_size_and_power, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
suppressMessages({
  suppressWarnings({

    load("./Data/type_I_results.Rdata")
    type_I_error_rates <- tapply(
      results$P_Value, 
      INDEX = list(results$L, results$Sample_Size, results$Alpha_Nominal),
      FUN = function(p_values) sum(p_values) / length(p_values)
    )

    type_I_error_results <- as.data.frame(as.table(type_I_error_rates))
    colnames(type_I_error_results) <- c("L", "Sample_Size", "Alpha_Nominal", "Type_I_Error_Rate")

    spread_results <- type_I_error_results %>%
      pivot_wider(names_from = Alpha_Nominal, values_from = Type_I_Error_Rate)

    load("./Data/results_power.Rdata")

    summary_stats <- results_power %>%
      pivot_wider(names_from = alpha_nominal, values_from = power)

    combined_results <- merge(spread_results, summary_stats, by = c("L", "Sample_Size")) %>%
      arrange(L, Sample_Size)

    selected_columns <- c("L", "Sample_Size", grep("^0\\.", colnames(combined_results), value = TRUE))

    if (length(selected_columns) == 8) {
      combined_results <- combined_results[, selected_columns, drop = FALSE]
      colnames(combined_results) <- c("$\\bm{L}$", "$\\bm{n}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$", 
                                      "$\\bm{1\\%}$", "$\\bm{5\\%}$", "$\\bm{10\\%}$")
    }

    combined_results[] <- lapply(combined_results, function(x) {
      if (is.numeric(x)) {
        if (all(x %% 1 == 0)) {
          sprintf("$%d$", x)  # Para enteros en formato LaTeX
        } else {
          ifelse(
            x < 0, 
            sprintf("$%.3f$", x),
            sprintf("$\\phantom{-}%.3f$", x)
          )
        }
      } else {
        x
      }
    })

  })
})

table_combined_result <- knitr::kable(
  combined_results,
  caption = "Size and Power of the $S_{\\widetilde{H}_{\\lambda}}(\\bm{Z})$ test statistic.",
  format = "latex",
  booktabs = TRUE,
  align = "ccccccccc",
  escape = FALSE,
  digits = 3,
  label = "table_size_power",
  centering = FALSE,
  table.envir = "table", 
  position="htb", 
  linesep = ""
) %>%
  add_header_above(c(" ", " ", "Size" = 3,  "Power" = 3)) %>%
  collapse_rows(columns = 1:2, latex_hline = "major", valign = "middle") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 7) %>%
  kable_styling(full_width = T)

print(table_combined_result)
```

# Applications to SAR Data {#sec:app}

In this section, we evaluate the proposed test statistic on SAR data to assess its effectiveness in detecting heterogeneity.
We applied the test to images covering urban, agricultural, and water regions. 
We analyzed images from London, United Kingdom; from the surroundings of Munich, Germany; and from  Dublin Port, Ireland, as shown in Figs.&nbsp;\ref{fig:london}(a), \ref{fig:munich}(a), and \ref{fig:dublin}(a).
Table&nbsp;\ref{tab:table_param} provides detailed acquisition parameters for these images.

<!-- ACF They all have HH polarization; the column is not informative, and can be deleted to gain space -->
\renewcommand{\arraystretch}{3}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

SAR_data <- data.frame(
  Site = c("London", "Munich", "Dublin" ),
  Mission = c("TanDEM-X", "UAVSAR", "TanDEM-X"),
  Band = c("X", "L", "X"),
  Polarization = c("HH", "HH", "HH"),
  Size = c("$2000\\times2000$", "$1024\\times1024$", "$1100\\times1100$"),
  L = c(1, 12, 16),
  Resol = c("$0.99/0.99$", "$4.9/7.2$", "$1.35/1.35$"),
  Date = c("12-11-2021", "16-04-2015", "03-09-2017")
)

colnames(SAR_data) <- c("\\textbf{Site}", "\\textbf{Mission}", "\\textbf{Band}", "\\textbf{Polarization}", "\\textbf{Size (pixels)}",  "$\\bm{L}$", "\\textbf{Resolution [m]} ", "\\textbf{Acquisition Date}")

SAR_data[] <- lapply(SAR_data, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.1f$", x), sprintf("$\\phantom{-}%.1f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

kable(SAR_data, caption = "Parameters of selected SAR images.",
      format = "latex",
      booktabs = TRUE,
      align = "cccccccc",
      escape = FALSE,
      digits = 2,
      label = "table_param",
      centering = FALSE,
      table.envir = "table", position = "hbt") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 21) %>% 
  kable_styling(full_width = TRUE) %>%
  column_spec(1, width = "3.0cm") %>%
  column_spec(2, width = "3.7cm") %>%
  column_spec(3, width = "1.5cm") %>%
  column_spec(4, width = "2.5cm") %>%
  column_spec(5, width = "4cm") %>%
  column_spec(6, width = "1.5cm") %>%
  column_spec(7, width = "3.0cm") %>%
  column_spec(8, width = "3.5cm")

```

We applied the test statistics $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$ and  $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$, based on Shannon entropy as described in&nbsp;[@Frery2024], using a local sliding window of size $7\times7$. 
<!-- ACF What is $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$? -->
The resulting $p$-value maps are presented in Figs.&nbsp;\ref{fig:london}(b), \ref{fig:london}(d), \ref{fig:munich}(b), \ref{fig:munich}(d), \ref{fig:dublin}(b), and \ref{fig:dublin}(d). 
These maps employ a color gradient, where darker regions indicate areas with higher roughness, while lighter regions correspond to smoother, less textured surfaces.

Figs.&nbsp;\ref{fig:london}(c), \ref{fig:london}(e), \ref{fig:munich}(c), \ref{fig:munich}(e), \ref{fig:dublin}(c), and \ref{fig:dublin}(e) 
are binary maps at significance level of $0.05$ that 
illustrate the decision-making process.
In these binary representations, white areas correspond to $p$-values greater than $0.05$, indicating that there is no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions where the $p$-values are less than $0.05$, indicating statistically significant heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations compared to the Shannon-based approach, due to the flexibility provided by the parameter $\lambda$. 
<!-- ACF Where is the Shannon-based approach shown? -->
<!-- When $\lambda \to 1$, Rényi entropy converges to Shannon entropy, resulting in similar heterogeneity detection.  -->

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/london_2000.png}
        \caption{ SAR image}
        \label{fig:1a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_renyi.png}
        \caption{ $p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:1b}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_renyi_L1_.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:1c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_london_Shannon_c1.png}
        \caption{$p$-values for $\tiny{S_{\widetilde{H}_{\text{AO}}}}$ }
        \label{fig:1d}
    \end{subfigure}
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_london_shannon.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:1e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.  }
    \label{fig:london}
\end{figure*}

```

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.17\textwidth}
        \includegraphics[width=\linewidth]{./Figures/munich_1024_2.png}
        \caption{SAR image}
        \label{fig:2a}
    \end{subfigure}
    \begin{subfigure}{0.23\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_muni_renyi.png}
        \caption{$p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:2b}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_munich_renyi.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:2c}
    \end{subfigure}
    \begin{subfigure}{0.23\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_pvalue_muni_Shan22.png}
        \caption{$p$-values for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:2d}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_munich_sh_AO_L12.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:2e}
    \end{subfigure}
    \caption{Detection of heterogeneous areas in a SAR image over Munich, Germany: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}
    \label{fig:munich}
\end{figure*}

```

```{=latex}
\begin{figure*}[hbt]
    \centering
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/dublin_1100_hh.png}
        \caption{SAR image}
        \label{fig:3a}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/dublin_renyi_09_w7_b100.png}
        \caption{$p$-values for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:3b}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_dublin_renyi_09_w7_b100.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\lambda}}}$}
        \label{fig:3c}
    \end{subfigure}
    \begin{subfigure}{0.22\textwidth}
        \includegraphics[width=\linewidth]{./Figures/AO_w7_L16_b100.png}
        \caption{$p$-values for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:3d}
    \end{subfigure}
    \begin{subfigure}{0.16\textwidth}
        \includegraphics[width=\linewidth]{./Figures/H_005_AO_w7_L16_b100.png}
        \caption{Binary map for $\small{S_{\widetilde{H}_{\text{AO}}}}$}
        \label{fig:3e}
    \end{subfigure}
   \caption{Detection of heterogeneous areas in a SAR image over Dublin Port, Ireland: comparison with the tests $\small{S_{\widetilde{H}_{\lambda}}}$ and $\small{S_{\widetilde{H}_{\text{AO}}}}$.}

    \label{fig:dublin}
\end{figure*}

```

# Conclusions {#sec:conclusion}

This study presented a new statistical approach based on Rényi entropy for detecting heterogeneity in SAR images, distinguishing between fully developed speckle and heterogeneous clutter. 
The statistics of the proposed test was computed non-parametrically and its performance was evaluated by a Monte Carlo study. 
The results showed that the method effectively controls the probability of Type&nbsp;I error while maintaining a high detection performance that improves as the sample size increases.

The performance of the Rényi-based test was evaluated by comparing its results with those of a test based on Shannon entropy.
<!-- ACF Again, where are the Shannon results? -->
This comparison was performed using $p$-values and binary maps. 
The results showed that while both tests successfully detected heterogeneous regions, the Rényi-based test performed better due to its flexibility in adjusting the $\lambda$ parameter, which allowed for finer differentiation of texture variations.

These results suggested that the Rényi entropy was a more accurate alternative to the Shannon entropy for detecting heterogeneity in SAR images, especially due to its adaptability.

Future research will focus on a more quantitative evaluation of both approaches, using statistical measures such as mean square error and other validation metrics to assess their accuracy in detecting heterogeneity.

<!-- # Acknowledgment {-} -->

<!-- This study was financed in part by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior (CAPES) and the Fundação de Amparo à Ciência e Tecnologia de Pernambuco (FACEPE), Brazil. -->

<!-- # Computational Information {-}   -->
<!-- This article was written using Quarto v6.4, RStudio v2024.12.0.467, and R v4.4.2 and is fully reproducible.  -->

<!-- Code is available at: <https://github.com/rjaneth/renyi-entropy>   -->

[]{.appendix options="Derivation of the Rényi Entropy of $\mathcal{G}^0_I$"}

Let $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$ as defined in Ref.&nbsp;[@Frery2024].
We compute $I = \int_0^\infty [f_{\mathcal{G}^0_I}(z)]^\lambda dz$ using the pdf  
\begin{equation*}
f_{\mathcal{G}^0_I}(z) = C \frac{z^{L-1}}{(\gamma + Lz)^{L-\alpha}},  
\quad C = \frac{L^L \Gamma(L - \alpha)}{\gamma^\alpha \Gamma(-\alpha) \Gamma(L)}. \label{eq:pdf_G0I}
\end{equation*}
Setting $t = Lz/\gamma$, we obtain  
\begin{equation*}
I = C^\lambda \frac{\gamma^{1-\lambda+\lambda\alpha}}{L^{\lambda L + 1 - \lambda}}  
\int_0^\infty \frac{t^{\lambda(L-1)}}{(1+t)^{\lambda(L-\alpha)}} dt.
\end{equation*}
Applying the Beta function identity, $B(a,b) = \int_0^\infty \frac{t^{a-1}}{(1+t)^{a+b}} dt$ with $a = \lambda(L-1) + 1$, $b = \lambda(-\alpha+1) - 1$, we get  
\begin{equation*}
I= \gamma^{\,1 - \lambda}\,
   L^{\,\lambda - 1}
   \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
   \,B(a,b).
\end{equation*}
Using the Rényi entropy definition in \eqref{E:entropy2}: $H_\lambda(Z) = (1-\lambda)^{-1} \ln I$,
and substituting $\gamma = -\mu(\alpha + 1)$, we obtain the final expression for $\mathcal{G}^0_I(\mu, \alpha, L)$ in \eqref{eq-HGI0}.

The code is available at: <https://github.com/rjaneth/renyi-entropy>

::: {.content-visible when-format="pdf"}
# References {-}
:::

<!-- [^issues-1023]: ["_[longtable not compatible with 2-column LaTeX documents](https://github.com/jgm/pandoc/issues/1023>)_",  -->

<!-- [^issues-2275]: See the issue here <https://github.com/quarto-dev/quarto-cli/issues/2275> -->

<!-- [IEEEXplore<sup>®</sup>]: <https://ieeexplore.ieee.org/> -->