### 2025-03-29 Text/Reports/report-Renyi/Appendix.Rmd

---
link-citations: true
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    number_sections: true
    toc_depth: 2
    fig_caption: true
    latex_engine: pdflatex
    template: ./svm-latex-ms.tex
bibliography: ../../Common/references.bib
header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   #- \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{position=bottom,
        labelfont={bf, small, stretch=1.17},
        labelsep=space,
        textfont={small, stretch=1.5},
        aboveskip=1pt, 
        belowskip=1pt,
        singlelinecheck=off,
        justification=centering}   
   #- \usepackage{placeins}
   - \usepackage{amsthm}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{example}{Example}
   - \usepackage{appendix}

biblio-style: apsr
title: "Rényi Entropy for Heterogeneity Identification in SAR Data"
thanks: ""
author:
- name: J. Alpala
  affiliation: 
abstract: " We present a statistical method for identifying roughness characteristics in synthetic aperture radar (SAR) intensity data through the use of Rényi entropy. 
Homogeneous regions, characterized by fully-developed speckle, are modeled with the Gamma distribution, while heterogeneous areas are described by the ${G}_I^0$ distribution, where fully-developed speckle assumptions do not apply. Our proposed hypothesis test, based on a non-parametric estimator of Renyi entropy, effectively distinguishes between homogeneous and heterogeneous regions. 
Results indicate that the Rényi entropy-based approach provides superior performance compared to previous methods using Shannon entropy, offering enhanced detection of heterogeneity in both simulated and real SAR data."
keywords: "Gamma distribution, heterogeneity, SAR, Rényi entropy, hypothesis tests "
date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: libertine
fontsize: 11pt
# spacing: double
endnote: no
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE)
# knitr::opts_chunk$set(echo = FALSE,cache=TRUE,
#                       message=FALSE, warning=FALSE,
#                       fig.path='figs/',
#                       cache.path = '_cache/',
#                       fig.process = function(x) {
#                       x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
#                       if (file.rename(x, x2)) x2 else x
#                       })

library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

source("../../../Code/R/MainFunctions/gamma_sar_sample.R")
source("../../../Code/R/MainFunctions/entropy_gamma_sar.R")
source("../../../Code/R/MainFunctions/entropy_renyi_gamma_sar.R")
source("../../../Code/R/MainFunctions/entropy_gI0.R")
source("../../../Code/R/MainFunctions/gi0_sample.R")

source("../../../Code/R/MainFunctions/van_es_estimator.R")
source("../../../Code/R/MainFunctions/correa_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_correa_estimator_log_mean.R")
source("../../../Code/R/MainFunctions/ebrahimi_estimator.R")
source("../../../Code/R/MainFunctions/noughabi_arghami_estimator.R")
source("../../../Code/R/MainFunctions/vasicek_estimator.R")
source("../../../Code/R/MainFunctions/al_omari_1_estimator.R")
source("../../../Code/R/MainFunctions/renyi_entropy_estimator_v1.R")

source("../../../Code/R/MainFunctions/bootstrap_van_es_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_correa_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_ebrahimi_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_noughabi_arghami_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_vasicek_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_al_omari_1_estimator.R")
source("../../../Code/R/MainFunctions/bootstrap_renyi_entropy_estimator_v1.R")
#The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
source("../../../Code/R/Programs/functions_sample_bias_mse.R")# read_ENVI_images
source("../../../Code/R/Programs/functions_sample_bias_mse_1.R")
source("../../../Code/R/Programs/read_ENVI_images.R")

```

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

\appendix

\section{ Derivation of the Rényi Entropy for the \( G_I^0 \) Distribution}\label{app1}

The Rényi entropy of order \( \lambda \) for a continuous random variable \( Z \) with density \( f_Z(z) \) is given by:

\[
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \left( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \right), \quad \lambda > 0, \, \lambda \neq 1.
\]

For \( Z \sim G_I^0(\alpha, \gamma, L) \), the density function is:

\[
f_Z(z) = C \cdot \frac{z^{L - 1}}{(\gamma + L z)^{L - \alpha}}, \quad \text{where} \quad C = \frac{L^L \Gamma(L - \alpha)}{\gamma^\alpha \Gamma(-\alpha) \Gamma(L)}.
\]

Raising \( f_Z(z) \) to the power \( \lambda \) and substituting into the entropy integral, we have:

\[
I = \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz = C^\lambda \int_{0}^{\infty} \frac{z^{\lambda(L - 1)}}{(\gamma + L z)^{\lambda(L - \alpha)}} \, dz.
\]

Using the substitution \( t = \frac{L z}{\gamma} \), \( z = \frac{\gamma t}{L} \), and \( dz = \frac{\gamma}{L} \, dt \), we get:

\[
I = C^\lambda \cdot \frac{\gamma^{1 + \lambda(\alpha - 1)}}{L^{1 + \lambda(L - 1)}} \int_{0}^{\infty} \frac{t^{\lambda(L - 1)}}{(1 + t)^{\lambda(L - \alpha)}} \, dt.
\]

This integral is expressed in terms of the Beta function:

\[
\int_{0}^{\infty} \frac{t^{a - 1}}{(1 + t)^{a + b}} \, dt = B(a, b), \quad \text{where} \quad a = \lambda(L - 1) + 1, \quad b = \lambda(-\alpha + 1) - 1.
\]

Thus,

\[
I = C^\lambda \cdot \frac{\gamma^{1 + \lambda(\alpha - 1)}}{L^{1 + \lambda(L - 1)}} \cdot B(a, b).
\]

We simplify the powers of \( \gamma \) and \( L \):

\[
\gamma^{1 + \lambda(\alpha - 1)} = \gamma^{1 - \lambda + \lambda \alpha}, \quad L^{1 + \lambda(L - 1)} = L^{\lambda L + 1 - \lambda}.
\]

Substituting \( C^\lambda = L^{L \lambda} \gamma^{-\alpha \lambda} \left( \frac{\Gamma(L - \alpha)}{\Gamma(-\alpha) \Gamma(L)} \right)^\lambda \), we obtain:

\[
I = \gamma^{1 - \lambda} L^{\lambda - 1} \left( \frac{\Gamma(L - \alpha)}{\Gamma(-\alpha) \Gamma(L)} \right)^\lambda B(a, b).
\]

The Rényi entropy is therefore:

\[
H_\lambda(Z) = \frac{1}{1 - \lambda} \left[ \ln \left( \gamma^{1 - \lambda} L^{\lambda - 1} \left( \frac{\Gamma(L - \alpha)}{\Gamma(-\alpha) \Gamma(L)} \right)^\lambda B(a, b) \right) \right].
\]

Simplifying:

\[
H_\lambda(G_I^0 ) = \ln \left( \frac{\gamma}{L} \right) + \frac{1}{1 - \lambda} \left[ \lambda (\ln \Gamma(L - \alpha) - \ln \Gamma(-\alpha) - \ln \Gamma(L)) + \ln B(a, b) \right].
\]

Using the Beta function property \( \ln B(a, b) = \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a + b) \), where \( a + b = \lambda(L - \alpha) \), we get the final expression:

\[
H_\lambda(G_I^0 ) = \ln \left( \frac{\gamma}{L} \right) + \frac{1}{1 - \lambda} \left[ \lambda (\ln \Gamma(L - \alpha) - \ln \Gamma(-\alpha) - \ln \Gamma(L)) + \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(\lambda(L - \alpha)) \right].
\]

Given that the mean \( \mu \) is related to \( \gamma \) by:

\[
\mu = -\frac{\gamma}{\alpha + 1} \quad \Rightarrow \quad \gamma = -\mu(\alpha + 1),
\]

substituting \( \gamma \) into the entropy expression:

\begin{multline*}
H_\lambda(G_I^0 ) = \ln\left( \frac{ -\mu(\alpha + 1) }{ L } \right)\\
+ \frac{1}{1 - \lambda} \left[ \lambda\left( \ln \Gamma(L - \alpha) - \ln \Gamma(-\alpha) - \ln \Gamma(L) \right) + \ln \Gamma\left( \lambda(L - 1) + 1 \right) + \ln \Gamma\left( \lambda(-\alpha + 1) - 1 \right) - \ln \Gamma\left( \lambda(L - \alpha) \right) \right].
\end{multline*}
Therefore:
\begin{multline}
H_\lambda(G_I^0 ) = \ln \mu+ \ln (-1-\alpha)-\ln L \\
+ \frac{1}{1 - \lambda} \left[ \lambda\left( \ln \Gamma(L - \alpha) - \ln \Gamma(-\alpha) - \ln \Gamma(L) \right) + \ln \Gamma\left( \lambda(L - 1) + 1 \right) + \ln \Gamma\left( \lambda(-\alpha + 1) - 1 \right) - \ln \Gamma\left( \lambda(L - \alpha) \right) \right].
\end{multline}

##  Derivation of the Rényi Entropy for the \( \Gamma_{\text{SAR}} \) Distribution

To compute the Rényi entropy of \( \Gamma_{\text{SAR}}(L, \mu) \), we need to evaluate the integral:

\[
I = \int_{0}^{\infty} [f_Z(z; L, \mu)]^\lambda \, dz = \left( \frac{L^L}{\Gamma(L) \mu^L} \right)^\lambda \int_{0}^{\infty} z^{\lambda(L - 1)} \, e^{-\lambda L z / \mu} \, dz.
\]

This integral can be evaluated using the Gamma integral formula:

\[
\int_{0}^{\infty} x^{p - 1} e^{-q x} \, dx = \frac{\Gamma(p)}{q^p}, \quad \text{for } p > 0, \ q > 0,
\]

where \( p = \lambda L - \lambda + 1 \) and \( q = \frac{\lambda L}{\mu} \). Substituting these values into the integral, we obtain:

\[
I = \left( \frac{L^L}{\Gamma(L) \mu^L} \right)^\lambda \frac{\Gamma(\lambda L - \lambda + 1)}{\left( \frac{\lambda L}{\mu} \right)^{\lambda L - \lambda + 1}}.
\]

The natural logarithm of the integral is then given by:

\[
\ln I = \lambda \left( L \ln L - L \ln \mu - \ln \Gamma(L) \right) + \ln \Gamma(\lambda L - \lambda + 1) - \left( \lambda L - \lambda + 1 \right) \left( \ln \lambda L - \ln \mu \right).
\]

The Rényi entropy \( H_\lambda(Z) \) can be computed by substituting \( \ln I \) into the definition:

\[
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln I.
\]

Expanding and simplifying the terms, the Rényi entropy of \( \Gamma_{\text{SAR}}(L, \mu) \) becomes:

\[
H_\lambda(\Gamma_{\text{SAR}}) = \frac{1}{\lambda - 1} \left[ \lambda \ln \Gamma(L) - \ln \Gamma\big( \lambda(L - 1) + 1 \big) - (\lambda - 1) \big( \ln L - \ln \mu \big) + \big( \lambda(L - 1) + 1 \big) \ln \lambda \right].
\]

# Derivation of the Rényi Entropy for the \(\Gamma_{\text{SAR}}(L, \mu)\) Distribution.

Let \(Z \sim \Gamma_{\text{SAR}}(L, \mu)\) with pdf
\begin{align}
f_{\Gamma_{\text{SAR}}}(z; L, \mu)
&= \frac{L^L}{\Gamma(L)\,\mu^L}\,z^{\,L - 1} 
  \exp\!\Bigl(-\tfrac{L z}{\mu}\Bigr),
\label{eq:pdf}
\end{align}
Define
\begin{align}
I 
&= \int_{0}^{\infty}\!\bigl[f_{\Gamma_{\text{SAR}}}(z; L,\mu)\bigr]^\lambda \,dz 
 = \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \int_{0}^{\infty} 
   z^{\,\lambda\,(L-1)} \exp\!\Bigl(-\tfrac{\lambda\,L}{\mu}\,z\Bigr)\,dz.
\label{eq:I}
\end{align}
Using the Gamma integral 
\(\displaystyle
  \int_{0}^{\infty} x^{p-1} e^{-qx}\,dx 
   = \frac{\Gamma(p)}{q^p},\)
with \(p = \lambda L - \lambda + 1\) and \(q = \tfrac{\lambda L}{\mu}\), it follows that
\begin{align}
I 
&= \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \frac{\Gamma(\lambda L - \lambda + 1)}
        {\Bigl(\tfrac{\lambda\,L}{\mu}\Bigr)^{\lambda L - \lambda + 1}}.
\label{eq:Iresult}
\end{align}
Taking the natural logarithm,
\begin{align}
\ln I 
&= \lambda\!\Bigl(L \ln L - L \ln \mu - \ln \Gamma(L)\Bigr)
   \;+\; \ln\Gamma\!\bigl(\lambda L - \lambda + 1\bigr)
   \;-\; \bigl(\lambda L - \lambda + 1\bigr)\,\Bigl(\ln(\lambda L) - \ln\mu\Bigr).
\label{eq:lnI}
\end{align}
The Rényi entropy is 
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln I.
\label{eq:RenyiDefinition}
\end{align}
By expanding \eqref{eq:lnI} and collecting terms in \(\ln L\) and \(\ln \mu\), one obtains
\begin{align}
\ln I 
&= (1 - \lambda)\bigl(\ln \mu - \ln L\bigr)
   \;-\;\lambda\,\ln \Gamma(L)
   \;+\;\ln\Gamma\!\bigl(\lambda(L-1)+1\bigr)
   \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln \lambda.
\label{eq:lnIsimplified}
\end{align}
Substituting \eqref{eq:lnIsimplified} into \eqref{eq:RenyiDefinition} and simplifying,
\begin{multline}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L,\mu)\bigr) 
= \ln \mu - \ln L 
  + \frac{1}{\,1-\lambda\,}
  \Bigl[
    -\lambda\,\ln\Gamma(L)
    + \ln\Gamma\!\bigl(\lambda\,(L-1)+1\bigr)
\\
    - \bigl(\lambda\,(L-1)+1\bigr)\,\ln\lambda
  \Bigr].
\label{eq:RenyiFinal}
\end{multline}
This completes the derivation.

# Derivation of the Rényi Entropy for the $\Gamma_{\text{SAR}}(L, \mu)$ Distribution.

Let $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ with pdf
\begin{align*}
f_{\Gamma_{\text{SAR}}}(z; L, \mu)
&= \frac{L^L}{\Gamma(L)\,\mu^L}\,z^{\,L - 1} 
  \exp\!\Bigl(-\tfrac{L z}{\mu}\Bigr).
\end{align*}
Define
\begin{align*}
I 
&= \int_{0}^{\infty}\!\bigl[f_{\Gamma_{\text{SAR}}}(z; L,\mu)\bigr]^\lambda \,dz 
 = \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \int_{0}^{\infty} 
   z^{\,\lambda\,(L-1)} \exp\!\Bigl(-\tfrac{\lambda\,L}{\mu}\,z\Bigr)\,dz.
\end{align*}
Using the Gamma integral 
$\displaystyle
  \int_{0}^{\infty} x^{p-1} e^{-qx}\,dx 
   = \frac{\Gamma(p)}{q^p},$
with $p = \lambda L - \lambda + 1$ and $q = \tfrac{\lambda L}{\mu}$, it follows that
\begin{align*}
I 
&= \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \frac{\Gamma(\lambda L - \lambda + 1)}
        {\Bigl(\tfrac{\lambda\,L}{\mu}\Bigr)^{\lambda L - \lambda + 1}}.
\end{align*}
Taking the natural logarithm,
\begin{align}
\ln I 
&= \lambda\!\Bigl(L \ln L - L \ln \mu - \ln \Gamma(L)\Bigr)
   \;+\; \ln\Gamma\!\bigl(\lambda L - \lambda + 1\bigr)
   \;-\; \bigl(\lambda L - \lambda + 1\bigr)\,\Bigl(\ln(\lambda L) - \ln\mu\Bigr).
\label{eq:lnI}
\end{align}
The Rényi entropy is 
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln I.
\label{eq:RenyiDefinition}
\end{align}
By expanding \eqref{eq:lnI} and collecting terms in $\ln L$ and $\ln \mu$, one obtains
\begin{align}
\ln I 
&= (1 - \lambda)\bigl(\ln \mu - \ln L\bigr)
   \;-\;\lambda\,\ln \Gamma(L)
   \;+\;\ln\Gamma\!\bigl(\lambda(L-1)+1\bigr)
   \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln \lambda.
\label{eq:lnIsimplified}
\end{align}
Substituting \eqref{eq:lnIsimplified} into \eqref{eq:RenyiDefinition} and simplifying,
\begin{multline}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L,\mu)\bigr) 
= \ln \mu - \ln L 
  + \frac{1}{\,1-\lambda\,}
  \Bigl[
    -\lambda\,\ln\Gamma(L)
    + \ln\Gamma\!\bigl(\lambda\,(L-1)+1\bigr)
    - \bigl(\lambda\,(L-1)+1\bigr)\,\ln\lambda
  \Bigr].
\label{eq:RenyiFinal}
\end{multline}
This completes the derivation.

# Derivation of the Rényi Entropy for the $G_I^0$ Distribution

The Rényi entropy of order $\lambda$ for a continuous random variable $Z$ with density $f_Z(z)$ is given by
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln \biggl( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \biggr), 
\quad \lambda > 0, \,\lambda \neq 1.
\label{eq:RenyiDefinition}
\end{align}

For $Z \sim G_I^0(\alpha, \gamma, L)$, the density function is
\begin{align}
f_{G_I^0}(z; \alpha, \gamma, L) 
&= \frac{L^L\,\Gamma(L-\alpha)}{\bigl[-\gamma\bigr]^{\alpha}\,\Gamma(-\alpha)\,\Gamma(L)}
   \,\frac{z^{\,L-1}}{\bigl(\gamma + L\,z\bigr)^{\,L-\alpha}}
\label{eq:pdfGI0}
\end{align}
where $\gamma > 0$ in this parametrization (noting $\gamma = -\mu(\alpha+1)$ for a negative $\mu$ when $\alpha<-1$).

Raising $f_Z(z)$ to the power $\lambda$ and substituting into the entropy integral,
\[
I 
= \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz 
= C^\lambda \int_{0}^{\infty} 
  \frac{z^{\lambda(L - 1)}}
       {\bigl(\gamma + L z\bigr)^{\,\lambda(L - \alpha)}} \, dz,
\]
where 
\[
C = \frac{L^L \,\Gamma(L - \alpha)}{\gamma^\alpha \,\Gamma(-\alpha)\,\Gamma(L)}.
\]
Using the change of variables $t = \frac{Lz}{\gamma}$, $z = \frac{\gamma\,t}{L}$, $dz = \frac{\gamma}{L}\,dt$, one obtains
\[
I 
= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
  \int_{0}^{\infty} \frac{t^{\lambda(L - 1)}}{\bigl(1 + t\bigr)^{\lambda(L - \alpha)}} \, dt.
\]
The integral is expressed in terms of the Beta function
\[
\int_{0}^{\infty} \frac{t^{\,a - 1}}{(1 + t)^{\,a + b}}\,dt
= B(a,b),
\]
with $a = \lambda(L - 1) + 1$ and $b = \lambda(-\alpha + 1) - 1$. Thus
\begin{align}
I 
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}} \,B(a,b).
\label{eq:Iresult}
\end{align}
By simplifying powers of $\gamma$ and $L$, and substituting 
$C^\lambda = L^{\lambda L} \,\gamma^{-\alpha \lambda}
   \Bigl(\frac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda,$
one obtains
\[
I 
= \gamma^{\,1 - \lambda}\,L^{\,\lambda - 1}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^{\!\lambda}
  B(a,b).
\]
Hence, by definition \eqref{eq:RenyiDefinition},
\[
H_\lambda(Z) 
= \frac{1}{\,1 - \lambda\,}\,
  \ln\Bigl[\gamma^{\,1 - \lambda}\,L^{\,\lambda - 1}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^{\!\lambda}
  B(a,b)\Bigr].
\]
Rewriting,
\[
H_\lambda(G_I^0 ) 
= \ln\!\Bigl(\tfrac{\gamma}{\,L}\Bigr)
  \;+\; \frac{1}{\,1 - \lambda\,}
  \Bigl[\,\lambda\bigl(\ln \Gamma(L - \alpha) - \ln \Gamma(-\alpha) - \ln \Gamma(L)\bigr)
         \;+\;\ln B(a,b)\Bigr].
\]
Using the property $\ln B(a,b) = \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a+b)$, where 
$a + b = \lambda(L - \alpha),$ we get
\[
H_\lambda(G_I^0 ) 
= \ln\!\Bigl(\tfrac{\gamma}{L}\Bigr)
  + \frac{1}{\,1 - \lambda\,}\biggl[\lambda\Bigl(\ln \Gamma(L - \alpha) 
     - \ln \Gamma(-\alpha) 
     - \ln \Gamma(L)\Bigr)
  + \ln \Gamma(a) 
  + \ln \Gamma(b) 
  - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)\biggr].
\]
Finally, recalling $\mu = -\frac{\gamma}{\alpha + 1}$ gives $\gamma = -\mu(\alpha+1)$, and substituting into the above expression:
\begin{multline}
H_\lambda(G_I^0 ) 
= \ln \mu + \ln(-\alpha - 1) - \ln L \\
+ \frac{1}{\,1 - \lambda\,}\Bigl[
   \lambda\bigl(\ln \Gamma(L - \alpha) 
       - \ln \Gamma(-\alpha) 
       - \ln \Gamma(L)\bigr)
   + \ln \Gamma\bigl(\lambda(L - 1) + 1\bigr) 
   + \ln \Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) 
   - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\label{eq:RenyiFinalGI0}
\end{multline}

# Derivation of the Rényi Entropy for the $G_I^0$ Distribution

The Rényi entropy of order $\lambda$ for a continuous random variable $Z$ with density $f_Z(z)$ is given by
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln \Bigl( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \Bigr),
\quad \lambda > 0, \;\lambda \neq 1.
\label{eq:RenyiDefinition}
\end{align}

For $Z \sim G_I^0(\alpha, \gamma, L)$, the density function is
\begin{align}
f_{G_I^0}(z; \alpha, \gamma, L) 
&= \frac{L^L\,\Gamma(L-\alpha)}{\bigl[-\gamma\bigr]^{\alpha}\,\Gamma(-\alpha)\,\Gamma(L)}
   \,\frac{z^{\,L-1}}{\bigl(\gamma + L\,z\bigr)^{\,L-\alpha}}
\label{eq:pdfGI0}
\end{align}
where $\gamma > 0$ in this parametrization (equivalently $\gamma = -\mu(\alpha+1)$ for a negative $\mu$ when $\alpha<-1$).

\medskip

\noindent
\textbf{Definition of the integral.}
Define
\begin{align}
I 
&= \int_{0}^{\infty} \bigl[ f_{G_I^0}(z; \alpha, \gamma, L) \bigr]^\lambda \,dz 
= C^\lambda \int_{0}^{\infty} 
  \frac{z^{\,\lambda(L - 1)}}
       {\bigl(\gamma + L\,z\bigr)^{\,\lambda(L - \alpha)}} 
  \, dz,
\label{eq:I}
\end{align}
where
\[
C 
= \frac{L^L \,\Gamma(L - \alpha)}{\gamma^\alpha \,\Gamma(-\alpha)\,\Gamma(L)}.
\]

To evaluate $I$, we use the change of variables $t = \tfrac{Lz}{\gamma}$, $z = \tfrac{\gamma\,t}{L}$, $dz = \tfrac{\gamma}{L}\,dt$. This gives
\begin{align}
I 
&= C^\lambda 
   \int_{0}^{\infty} 
   \Bigl(\tfrac{\gamma\,t}{L}\Bigr)^{\lambda(L - 1)}
   \,\Bigl(\gamma + L\,\tfrac{\gamma\,t}{L}\Bigr)^{-\lambda(L - \alpha)} 
   \,\tfrac{\gamma}{L}\, dt
\nonumber\\
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
  \int_{0}^{\infty} 
  \frac{t^{\lambda(L - 1)}}{(1 + t)^{\lambda(L - \alpha)}}
  \, dt.
\end{align}
Using the Beta-function result 
\[
\int_{0}^{\infty} \frac{t^{\,a - 1}}{(1 + t)^{\,a + b}} \, dt 
= B(a,b),
\]
with $a = \lambda(L - 1) + 1$ and $b = \lambda(-\alpha + 1) - 1$, we have
\begin{align}
I 
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \,B(a,b).
\label{eq:Ibeta}
\end{align}
Next, we simplify $\gamma^{\,1 + \lambda(\alpha - 1)} = \gamma^{\,1 - \lambda + \lambda\alpha}$ and 
$L^{\,1 + \lambda(L - 1)} = L^{\,\lambda L + 1 - \lambda}$. 
Since 
\[
C^\lambda 
= \Bigl(\tfrac{L^L}{\gamma^\alpha \,\Gamma(-\alpha)\,\Gamma(L)}\,\Gamma(L - \alpha)\Bigr)^\lambda
= L^{\lambda L} \,\gamma^{-\alpha \lambda}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda,
\]
we obtain
\[
I
= \gamma^{\,1 - \lambda}\,L^{\,\lambda - 1}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
  \;B(a,b).
\]

\smallskip

\noindent
\textbf{Rényi entropy.}
By \eqref{eq:RenyiDefinition},
\[
H_\lambda(Z) 
= \frac{1}{\,1 - \lambda\,} \,\ln I
= \frac{1}{\,1 - \lambda\,}
  \,\ln\!\Bigl[
    \gamma^{\,1 - \lambda}\,
    L^{\,\lambda - 1}
    \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
    B(a,b)
  \Bigr].
\]
Hence,
\[
H_\lambda(G_I^0)
= \ln\!\Bigl(\tfrac{\gamma}{\,L}\Bigr)
  \;+\; \frac{1}{\,1 - \lambda\,}
  \Bigl[\,
    \lambda\bigl(\ln \Gamma(L - \alpha) 
          - \ln \Gamma(-\alpha) 
          - \ln \Gamma(L)\bigr)
    \;+\; \ln B(a,b)
  \Bigr].
\]
Using the property 
$\ln B(a,b) = \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a + b)$, 
where $a + b = \lambda(L - \alpha)$, one obtains
\[
H_\lambda(G_I^0)
= \ln\!\Bigl(\tfrac{\gamma}{L}\Bigr)
 + \frac{1}{\,1 - \lambda\,}
   \Bigl[
     \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
   + \ln \Gamma(a)
   + \ln \Gamma(b)
   - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
   \Bigr].
\]
Finally, noting $\mu = -\tfrac{\gamma}{\alpha + 1}$ implies $\gamma = -\mu(\alpha + 1)$, and substituting, we get
\begin{multline}
H_\lambda(G_I^0)
= \ln \mu \;+\; \ln(-\alpha - 1) \;-\; \ln L 
\\
+ \frac{1}{\,1 - \lambda\,}\Bigl[
   \lambda\Bigl(\ln \Gamma(L - \alpha) 
       - \ln \Gamma(-\alpha) 
       - \ln \Gamma(L)\Bigr)
   + \ln \Gamma\bigl(\lambda(L - 1) + 1\bigr) 
   + \ln \Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) 
   - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
\Bigr].
\label{eq:RenyiGI0Final}
\end{multline}