### 2025-05-20 Text/Articles/RENYI_ENTROPY_SAR_2025/Reviews/Responses-Reviewer2.tex

\documentclass[11pt]{report}

\usepackage[pdftex]{geometry}
\geometry{a4paper,left=2.8cm,right=2.8cm,top=1.5cm,bottom=1.5cm,twoside}

\usepackage{amssymb}
\usepackage{enumerate}
\usepackage{graphicx}
	\graphicspath{./Comments/}
\usepackage[many]{tcolorbox}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{subfigure}
\usepackage{bm,bbm}
\usepackage{float}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\newtcolorbox{reviewbox}[1]{
colback = white!5!white,
colframe = purple!75!black,
fonttitle=\bfseries,
title=#1
}

\newtcolorbox{responsebox}[1]{
	colback = white!5!white,
	colframe = white!75!black,
	fonttitle=\bfseries,
	title=#1
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

\begin{center}
\large{\textbf{Response Letter to Reviewer \# 2}}

\vglue 0.3cm

\huge{ Quantifying Roughness in SAR Imagery with the Rényi Entropy\\ (GRSL-00722-2025)}
\end{center}

%\authors
\begin{center}
\textbf{Janeth Alpala,   Abraao D.\ C.\ Nascimento, Alejandro C.\ Frery }
\end{center}

\date{\today}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{2cm}
\noindent Dear Editors
\bigskip

\noindent We are grateful to the reviewers for the valuable comments and the time spent on checking our manuscript. 
We have addressed their comments towards improving the paper contents and presentation. 
Below we detail the changes and updates incorporated into the manuscript in this round of review.

\medskip

%-------------------------------------
\begin{reviewbox}{Comment 1}
1. The language needs further improvement and should be reviewed by a language expert.
\end{reviewbox}

\begin{responsebox}{Response}
Text

\end{responsebox}

\begin{reviewbox}{Comment 2}
2. The scientific content of the article is weak and requires further expansion.
\end{reviewbox}

\begin{responsebox}{Response}

Text

\end{responsebox}

\begin{reviewbox}{Comment 3}
 3. I did not observe any comparison of the authors' work's novelty index with the latest articles published in the field for the current years.
\end{reviewbox}

\begin{responsebox}{Response}

Text

\end{responsebox}
\begin{reviewbox}{Comment 4}
4. Conclusion requires redrafting.
\end{reviewbox}
\begin{responsebox}{Response}
Text

\end{responsebox}

\end{document}