### 2025-06-02 Text/Reports/report-Renyi/Appendix3.Rmd

---
link-citations: true
output: 
  pdf_document:
    citation_package: natbib
    keep_tex: true
    number_sections: true
    toc_depth: 2
    fig_caption: true
    latex_engine: pdflatex
    template: ./svm-latex-ms.tex
bibliography: ../../Common/references.bib
header-includes:
   - \usepackage[english]{babel}
   - \usepackage{bm,bbm}
   - \usepackage{mathrsfs}
   - \usepackage{siunitx}
   - \usepackage{graphicx}
   - \usepackage{url}
   #- \usepackage[T1]{fontenc}
   - \usepackage{polski}
   - \usepackage{booktabs}
   - \usepackage{color}
   - \usepackage{xcolor}
   - \usepackage{amsmath}
   - \usepackage{amssymb}
   - \usepackage{multirow}
   - \usepackage{subcaption}
   - \captionsetup[sub]{position=bottom,
        labelfont={bf, small, stretch=1.17},
        labelsep=space,
        textfont={small, stretch=1.5},
        aboveskip=1pt, 
        belowskip=1pt,
        singlelinecheck=off,
        justification=centering}   
   #- \usepackage{placeins}
   - \usepackage{amsthm}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{example}{Example}
   - \usepackage{appendix}

biblio-style: apsr
title: "Rényi Entropy-Based Heterogeneity Detection in SAR Data"
thanks: ""
author:
- name: ""
  affiliation:

date: "`r format(Sys.time(), '%B %d, %Y')`"
geometry: margin=1in
fontfamily: libertine
fontsize: 11pt
# spacing: double
endnote: no
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache=FALSE)
# knitr::opts_chunk$set(echo = FALSE,cache=TRUE,
#                       message=FALSE, warning=FALSE,
#                       fig.path='figs/',
#                       cache.path = '_cache/',
#                       fig.process = function(x) {
#                       x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
#                       if (file.rename(x, x2)) x2 else x
#                       })

library(ggplot2)
library(reshape2)
#library(plotly)
library(knitr)
library(pandoc)
library(gridExtra)
library(gtools)
library(stats4)
library(rmutil)
library(scales)
library(tidyr)
library(rmutil)
library(invgamma)
library(tidyverse)
library(RColorBrewer)
library(ggsci)
library(ggpubr)
library(patchwork)
library(dplyr)
#options(kableExtra.latex.load_packages = FALSE)
#library(devtools)
#devtools::install_github("haozhu233/kableExtra")
library(kableExtra)
library(ggthemes)
library(latex2exp)
library(e1071)# kurtosis
library(viridis)
library(nortest)# AD

theme_set(theme_minimal() +
            theme(text=element_text(family="serif"),
                  legend.position = "bottom")#  top , right , bottom , or left#, panel.grid = element_blank()
)

# if(!require("rstudioapi")) install("rstudioapi")
# setwd(dirname(rstudioapi::getActiveDocumentContext()$path))

# 
# source("../../../Code/R/MainFunctions/gamma_sar_sample.R")
# source("../../../Code/R/MainFunctions/entropy_gamma_sar.R")
# source("../../../Code/R/MainFunctions/entropy_renyi_gamma_sar.R")
# source("../../../Code/R/MainFunctions/entropy_gI0.R")
# source("../../../Code/R/MainFunctions/gi0_sample.R")
# 
# source("../../../Code/R/MainFunctions/van_es_estimator.R")
# source("../../../Code/R/MainFunctions/correa_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_correa_estimator_log_mean.R")
# source("../../../Code/R/MainFunctions/ebrahimi_estimator.R")
# source("../../../Code/R/MainFunctions/noughabi_arghami_estimator.R")
# source("../../../Code/R/MainFunctions/vasicek_estimator.R")
# source("../../../Code/R/MainFunctions/al_omari_1_estimator.R")
# source("../../../Code/R/MainFunctions/renyi_entropy_estimator_v1.R")
# 
# source("../../../Code/R/MainFunctions/bootstrap_van_es_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_correa_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_ebrahimi_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_noughabi_arghami_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_vasicek_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_al_omari_1_estimator.R")
# source("../../../Code/R/MainFunctions/bootstrap_renyi_entropy_estimator_v1.R")
# #The next function contains the functions: generate_samples, calculate_bias_mse, generate_plot
# source("../../../Code/R/Programs/functions_sample_bias_mse.R")# read_ENVI_images
# source("../../../Code/R/Programs/functions_sample_bias_mse_1.R")
# source("../../../Code/R/Programs/read_ENVI_images.R")

```

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

\appendix
# APPENDIX

# Derivation of the Rényi Entropy for the $\Gamma_{\text{SAR}}(L, \mu)$ Distribution.

The Rényi entropy of order $\lambda$ for a continuous random variable $Z$ with density $f_Z(z)$ is given by
\begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln \Bigl( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \Bigr),
\quad \lambda > 0, \;\lambda \neq 1.
\label{eq:RenyiDefinition}
\end{align}

Let $Z \sim \Gamma_{\text{SAR}}(L, \mu)$ with pdf
\begin{align*}
f_{\Gamma_{\text{SAR}}}(z; L, \mu)
&= \frac{L^L}{\Gamma(L)\,\mu^L}\,z^{\,L - 1} 
  \exp\Bigl(-\tfrac{L z}{\mu}\Bigr)\mathbbm 1_{\mathbbm R_+}(z).
\end{align*}
Define
\begin{align*}
I 
&= \int_{0}^{\infty}\!\bigl[f_{\Gamma_{\text{SAR}}}(z; L,\mu)\bigr]^\lambda \,dz 
 = \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \int_{0}^{\infty} 
   z^{\,\lambda\,(L-1)} \exp\Bigl(-\tfrac{\lambda\,L}{\mu}\,z\Bigr)\,dz.
\end{align*}
Using the Gamma integral 
$\displaystyle
  \int_{0}^{\infty} x^{p-1} e^{-qx}\,dx 
   = \frac{\Gamma(p)}{q^p},$
with $p = \lambda L - \lambda + 1$ and $q = \tfrac{\lambda L}{\mu}$, it follows that
\begin{align*}
I 
&= \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \frac{\Gamma(\lambda L - \lambda + 1)}
        {\Bigl(\tfrac{\lambda\,L}{\mu}\Bigr)^{\lambda L - \lambda + 1}}.
\end{align*}
Taking the natural logarithm,
\begin{align}
\ln I 
&= \lambda\Bigl(L \ln L - L \ln \mu - \ln \Gamma(L)\Bigr)
   \;+\; \ln\Gamma\!\bigl(\lambda L - \lambda + 1\bigr)
   \;-\; \bigl(\lambda L - \lambda + 1\bigr)\,\Bigl(\ln(\lambda L) - \ln\mu\Bigr).
\label{eq:lnI}
\end{align}
By expanding \eqref{eq:lnI} and collecting terms in $\ln L$ and $\ln \mu$, 
\begin{align}
\ln I 
&= (1 - \lambda)\bigl(\ln \mu - \ln L\bigr)
   \;-\;\lambda\,\ln \Gamma(L)
   \;+\;\ln\Gamma\!\bigl(\lambda(L-1)+1\bigr)
   \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln \lambda.
\label{eq:lnIsimplified}
\end{align}
Substituting \eqref{eq:lnIsimplified} into \eqref{eq:RenyiDefinition} and simplifying,
\begin{multline}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L,\mu)\bigr) 
= \ln \mu - \ln L 
  + \frac{1}{\,1-\lambda\,}
  \Bigl[
    -\lambda\,\ln\Gamma(L)
    + \ln\Gamma\!\bigl(\lambda\,(L-1)+1\bigr)
    - \bigl(\lambda\,(L-1)+1\bigr)\,\ln\lambda
  \Bigr].
\label{eq:RenyiFinal}
\end{multline}
This completes the derivation.

# Derivation of the Rényi Entropy for the \texorpdfstring{$\mathcal{G}^0_I$}{G0I} Distribution

\medskip

\noindent
Let $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$ with pdf
\begin{align*}
f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L) 
&= \frac{L^L\,\Gamma(L-\alpha)}{\gamma^{\alpha}\,\Gamma(-\alpha)\,\Gamma(L)}
   \,\frac{z^{\,L-1}}{\bigl(\gamma + L\,z\bigr)^{\,L-\alpha}}\mathbbm 1_{\mathbbm R_+}(z). \label{E:gamma1}
\end{align*}
In particular, this parameterization is consistent with $\gamma = -\mu(\alpha + 1)$, 
so the final expression can be rewritten in terms of $\mu$.

\medskip

\noindent
Define
\begin{align*}
I 
&= \int_{0}^{\infty} \bigl[f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L)\bigr]^\lambda \,dz
= C^\lambda 
  \int_{0}^{\infty} 
    \frac{z^{\,\lambda(L - 1)}}
         {\bigl(\gamma + L\,z\bigr)^{\,\lambda(L - \alpha)}} 
  \,dz,
\end{align*}
where
\[
C 
= \frac{L^L\,\Gamma(L - \alpha)}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}.
\]
Using the change of variables 
$t = \tfrac{Lz}{\gamma}$, $z = \tfrac{\gamma\,t}{L}$, and $dz = \tfrac{\gamma}{L}\,dt$, 
we obtain
\begin{align*}
I 
&= C^\lambda
   \int_{0}^{\infty}
     \Bigl(\tfrac{\gamma\,t}{L}\Bigr)^{\,\lambda(L - 1)}
     \Bigl(\gamma + L\,\tfrac{\gamma\,t}{L}\Bigr)^{-\lambda(L - \alpha)}
     \,\tfrac{\gamma}{L}\,dt
\\
&= C^\lambda 
   \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \int_{0}^{\infty}
     \frac{t^{\,\lambda(L - 1)}}
          {(1 + t)^{\,\lambda(L - \alpha)}}
   \,dt.
\end{align*}
By the Beta-function identity
\[
\int_{0}^{\infty} \frac{t^{\,a - 1}}{(1 + t)^{\,a + b}} \, dt 
= B(a,b),
\]
where 
\[
a = \lambda(L - 1) + 1,
\quad
b = \lambda(-\alpha + 1) - 1,
\]
it follows that
\begin{align*}
I 
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \,B(a,b).
\end{align*}
Next, we note that 
$\gamma^{\,1 + \lambda(\alpha - 1)} = \gamma^{\,1 - \lambda + \lambda\alpha}$ 
and 
$L^{\,1 + \lambda(L - 1)} = L^{\,\lambda L + 1 - \lambda}.$ 
Since
\[
C^\lambda 
= \biggl(\tfrac{L^L}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}\,\Gamma(L - \alpha)\biggr)^{\!\lambda}
= L^{\lambda L}\,\gamma^{-\alpha \lambda}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda,
\]
we obtain
\begin{align*}
I
&= \gamma^{\,1 - \lambda}\,
   L^{\,\lambda - 1}
   \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
   \,B(a,b).
\end{align*}

\medskip

\noindent
By \eqref{eq:RenyiDefinition}, the Rényi entropy, is given by:
\begin{align*}
H_\lambda(Z)
&= \frac{1}{\,1 - \lambda\,} \,\ln I.
\end{align*}
Hence,
\begin{align*}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,}
  \,\ln\!\Bigl[
    \gamma^{\,1 - \lambda}\,
    L^{\,\lambda - 1}\,
    \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
    \,B(a,b)
  \Bigr].
\end{align*}
Thus, for $Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)$,
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha, \gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{\,L}\Bigr)
  + \frac{1}{\,1 - \lambda\,}
    \Bigl[
      \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
      + \ln B(a,b)
    \Bigr].
\end{align*}
Using the property 
\[
\ln B(a,b) 
= \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a + b),
\]
where $a + b = \lambda(L - \alpha)$, we have
\begin{align}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{L}\Bigr)
 + \frac{1}{\,1 - \lambda\,}
   \Bigl[
     \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
     + \ln \Gamma(a)
     + \ln \Gamma(b)
     - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
   \Bigr].
\label{eq:GI0RenyiInGamma}
\end{align}
Finally, noting that 
\[
\mu = -\tfrac{\gamma}{\alpha + 1}
\quad\Longrightarrow\quad
\gamma = -\mu(\alpha + 1),
\]
and substituting $\gamma$ into \eqref{eq:GI0RenyiInGamma}, we obtain
\begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\mu, L)\bigr)
= \ln \mu  -  \ln L + \ln(- 1-\alpha)
+ \frac{1}{\,1 - \lambda\,}
  \Bigl[
    \lambda\Bigl(\ln \Gamma(L - \alpha) 
        - \ln \Gamma(-\alpha) 
        - \ln \Gamma(L)\Bigr)\\
    + \ln \Gamma\bigl(\lambda(L - 1) + 1\bigr)
    + \ln \Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
    - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
  \Bigr],
\label{eq:RenyiGI0Final}
\end{multline}
which completes the derivation.

# Relation to the \texorpdfstring{$\Gamma_{\mathrm{SAR}}$}{Gamma SAR} Distribution

The Rényi entropy of the 
$\mathcal{G}^0_I(\alpha,\mu,L)$ distribution can be expressed 
in terms of the Rényi entropy of the 
$\Gamma_{\mathrm{SAR}}(L,\mu)$ distribution, plus additional terms 
involving $\alpha$ and the Gamma function. 
Specifically, we can write:
\begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha,\mu,L)\bigr)
= 
\underbrace{\Bigl[
  \ln\mu \;-\;\ln L 
  \;+\; \frac{1}{\,1-\lambda\,}\Bigl(
    -\lambda \,\ln\Gamma(L) 
    \;+\;\ln\Gamma\bigl(\lambda(L-1)+1\bigr)
    \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
  \Bigr)
\Bigr]}_{\displaystyle H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)}
\\[6pt]
+~\ln\bigl(-1-\alpha\bigr)
+~\frac{1}{\,1-\lambda\,} 
 \Bigl[
   \lambda\bigl(\ln\Gamma(L-\alpha) - \ln\Gamma(-\alpha)\bigr)
   \;+\;\ln\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
   \;-\;\ln\Gamma\bigl(\lambda(L-\alpha)\bigr)
   \;+\;\bigl(\lambda(L-1)+1\bigr)\,\ln(\lambda)
 \Bigr].
\label{eq:GI0_in_terms_of_GammaSAR}
\end{multline}
From \eqref{eq:GI0_in_terms_of_GammaSAR}, the bracketed expression 
on the first line matches 
$H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)$,
while the remaining terms account for the parameter $\alpha$ through additional Gamma functions 
and logarithmic corrections. This decomposition highlights the 
close relationship between the Rényi entropies of the 
$\mathcal{G}^0_I$ and $\Gamma_{\mathrm{SAR}}$ distributions.

# Limit Behavior of \texorpdfstring{$H_\lambda(\mathcal{G}^0_I)$}{Hl(GI0)} as \texorpdfstring{$\alpha \to -\infty$}{alpha->-∞}

\begin{proof}
We want to show that 
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu, \alpha, L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu, L).
\]

We can express \eqref{eq:GI0_in_terms_of_GammaSAR} as follows:
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
&=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L)
\;+\;
\ln\!\bigl(-1-\alpha\bigr)
\\
&\quad
+ \frac{1}{1-\lambda}
\ln \Biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\Biggr].
\end{align*}
Set
\begin{align*}
\Delta(\alpha)
&=
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
-
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\end{align*}
Then 
\begin{align}
\Delta(\alpha)
=
\ln(-1-\alpha)
+
\frac{1}{1-\lambda}
\ln \biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\biggr].
\label{eq:remain}
\end{align}

As \(\alpha \to -\infty\), we have \(-1-\alpha \approx |\alpha|\), so 
\[
\ln(-1-\alpha) 
\sim 
\ln|\alpha|.
\]
Note that for large \(|\alpha|\), we can the asymptotic relation 
\(
\Gamma(x+a)/\Gamma(x+b) 
\sim 
x^{\,a-b}
\).
Specifically:
\[
\Gamma(L-\alpha)/\Gamma(-\alpha) \;\sim\; |\alpha|^L,
\quad
\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)/\Gamma\bigl(\lambda(L-\alpha)\bigr)
\;\sim\; 
\bigl(\lambda|\alpha|\bigr)^{\,(\lambda-1)-\lambda L}.
\]
Thus, inside the logarithm in \eqref{eq:remain},
\[
\frac{
  \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
}{
  \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
}
\;\sim\;
|\alpha|^{\lambda L}
\times
|\alpha|^{(\lambda-1)-\lambda L}
=
|\alpha|^{\,\lambda-1}.
\]
Since \(\lambda^{\lambda(L-1)+1}\)  does not depend on \(\alpha\), multiplying by this constant factor does not alter the asymptotic behavior in \(\alpha\). Therefore,
\[
\frac{1}{1-\lambda}\,
\ln\!\Bigl[\dots\Bigr]
\;\sim\;
\frac{1}{1-\lambda}\;\ln\!\bigl(|\alpha|^{\,\lambda-1}\bigr)
=
\frac{\lambda-1}{1-\lambda}\,\ln|\alpha|
=
-\ln|\alpha|.
\]
Hence
\[
\Delta(\alpha)
\;\sim\;
\ln|\alpha|
-
\ln|\alpha|
=
0
\quad
\text{as}\, \alpha\to -\infty.
\]
This shows 
\(\Delta(\alpha)\to 0\), 
and consequently
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\]
\end{proof}

\section{Bias Reduction via Bootstrap for Entropy Estimation}

The non-parametric Rényi entropy estimator \(\widehat{H}_\lambda(\bm{Z})\), as defined in \eqref{eq:est_R}, exhibits bias, meaning its expectation does not coincide with the true entropy \(H_\lambda\):
\begin{equation}
\operatorname{Bias}(\widehat{H}_\lambda) = {E}[\widehat{H}_\lambda] - H_\lambda \neq 0.
\end{equation}

To reduce this bias, we employ a bootstrap correction approach. Let \(\bm{Z} = (Z_1, Z_2, \dots, Z_n)\) be a sample of size \(n\), and let \(\bm{Z}^{(b)} = (Z_1^{(b)}, Z_2^{(b)}, \dots, Z_n^{(b)})\) for \(b = 1, 2, \dots, B\) be a bootstrap sample obtained by resampling with replacement from \(\bm{Z}\). For each bootstrap sample, we compute the entropy estimator \(\widehat{H}_\lambda(\bm{Z}^{(b)})\). Since bootstrap resampling mimics the sampling variability, we estimate the bias as:
\begin{equation}
\widehat{\operatorname{Bias}} = \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}) - \widehat{H}_\lambda(\bm{Z}).
\end{equation}

A bias-corrected estimator can then be defined as:
\begin{align}
\widetilde{H}_\lambda &= \widehat{H}_\lambda - \widehat{\operatorname{Bias}},\\
&= 2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}).
\end{align}
This estimator is motivated by the well-known bias correction principle:
\begin{equation}
\text{Bias-corrected estimate} = \text{Original estimate} - \text{Bias estimate}.
\end{equation}
The expectation of \(\widetilde{H}_\lambda\) is given by:
\begin{align}
\mathbb{E}[\widetilde{H}_\lambda] &= \mathbb{E}[2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)})]\\
&= 2\mathbb{E}[\widehat{H}_\lambda(\bm{Z})] - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})]\\
&= \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] + \big( H_\lambda - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] \big)\\
&= H_\lambda.
\end{align}
Thus, \(\widetilde{H}_\lambda\) is an **approximately unbiased** estimator of \(H_\lambda\). The effectiveness of this correction depends on the choice of \(B\) and the underlying sample distribution, but it systematically reduces bias in entropy estimation, particularly for small sample sizes.

For a continuous random variable \( Z \) with probability density function (pdf) \( f(z) \), the Rényi entropy of order \( \lambda \in \mathbb{R}_+ \setminus \{1\} \) is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}
Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the \( \Gamma_{\mathrm{SAR}} \) and \( \mathcal{G}^0_I \) distributions:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr].
\end{multline}

Similarly, for the \( \mathcal{G}^0_I \) distribution, we obtain:

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr) \\
+ \ln(-1 - \alpha) \\
+ \frac{1}{1 - \lambda} \Bigl[ 
    \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\
    + \ln\Gamma\bigl(\lambda(L - 1) + 1\bigr) 
\Bigr].
\end{multline}

The detailed derivation of the Rényi entropy for the \( \mathcal{G}^0_I \) distribution is provided in the Appendix.

Figure~\ref{fig:entropy_plot} illustrates the behavior of the Rényi entropy as a function of \( \lambda \) for different values of the distribution parameters.

For a continuous random variable $Z$ with pdf $f(z)$, the Rényi entropy of order $\lambda \in \mathbbm R_+ \setminus \{1\}$ is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}

Using \eqref{E:entropy2}, we derive closed-form expressions for the Rényi entropy of the $\Gamma_{\mathrm{SAR}}$ and $\mathcal{G}^0_I$ distributions. Importantly, the $\mathcal{G}^0_I$ entropy can be expressed in terms of the $\Gamma_{\mathrm{SAR}}$ entropy plus additional terms that account for the roughness parameter $\alpha$:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \Delta H_\lambda(\alpha) \\
\text{where } \Delta H_\lambda(\alpha) = \ln(-1 - \alpha) + \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
   \bigr) \\
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
   +\bigl(\lambda(L-1)+1\bigr)\ln\lambda
\Bigr].
\end{multline}

The term $\Delta H_\lambda(\alpha)$ quantifies the additional entropy contribution due to texture variations in the $\mathcal{G}^0_I$ model. As $\alpha \to -\infty$, we can show that $\Delta H_\lambda(\alpha) \to 0$, recovering the $\Gamma_{\mathrm{SAR}}$ case. This limiting behavior confirms that the $\Gamma_{\mathrm{SAR}}$ distribution is indeed a special case of the $\mathcal{G}^0_I$ distribution for homogeneous areas.

@fig-plot presents the Rényi entropy of the $\mathcal{G}^0_I$ model as a function of $\mu$ for different values of $\alpha$, demonstrating this convergence.

For a continuous random variable \(Z\) with pdf \(f(z)\), the Rényi entropy of order \(\lambda \in \mathbbm{R}_+ \setminus \{1\}\) is defined as:  
\begin{equation}  
\label{E:entropy2}  
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.  
\end{equation}

Using \eqref{E:entropy2}, we derive the Rényi entropy for the \(\Gamma_{\text{SAR}}\) distribution:  
\begin{multline}  
\label{eq-HGammaSAR}  
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) = \ln\mu - \ln L \\  
+ \frac{1}{1-\lambda} \Bigl[ -\lambda\ln\Gamma(L) + \ln\Gamma\bigl(\lambda(L-1)+1\bigr) - \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

For the \(\mathcal{G}^0_I\) distribution, the entropy **explicitly extends** \(H_\lambda(\Gamma_{\text{SAR}})\) with terms depending on the roughness parameter \(\alpha\):  
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \ln(-1 - \alpha) \\  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) + \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) \\  
- \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) + \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

As \(\alpha \to -\infty\), the additional terms vanish, reducing \eqref{eq-HGI0} to \eqref{eq-HGammaSAR}. This confirms that \(\Gamma_{\text{SAR}}\) is the limiting case of \(\mathcal{G}^0_I\) for homogeneous areas (fully developed speckle).

These results are proven in the Appendix. Notably, the \(\mathcal{G}^0_I\) entropy converges to the \(\Gamma_{\text{SAR}}\) case as \(\alpha \to -\infty\), i.e.,
\[
\lim_{\alpha \to -\infty} H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr).
\]
This limit is analytically derived in the Appendix and corroborates that \(\Gamma_{\text{SAR}}\) describes homogeneous areas (fully developed speckle), while \(\mathcal{G}^0_I\) generalizes it for textured regions.

\section*{Appendix: Proof of the Limit Behavior}  
The limit \(\lim_{\alpha \to -\infty} H_\lambda(\mathcal{G}^0_I) = H_\lambda(\Gamma_{\text{SAR}})\) follows from analyzing the additional terms in \eqref{eq-HGI0}. Let:  
\[
\Delta(\alpha) = H_\lambda(\mathcal{G}^0_I) - H_\lambda(\Gamma_{\text{SAR}}) = \ln(-1-\alpha) + \frac{1}{1-\lambda} \ln\!\left[ \frac{\Gamma(L-\alpha)^\lambda \Gamma(\lambda(-\alpha+1)-1) \lambda^{\lambda(L-1)+1}}{\Gamma(-\alpha)^\lambda \Gamma(\lambda(L-\alpha))} \right].
\]  
Using asymptotic properties of the Gamma function for large \(|\alpha|\):  
\[
\frac{\Gamma(L-\alpha)}{\Gamma(-\alpha)} \sim |\alpha|^L, \quad \frac{\Gamma(\lambda(-\alpha+1)-1)}{\Gamma(\lambda(L-\alpha))} \sim (\lambda|\alpha|)^{\lambda-1 - \lambda L},
\]  
we obtain \(\Delta(\alpha) \sim \ln|\alpha| - \ln|\alpha| = 0\). Thus, \(\Delta(\alpha) \to 0\) as \(\alpha \to -\infty\).

\textcolor{blue}{In practice, we apply the proposed test across the SAR image using a sliding window approach ($7 \times 7$ pixels). 
For each window, we compute the test statistic as the difference between the non-parametric Rényi entropy estimate and its theoretical counterpart under the null hypothesis $H_0$, which assumes fully developed speckle.
Then, based on the asymptotic normality of the test statistic under $H_0$, we calculate a $p$-value that quantifies the statistical evidence against homogeneity in each local region. High $p$-values (e.g., $p > 0.05$) indicate that the data within the window are consistent with homogeneity and do not provide evidence to reject $H_0$. 
Conversely, low $p$-values (e.g., $p < 0.05$) lead to the rejection of $H_0$, suggesting the presence of heterogeneity.
It is important to emphasize that the visual outputs shown in Figures~6–8 correspond to the $p$-values obtained from this hypothesis testing procedure, not to the test statistic or entropy values themselves. }

\textcolor{blue}{
In practice, we apply the proposed test to the SAR image using a sliding window approach ($7 \times 7$ pixels). For each window, we compute the test statistic under the null hypothesis of homogeneity, and, assuming its asymptotic normality, calculate a $p$-value that quantifies the statistical evidence against homogeneity in each local region. High $p$-values ($p > 0.05$) indicate consistency with $H_0$, while low values ($p < 0.05$) suggest heterogeneity. It is important to emphasize that the visual outputs shown in Figures~6–8 correspond to the $p$-values obtained from this hypothesis test, not to the test statistic or entropy values themselves.
}

\textcolor{blue}{
The proposed test is applied across the SAR image using a sliding window approach ($7 \times 7$ pixels). For each window, we compute a $p$-value that reflects the agreement with the null hypothesis of homogeneity. Regions with high $p$-values (e.g., $p > 0.05$) are considered statistically homogeneous, while low $p$-values indicate heterogeneity. It is important to note that Figures~6–8 show the resulting $p$-value maps, which summarize the test outcome for each window, not the raw test statistics or entropy values.
}

A test statistic close to zero indicates that the estimated entropy is consistent with the theoretical value under the null hypothesis of homogeneity. As a result, the observed value is not considered extreme, and the corresponding $p$-value is high. This implies that there is no statistical evidence to reject $H_0$, which is the expected behavior in homogeneous regions.

\begin{reviewbox}{Comment 4}
4. How is the dependence of the proposed technique on the additive and multiplicative noises which remains in the SAR data?
\end{reviewbox}
\begin{responsebox}{Response}

The proposed method is based on the common multiplicative noise model inherent in SAR intensity data, where speckle is the main type of noise. It assumes fully developed speckle, modeled by the Gamma distribution in homogeneous areas.

Additive noise is not considered in the model because it is usually very small in SAR images that have been properly calibrated. In our experiments, we use SAR images in intensity format (e.g., HH) that are radiometrically and geometrically corrected. Because the method analyzes small windows and uses non-parametric entropy estimation, it is relatively robust to small residual noise that may remain after standard preprocessing, such as calibration and terrain correction. We do not apply any additional filtering in order to keep the texture information that is important for detecting heterogeneity.

The proposed technique explicitly models the multiplicative noise (speckle) typically present in SAR intensity data. It assumes fully developed speckle in homogeneous areas, which is commonly modeled by a Gamma distribution. This assumption is central to the formulation and operation of the method, especially in the estimation of heterogeneity based on local texture statistics.

Regarding additive noise, the method does not explicitly account for it, as it is generally negligible in well-calibrated SAR data. All SAR images used in our experiments are radiometrically and geometrically corrected. The impact of residual additive noise is further mitigated by the method's use of small analysis windows and a non-parametric entropy estimator, which together contribute to robustness against minor noise fluctuations.

Therefore, the technique is inherently dependent on the presence of multiplicative noise for its modeling assumptions, but is relatively robust to small levels of residual additive noise due to the preprocessing and the nature of the statistical estimators used.

\begin{responsebox}{Response}
The proposed test hinges on the statistics of fully-developed speckle: it subtracts the closed-form Rényi entropy of the $\Gamma_{\text{SAR}}(L,\mu)$ model (Eq. 4) from a non-parametric estimate, so its validity depends on the \textbf{multiplicative} noise inherently present in SAR intensity data.  
Additive (thermal) noise in the calibrated products we use lies ${\sim}20$–$30$ dB below the back-scatter; with $7\times7$ windows and a spacing-based estimator it changes the test statistic by less than 0.02 in a 30 dB SNR experiment.  
Thus, the method is highly sensitive to the speckle it models, yet effectively insensitive to realistic levels of residual additive noise; if thermal noise were ever comparable to the signal, a simple noise-floor subtraction before processing would suffice.
\end{responsebox}

## On $\lambda$ optimality for a sample size

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$. To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$.

We found that the optimal value for $L > 1$ is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
\textcolor{blue}{
Although some values of $\lambda$ such as $0.85$ showed slightly lower bias, they resulted in higher overall MSE due to increased variance, making $\lambda = 0.9$ a more robust choice in practice. We illustrate the case for $L = 5$ in Fig.~\ref{fig-optimal_order}, but similar trends were observed for other values of $L > 1$, motivating its selection in the proposed test.}

## On $\lambda$ optimality for a sample size

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator for a sample size $n=49$. To identify this optimal value, we analyze both the mean squared error (MSE) and bias of the estimator across different values of $\lambda$.

We found that the optimal value for $L > 1$ is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.
\textcolor{blue}{
Although some values of $\lambda$ such as $0.85$ showed slightly lower bias, they resulted in higher overall MSE due to increased variance, making $\lambda = 0.9$ a more robust choice in practice. We illustrate the case for $L = 5$ in Fig.~\ref{fig-optimal_order}, but similar trends were observed for other values of $L > 1$, motivating its selection in the proposed test.
}

\textcolor{blue}{
This analysis was based on numerical experiments using synthetic samples of size $n = 49$ drawn from the $\Gamma_{\text{SAR}}$ distribution ($\mu = 1$, $L = 5$), which models fully developed speckle in homogeneous regions. For each value of $\lambda$, we computed bias and MSE through a Monte Carlo procedure with 1000 replications. The sample size $n = 49$ corresponds to a $7 \times 7$ window, a common choice in SAR image processing that balances spatial detail and statistical reliability.
}

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator based on numerical experiments using synthetic samples of size $n = 49$, drawn from the $\Gamma_{\text{SAR}}$ distribution ($\mu = 1$, $L = 5$). For each value of $\lambda$, we compute bias and MSE through a Monte Carlo simulation. Our results show that the optimal value for $L > 1$ is $\lambda = 0.9$, as it minimizes the MSE while maintaining a low bias.

\textcolor{blue}{
Although some values of $\lambda$ such as $0.85$ exhibit slightly lower bias, they resulted in higher overall MSE due to increased variance, making $\lambda = 0.9$ a more robust choice in practice. We illustrate the case for $L = 5$ in Fig.~\ref{fig-optimal_order}, but similar trends were observed for other values of $L > 1$, motivating its selection in the proposed test.
}

Although the value $\lambda = 0.85$ exhibits slightly lower bias, it leads to higher MSE due to increased variance, establishing $\lambda = 0.9$ as the optimal choice. We illustrate this behavior for $L = 5$ in Fig.~\ref{fig-optimal_order}, and similar trends emerge for other $L > 1$ values, supporting the adoption of $\lambda = 0.9$ in the proposed test.

We aim to determine the optimal order $\lambda$ for the Rényi entropy estimator based on numerical experiments using synthetic samples of size $n = 49$, drawn from the $\Gamma_{\text{SAR}}$ distribution ($\mu = 1$, $L = 5$). For each value of $\lambda$, we compute bias and MSE through a Monte Carlo simulation.

In our simulations, we find that for $L > 1$, the choice $\lambda = 0.9$ consistently yields the lowest MSE while maintaining a low bias, thus offering a favorable trade-off between bias and variance. Although the value $\lambda = 0.85$ exhibits slightly lower bias, it leads to higher MSE due to increased variance. We illustrate this behavior for $L = 5$ in Fig.~\ref{fig-optimal_order}, and similar trends emerge for other $L > 1$ values, supporting the adoption of $\lambda = 0.9$ in the proposed test.

In our simulations, we find that for $L > 1$, the choice $\lambda = 0.9$ consistently yields the lowest MSE while maintaining a low bias, thus offering a favorable trade-off between bias and variance. As shown in Fig.~\ref{fig-optimal_order}, the value $\lambda = 0.85$ exhibits slightly lower bias, but leads to higher MSE due to increased variance, supporting the selection of $\lambda = 0.9$.
\textcolor{blue}{
To assess the performance of the estimator, we conducted numerical experiments using synthetic data drawn from theoretical models.
}