### 2025-06-02 Text/Reports/report-Renyi/Appendix3.tex

\documentclass[11pt,]{article}
\usepackage[left=1in,top=1in,right=1in,bottom=1in]{geometry}
\newcommand*{\authorfont}{\fontfamily{phv}\selectfont}
\usepackage[]{libertine}

  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}

\usepackage{abstract}
\renewcommand{\abstractname}{}    % clear the title
\renewcommand{\absnamepos}{empty} % originally center

\renewenvironment{abstract}
 {{%
    \setlength{\leftmargin}{0mm}
    \setlength{\rightmargin}{\leftmargin}%
  }%
  \relax}
 {\endlist}

\makeatletter
\def\@maketitle{%
  \newpage
%  \null
%  \vskip 2em%
%  \begin{center}%
  \let \footnote \thanks
    {\fontsize{18}{20}\selectfont\raggedright  \setlength{\parindent}{0pt} \@title \par}%
}
%\fi
\makeatother

\setcounter{secnumdepth}{0}

\title{Rényi Entropy-Based Heterogeneity Detection in SAR Data  }

\author{\Large \vspace{0.05in} \newline\normalsize\emph{}  }

\date{}

\usepackage{titlesec}

\titleformat*{\section}{\normalsize\bfseries}
\titleformat*{\subsection}{\normalsize\itshape}
\titleformat*{\subsubsection}{\normalsize\itshape}
\titleformat*{\paragraph}{\normalsize\itshape}
\titleformat*{\subparagraph}{\normalsize\itshape}

\usepackage{natbib}
\bibliographystyle{apsr}
\usepackage[strings]{underscore} % protect underscores in most circumstances

\newtheorem{hypothesis}{Hypothesis}
\usepackage{setspace}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage[english]{babel}
\usepackage{bm,bbm}
\usepackage{mathrsfs}
\usepackage{siunitx}
\usepackage{graphicx}
\usepackage{url}
\usepackage{polski}
\usepackage{booktabs}
\usepackage{color}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{multirow}
\usepackage{subcaption}
\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=1pt, belowskip=1pt, singlelinecheck=off, justification=centering}
\usepackage{amsthm}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{theorem}{Theorem}
\newtheorem{example}{Example}
\usepackage{appendix}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

% move the hyperref stuff down here, after header-includes, to allow for - \usepackage{hyperref}

\makeatletter
\@ifpackageloaded{hyperref}{}{%
\ifxetex
  \PassOptionsToPackage{hyphens}{url}\usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \PassOptionsToPackage{hyphens}{url}\usepackage[draft,unicode=true]{hyperref}
\fi
}

\@ifpackageloaded{color}{
    \PassOptionsToPackage{usenames,dvipsnames}{color}
}{%
    \usepackage[usenames,dvipsnames]{color}
}
\makeatother
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={ ()},
             pdfkeywords = {},  
            pdftitle={Rényi Entropy-Based Heterogeneity Detection in SAR
Data},
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls

% Add an option for endnotes. -----

% add tightlist ----------
\providecommand{\tightlist}{%
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% add some other packages ----------

% \usepackage{multicol}
% This should regulate where figures float
% See: https://tex.stackexchange.com/questions/2275/keeping-tables-figures-close-to-where-they-are-mentioned
\usepackage[section]{placeins}

\begin{document}

% \pagenumbering{arabic}% resets `page` counter to 1 
%

% \maketitle

{% \usefont{T1}{pnc}{m}{n}
\setlength{\parindent}{0pt}
\thispagestyle{plain}
{\fontsize{18}{20}\selectfont\raggedright 
\maketitle  % title \par

}

{
   \vskip 13.5pt\relax \normalsize\fontsize{11}{12} 
\textbf{\authorfont } \hskip 15pt \emph{\small }

}

}

\vskip -8.5pt

 % removetitleabstract

\noindent

\newcommand{\bias}{\operatorname{Bias}}
\newcommand{\widebar}[1]{\overline{#1}}

\appendix

\section{APPENDIX}\label{appendix}

\section{\texorpdfstring{Derivation of the Rényi Entropy for the
\(\Gamma_{\text{SAR}}(L, \mu)\)
Distribution.}{Derivation of the Rényi Entropy for the \textbackslash Gamma\_\{\textbackslash text\{SAR\}\}(L, \textbackslash mu) Distribution.}}\label{derivation-of-the-ruxe9nyi-entropy-for-the-gamma_textsarl-mu-distribution.}

The Rényi entropy of order \(\lambda\) for a continuous random variable
\(Z\) with density \(f_Z(z)\) is given by \begin{align}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,} \,\ln \Bigl( \int_{0}^{\infty} [f_Z(z)]^\lambda \, dz \Bigr),
\quad \lambda > 0, \;\lambda \neq 1.
\label{eq:RenyiDefinition}
\end{align}

Let \(Z \sim \Gamma_{\text{SAR}}(L, \mu)\) with pdf \begin{align*}
f_{\Gamma_{\text{SAR}}}(z; L, \mu)
&= \frac{L^L}{\Gamma(L)\,\mu^L}\,z^{\,L - 1} 
  \exp\Bigl(-\tfrac{L z}{\mu}\Bigr)\mathbbm 1_{\mathbbm R_+}(z).
\end{align*} Define \begin{align*}
I 
&= \int_{0}^{\infty}\!\bigl[f_{\Gamma_{\text{SAR}}}(z; L,\mu)\bigr]^\lambda \,dz 
 = \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \int_{0}^{\infty} 
   z^{\,\lambda\,(L-1)} \exp\Bigl(-\tfrac{\lambda\,L}{\mu}\,z\Bigr)\,dz.
\end{align*} Using the Gamma integral \(\displaystyle
  \int_{0}^{\infty} x^{p-1} e^{-qx}\,dx
   = \frac{\Gamma(p)}{q^p},\) with \(p = \lambda L - \lambda + 1\) and
\(q = \tfrac{\lambda L}{\mu}\), it follows that \begin{align*}
I 
&= \biggl(\frac{L^L}{\Gamma(L)\,\mu^L}\biggr)^{\!\lambda}
   \frac{\Gamma(\lambda L - \lambda + 1)}
        {\Bigl(\tfrac{\lambda\,L}{\mu}\Bigr)^{\lambda L - \lambda + 1}}.
\end{align*} Taking the natural logarithm, \begin{align}
\ln I 
&= \lambda\Bigl(L \ln L - L \ln \mu - \ln \Gamma(L)\Bigr)
   \;+\; \ln\Gamma\!\bigl(\lambda L - \lambda + 1\bigr)
   \;-\; \bigl(\lambda L - \lambda + 1\bigr)\,\Bigl(\ln(\lambda L) - \ln\mu\Bigr).
\label{eq:lnI}
\end{align} By expanding \eqref{eq:lnI} and collecting terms in
\(\ln L\) and \(\ln \mu\), \begin{align}
\ln I 
&= (1 - \lambda)\bigl(\ln \mu - \ln L\bigr)
   \;-\;\lambda\,\ln \Gamma(L)
   \;+\;\ln\Gamma\!\bigl(\lambda(L-1)+1\bigr)
   \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln \lambda.
\label{eq:lnIsimplified}
\end{align} Substituting \eqref{eq:lnIsimplified} into
\eqref{eq:RenyiDefinition} and simplifying, \begin{multline}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L,\mu)\bigr) 
= \ln \mu - \ln L 
  + \frac{1}{\,1-\lambda\,}
  \Bigl[
    -\lambda\,\ln\Gamma(L)
    + \ln\Gamma\!\bigl(\lambda\,(L-1)+1\bigr)
    - \bigl(\lambda\,(L-1)+1\bigr)\,\ln\lambda
  \Bigr].
\label{eq:RenyiFinal}
\end{multline} This completes the derivation.

\section{\texorpdfstring{Derivation of the Rényi Entropy for the
\texorpdfstring{$\mathcal{G}^0_I$}{G0I}
Distribution}{Derivation of the Rényi Entropy for the  Distribution}}\label{derivation-of-the-ruxe9nyi-entropy-for-the-distribution}

\medskip

\noindent Let \(Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)\) with pdf
\begin{align*}
f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L) 
&= \frac{L^L\,\Gamma(L-\alpha)}{\gamma^{\alpha}\,\Gamma(-\alpha)\,\Gamma(L)}
   \,\frac{z^{\,L-1}}{\bigl(\gamma + L\,z\bigr)^{\,L-\alpha}}\mathbbm 1_{\mathbbm R_+}(z). \label{E:gamma1}
\end{align*} In particular, this parameterization is consistent with
\(\gamma = -\mu(\alpha + 1)\), so the final expression can be rewritten
in terms of \(\mu\).

\medskip

\noindent Define \begin{align*}
I 
&= \int_{0}^{\infty} \bigl[f_{\mathcal{G}^0_I}(z; \alpha, \gamma, L)\bigr]^\lambda \,dz
= C^\lambda 
  \int_{0}^{\infty} 
    \frac{z^{\,\lambda(L - 1)}}
         {\bigl(\gamma + L\,z\bigr)^{\,\lambda(L - \alpha)}} 
  \,dz,
\end{align*} where \[
C 
= \frac{L^L\,\Gamma(L - \alpha)}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}.
\] Using the change of variables \(t = \tfrac{Lz}{\gamma}\),
\(z = \tfrac{\gamma\,t}{L}\), and \(dz = \tfrac{\gamma}{L}\,dt\), we
obtain \begin{align*}
I 
&= C^\lambda
   \int_{0}^{\infty}
     \Bigl(\tfrac{\gamma\,t}{L}\Bigr)^{\,\lambda(L - 1)}
     \Bigl(\gamma + L\,\tfrac{\gamma\,t}{L}\Bigr)^{-\lambda(L - \alpha)}
     \,\tfrac{\gamma}{L}\,dt
\\
&= C^\lambda 
   \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \int_{0}^{\infty}
     \frac{t^{\,\lambda(L - 1)}}
          {(1 + t)^{\,\lambda(L - \alpha)}}
   \,dt.
\end{align*} By the Beta-function identity \[
\int_{0}^{\infty} \frac{t^{\,a - 1}}{(1 + t)^{\,a + b}} \, dt 
= B(a,b),
\] where \[
a = \lambda(L - 1) + 1,
\quad
b = \lambda(-\alpha + 1) - 1,
\] it follows that \begin{align*}
I 
&= C^\lambda \,\frac{\gamma^{\,1+\lambda(\alpha - 1)}}{L^{\,1+\lambda(L - 1)}}
   \,B(a,b).
\end{align*} Next, we note that
\(\gamma^{\,1 + \lambda(\alpha - 1)} = \gamma^{\,1 - \lambda + \lambda\alpha}\)
and \(L^{\,1 + \lambda(L - 1)} = L^{\,\lambda L + 1 - \lambda}.\) Since
\[
C^\lambda 
= \biggl(\tfrac{L^L}{\gamma^\alpha\,\Gamma(-\alpha)\,\Gamma(L)}\,\Gamma(L - \alpha)\biggr)^{\!\lambda}
= L^{\lambda L}\,\gamma^{-\alpha \lambda}
  \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda,
\] we obtain \begin{align*}
I
&= \gamma^{\,1 - \lambda}\,
   L^{\,\lambda - 1}
   \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
   \,B(a,b).
\end{align*}

\medskip

\noindent By \eqref{eq:RenyiDefinition}, the Rényi entropy, is given by:
\begin{align*}
H_\lambda(Z)
&= \frac{1}{\,1 - \lambda\,} \,\ln I.
\end{align*} Hence, \begin{align*}
H_\lambda(Z) 
&= \frac{1}{\,1 - \lambda\,}
  \,\ln\!\Bigl[
    \gamma^{\,1 - \lambda}\,
    L^{\,\lambda - 1}\,
    \Bigl(\tfrac{\Gamma(L - \alpha)}{\Gamma(-\alpha)\,\Gamma(L)}\Bigr)^\lambda
    \,B(a,b)
  \Bigr].
\end{align*} Thus, for \(Z \sim \mathcal{G}^0_I(\alpha, \gamma, L)\),
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha, \gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{\,L}\Bigr)
  + \frac{1}{\,1 - \lambda\,}
    \Bigl[
      \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
      + \ln B(a,b)
    \Bigr].
\end{align*} Using the property \[
\ln B(a,b) 
= \ln \Gamma(a) + \ln \Gamma(b) - \ln \Gamma(a + b),
\] where \(a + b = \lambda(L - \alpha)\), we have \begin{align}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\gamma, L)\bigr)
&= \ln\Bigl(\tfrac{\gamma}{L}\Bigr)
 + \frac{1}{\,1 - \lambda\,}
   \Bigl[
     \lambda\bigl(\ln \Gamma(L - \alpha) 
            - \ln \Gamma(-\alpha) 
            - \ln \Gamma(L)\bigr)
     + \ln \Gamma(a)
     + \ln \Gamma(b)
     - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
   \Bigr].
\label{eq:GI0RenyiInGamma}
\end{align} Finally, noting that \[
\mu = -\tfrac{\gamma}{\alpha + 1}
\quad\Longrightarrow\quad
\gamma = -\mu(\alpha + 1),
\] and substituting \(\gamma\) into \eqref{eq:GI0RenyiInGamma}, we
obtain \begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I( \alpha,\mu, L)\bigr)
= \ln \mu  -  \ln L + \ln(- 1-\alpha)
+ \frac{1}{\,1 - \lambda\,}
  \Bigl[
    \lambda\Bigl(\ln \Gamma(L - \alpha) 
        - \ln \Gamma(-\alpha) 
        - \ln \Gamma(L)\Bigr)\\
    + \ln \Gamma\bigl(\lambda(L - 1) + 1\bigr)
    + \ln \Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
    - \ln \Gamma\bigl(\lambda(L - \alpha)\bigr)
  \Bigr],
\label{eq:RenyiGI0Final}
\end{multline} which completes the derivation.

\section{\texorpdfstring{Relation to the
\texorpdfstring{$\Gamma_{\mathrm{SAR}}$}{Gamma SAR}
Distribution}{Relation to the  Distribution}}\label{relation-to-the-distribution}

The Rényi entropy of the \(\mathcal{G}^0_I(\alpha,\mu,L)\) distribution
can be expressed in terms of the Rényi entropy of the
\(\Gamma_{\mathrm{SAR}}(L,\mu)\) distribution, plus additional terms
involving \(\alpha\) and the Gamma function. Specifically, we can write:
\begin{multline}
H_\lambda\bigl(\mathcal{G}^0_I(\alpha,\mu,L)\bigr)
= 
\underbrace{\Bigl[
  \ln\mu \;-\;\ln L 
  \;+\; \frac{1}{\,1-\lambda\,}\Bigl(
    -\lambda \,\ln\Gamma(L) 
    \;+\;\ln\Gamma\bigl(\lambda(L-1)+1\bigr)
    \;-\;\bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
  \Bigr)
\Bigr]}_{\displaystyle H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)}
\\[6pt]
+~\ln\bigl(-1-\alpha\bigr)
+~\frac{1}{\,1-\lambda\,} 
 \Bigl[
   \lambda\bigl(\ln\Gamma(L-\alpha) - \ln\Gamma(-\alpha)\bigr)
   \;+\;\ln\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
   \;-\;\ln\Gamma\bigl(\lambda(L-\alpha)\bigr)
   \;+\;\bigl(\lambda(L-1)+1\bigr)\,\ln(\lambda)
 \Bigr].
\label{eq:GI0_in_terms_of_GammaSAR}
\end{multline} From \eqref{eq:GI0_in_terms_of_GammaSAR}, the bracketed
expression on the first line matches
\(H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L,\mu)\bigr)\), while the
remaining terms account for the parameter \(\alpha\) through additional
Gamma functions and logarithmic corrections. This decomposition
highlights the close relationship between the Rényi entropies of the
\(\mathcal{G}^0_I\) and \(\Gamma_{\mathrm{SAR}}\) distributions.

\section{\texorpdfstring{Limit Behavior of
\texorpdfstring{$H_\lambda(\mathcal{G}^0_I)$}{Hl(GI0)} as
\texorpdfstring{$\alpha \to -\infty$}{alpha->-∞}}{Limit Behavior of  as }}\label{limit-behavior-of-as}

\begin{proof}
We want to show that 
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu, \alpha, L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu, L).
\]

We can express \eqref{eq:GI0_in_terms_of_GammaSAR} as follows:
\begin{align*}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
&=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L)
\;+\;
\ln\!\bigl(-1-\alpha\bigr)
\\
&\quad
+ \frac{1}{1-\lambda}
\ln \Biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\Biggr].
\end{align*}
Set
\begin{align*}
\Delta(\alpha)
&=
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
-
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\end{align*}
Then 
\begin{align}
\Delta(\alpha)
=
\ln(-1-\alpha)
+
\frac{1}{1-\lambda}
\ln \biggl[
  \frac{
    \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)\,\lambda^{\lambda(L-1)+1}
  }{
    \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
  }
\biggr].
\label{eq:remain}
\end{align}

As \(\alpha \to -\infty\), we have \(-1-\alpha \approx |\alpha|\), so 
\[
\ln(-1-\alpha) 
\sim 
\ln|\alpha|.
\]
Note that for large \(|\alpha|\), we can the asymptotic relation 
\(
\Gamma(x+a)/\Gamma(x+b) 
\sim 
x^{\,a-b}
\).
Specifically:
\[
\Gamma(L-\alpha)/\Gamma(-\alpha) \;\sim\; |\alpha|^L,
\quad
\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)/\Gamma\bigl(\lambda(L-\alpha)\bigr)
\;\sim\; 
\bigl(\lambda|\alpha|\bigr)^{\,(\lambda-1)-\lambda L}.
\]
Thus, inside the logarithm in \eqref{eq:remain},
\[
\frac{
  \Gamma(L-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(-\alpha+1)-1\bigr)
}{
  \Gamma(-\alpha)^{\lambda}\,\Gamma\bigl(\lambda(L-\alpha)\bigr)
}
\;\sim\;
|\alpha|^{\lambda L}
\times
|\alpha|^{(\lambda-1)-\lambda L}
=
|\alpha|^{\,\lambda-1}.
\]
Since \(\lambda^{\lambda(L-1)+1}\)  does not depend on \(\alpha\), multiplying by this constant factor does not alter the asymptotic behavior in \(\alpha\). Therefore,
\[
\frac{1}{1-\lambda}\,
\ln\!\Bigl[\dots\Bigr]
\;\sim\;
\frac{1}{1-\lambda}\;\ln\!\bigl(|\alpha|^{\,\lambda-1}\bigr)
=
\frac{\lambda-1}{1-\lambda}\,\ln|\alpha|
=
-\ln|\alpha|.
\]
Hence
\[
\Delta(\alpha)
\;\sim\;
\ln|\alpha|
-
\ln|\alpha|
=
0
\quad
\text{as}\, \alpha\to -\infty.
\]
This shows 
\(\Delta(\alpha)\to 0\), 
and consequently
\[
\lim_{\alpha \to -\infty}
H_\lambda\bigl(\mathcal{G}^0_I\bigr)(\mu,\alpha,L)
=
H_\lambda\bigl(\Gamma_{\mathrm{SAR}}\bigr)(\mu,L).
\]
\end{proof}

\section{Bias Reduction via Bootstrap for Entropy Estimation}

The non-parametric Rényi entropy estimator
\(\widehat{H}_\lambda(\bm{Z})\), as defined in \eqref{eq:est_R},
exhibits bias, meaning its expectation does not coincide with the true
entropy \(H_\lambda\): \begin{equation}
\operatorname{Bias}(\widehat{H}_\lambda) = {E}[\widehat{H}_\lambda] - H_\lambda \neq 0.
\end{equation}

To reduce this bias, we employ a bootstrap correction approach. Let
\(\bm{Z} = (Z_1, Z_2, \dots, Z_n)\) be a sample of size \(n\), and let
\(\bm{Z}^{(b)} = (Z_1^{(b)}, Z_2^{(b)}, \dots, Z_n^{(b)})\) for
\(b = 1, 2, \dots, B\) be a bootstrap sample obtained by resampling with
replacement from \(\bm{Z}\). For each bootstrap sample, we compute the
entropy estimator \(\widehat{H}_\lambda(\bm{Z}^{(b)})\). Since bootstrap
resampling mimics the sampling variability, we estimate the bias as:
\begin{equation}
\widehat{\operatorname{Bias}} = \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}) - \widehat{H}_\lambda(\bm{Z}).
\end{equation}

A bias-corrected estimator can then be defined as: \begin{align}
\widetilde{H}_\lambda &= \widehat{H}_\lambda - \widehat{\operatorname{Bias}},\\
&= 2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)}).
\end{align} This estimator is motivated by the well-known bias
correction principle: \begin{equation}
\text{Bias-corrected estimate} = \text{Original estimate} - \text{Bias estimate}.
\end{equation} The expectation of \(\widetilde{H}_\lambda\) is given by:
\begin{align}
\mathbb{E}[\widetilde{H}_\lambda] &= \mathbb{E}[2\widehat{H}_\lambda(\bm{Z}) - \frac{1}{B} \sum_{b=1}^{B} \widehat{H}_\lambda(\bm{Z}^{(b)})]\\
&= 2\mathbb{E}[\widehat{H}_\lambda(\bm{Z})] - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})]\\
&= \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] + \big( H_\lambda - \mathbb{E}[\widehat{H}_\lambda(\bm{Z})] \big)\\
&= H_\lambda.
\end{align} Thus, \(\widetilde{H}_\lambda\) is an \textbf{approximately
unbiased} estimator of \(H_\lambda\). The effectiveness of this
correction depends on the choice of \(B\) and the underlying sample
distribution, but it systematically reduces bias in entropy estimation,
particularly for small sample sizes.

For a continuous random variable \(Z\) with probability density function
(pdf) \(f(z)\), the Rényi entropy of order
\(\lambda \in \mathbb{R}_+ \setminus \{1\}\) is defined as:
\begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation} Using \eqref{E:entropy2}, we derive closed-form
expressions for the Rényi entropy of the \(\Gamma_{\mathrm{SAR}}\) and
\(\mathcal{G}^0_I\) distributions:

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr].
\end{multline}

Similarly, for the \(\mathcal{G}^0_I\) distribution, we obtain:

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\mathrm{SAR}}(L, \mu)\bigr) \\
+ \ln(-1 - \alpha) \\
+ \frac{1}{1 - \lambda} \Bigl[ 
    \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) \\
    + \ln\Gamma\bigl(\lambda(L - 1) + 1\bigr) 
\Bigr].
\end{multline}

The detailed derivation of the Rényi entropy for the \(\mathcal{G}^0_I\)
distribution is provided in the Appendix.

Figure\textasciitilde{}\ref{fig:entropy_plot} illustrates the behavior
of the Rényi entropy as a function of \(\lambda\) for different values
of the distribution parameters.

For a continuous random variable \(Z\) with pdf \(f(z)\), the Rényi
entropy of order \(\lambda \in \mathbbm R_+ \setminus \{1\}\) is defined
as: \begin{equation}
\label{E:entropy2}
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.
\end{equation}

Using \eqref{E:entropy2}, we derive closed-form expressions for the
Rényi entropy of the \(\Gamma_{\mathrm{SAR}}\) and \(\mathcal{G}^0_I\)
distributions. Importantly, the \(\mathcal{G}^0_I\) entropy can be
expressed in terms of the \(\Gamma_{\mathrm{SAR}}\) entropy plus
additional terms that account for the roughness parameter \(\alpha\):

\begin{multline}
\label{eq-HGammaSAR}
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr)
= 
\ln \mu - \ln L + \frac{1}{1-\lambda}
\Bigl[
  -\lambda\,\ln\Gamma(L)  \\ + \ln\Gamma\bigl(\lambda(L-1)+1\bigr)  - \bigl(\lambda(L-1)+1\bigr)\,\ln\lambda
\Bigr],
\end{multline}

\begin{multline}
\label{eq-HGI0}
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \Delta H_\lambda(\alpha) \\
\text{where } \Delta H_\lambda(\alpha) = \ln(-1 - \alpha) + \frac{1}{1 - \lambda}
\Bigl[
   \lambda\bigl(
      \ln\Gamma(L - \alpha)
      -\ln\Gamma(-\alpha)
   \bigr) \\
   +\ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr)
   -\ln\Gamma\bigl(\lambda(L - \alpha)\bigr)
   +\bigl(\lambda(L-1)+1\bigr)\ln\lambda
\Bigr].
\end{multline}

The term \(\Delta H_\lambda(\alpha)\) quantifies the additional entropy
contribution due to texture variations in the \(\mathcal{G}^0_I\) model.
As \(\alpha \to -\infty\), we can show that
\(\Delta H_\lambda(\alpha) \to 0\), recovering the
\(\Gamma_{\mathrm{SAR}}\) case. This limiting behavior confirms that the
\(\Gamma_{\mathrm{SAR}}\) distribution is indeed a special case of the
\(\mathcal{G}^0_I\) distribution for homogeneous areas.

\citet{fig-plot} presents the Rényi entropy of the \(\mathcal{G}^0_I\)
model as a function of \(\mu\) for different values of \(\alpha\),
demonstrating this convergence.

For a continuous random variable \(Z\) with pdf \(f(z)\), the Rényi
entropy of order \(\lambda \in \mathbbm{R}_+ \setminus \{1\}\) is
defined as:\\
\begin{equation}  
\label{E:entropy2}  
H_\lambda(Z) = \frac{1}{1 - \lambda} \ln \int_{-\infty}^{\infty} [f(z)]^\lambda \, dz.  
\end{equation}

Using \eqref{E:entropy2}, we derive the Rényi entropy for the
\(\Gamma_{\text{SAR}}\) distribution:\\
\begin{multline}  
\label{eq-HGammaSAR}  
H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) = \ln\mu - \ln L \\  
+ \frac{1}{1-\lambda} \Bigl[ -\lambda\ln\Gamma(L) + \ln\Gamma\bigl(\lambda(L-1)+1\bigr) - \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

For the \(\mathcal{G}^0_I\) distribution, the entropy \textbf{explicitly
extends} \(H_\lambda(\Gamma_{\text{SAR}})\) with terms depending on the
roughness parameter \(\alpha\):\\
\begin{multline}  
\label{eq-HGI0}  
H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr) + \ln(-1 - \alpha) \\  
+ \frac{1}{1 - \lambda} \Bigl[ \lambda\bigl(\ln\Gamma(L - \alpha) - \ln\Gamma(-\alpha)\bigr) + \ln\Gamma\bigl(\lambda(-\alpha + 1) - 1\bigr) \\  
- \ln\Gamma\bigl(\lambda(L - \alpha)\bigr) + \bigl(\lambda(L-1)+1\bigr)\ln\lambda \Bigr].  
\end{multline}

As \(\alpha \to -\infty\), the additional terms vanish, reducing
\eqref{eq-HGI0} to \eqref{eq-HGammaSAR}. This confirms that
\(\Gamma_{\text{SAR}}\) is the limiting case of \(\mathcal{G}^0_I\) for
homogeneous areas (fully developed speckle).

These results are proven in the Appendix. Notably, the
\(\mathcal{G}^0_I\) entropy converges to the \(\Gamma_{\text{SAR}}\)
case as \(\alpha \to -\infty\), i.e., \[
\lim_{\alpha \to -\infty} H_\lambda\bigl(\mathcal{G}^0_I(\mu, \alpha, L)\bigr) = H_\lambda\bigl(\Gamma_{\text{SAR}}(L, \mu)\bigr).
\] This limit is analytically derived in the Appendix and corroborates
that \(\Gamma_{\text{SAR}}\) describes homogeneous areas (fully
developed speckle), while \(\mathcal{G}^0_I\) generalizes it for
textured regions.

\section*{Appendix: Proof of the Limit Behavior}

The limit
\(\lim_{\alpha \to -\infty} H_\lambda(\mathcal{G}^0_I) = H_\lambda(\Gamma_{\text{SAR}})\)
follows from analyzing the additional terms in \eqref{eq-HGI0}. Let:\\
\[
\Delta(\alpha) = H_\lambda(\mathcal{G}^0_I) - H_\lambda(\Gamma_{\text{SAR}}) = \ln(-1-\alpha) + \frac{1}{1-\lambda} \ln\!\left[ \frac{\Gamma(L-\alpha)^\lambda \Gamma(\lambda(-\alpha+1)-1) \lambda^{\lambda(L-1)+1}}{\Gamma(-\alpha)^\lambda \Gamma(\lambda(L-\alpha))} \right].
\]\\
Using asymptotic properties of the Gamma function for large
\(|\alpha|\):\\
\[
\frac{\Gamma(L-\alpha)}{\Gamma(-\alpha)} \sim |\alpha|^L, \quad \frac{\Gamma(\lambda(-\alpha+1)-1)}{\Gamma(\lambda(L-\alpha))} \sim (\lambda|\alpha|)^{\lambda-1 - \lambda L},
\]\\
we obtain \(\Delta(\alpha) \sim \ln|\alpha| - \ln|\alpha| = 0\). Thus,
\(\Delta(\alpha) \to 0\) as \(\alpha \to -\infty\).

\newpage
\singlespacing 
\bibliography{../../Common/references.bib}

\end{document}