### 2025-06-02 Text/Thesis/Chapters/Chapter1.qmd

<!-- # --- -->
<!-- # format: -->
<!-- #   pdf: -->
<!-- #     documentclass: ../MastersDoctoralThesis -->
<!-- #     classoption: ["11pt", "english", "singlespacing", "headsepline"] -->
<!-- #     bibliography: ../example.bib -->
<!-- # --- -->

# INTRODUCTION {#sec-Chapter1}

The technology of Synthetic Aperture Radar (SAR) imaging operates on the comprehensive aperture principle to produce high-resolution SAR images.
In contrast to conventional optical remote sensing images, SAR images remain unaffected by external factors, making them suitable for identifying ground targets in diverse weather conditions and expansive areas.  
Consequently, SAR imagery has become essential for environmental monitoring, crop mapping, ship target detection, and climatology.  
However, the effective use of SAR
data depends on a thorough understanding of its statistical properties
because it is corrupted by speckle. This noise-like interference effect
is inherent in SAR data due to the coherent nature of the imaging
process.

Speckle in intensity format is non-Gaussian. 
Thus, SAR data require reliable statistical models for accurate processing. 
The $\mathcal{G}^0$ distribution, which is suitable for SAR data, includes
the Gamma law as the limiting case for fully-developed
speckle and provides flexibility with fewer
parameters for analysis.

When deciding which model is the best, practitioners face a problem. 
On the one hand, if they opt for the Gamma law when the data come from the
$\mathcal{G}^0$ distribution, they lose all the information about the
number of scatterers, which is revealed by one of the parameters of the
latter model. 
On the other hand, if they apply the $\mathcal{G}^0$ distribution under fully developed speckle, maximum
likelihood estimation is tricky: bias increases making estimation
unreliable, and the likelihood is
flat, so numerical optimization may not
converge.

Our work aims to improve the identification of potential roughness
features in SAR intensity data. 
Physical modeling of SAR data allows the
use of the Gamma distribution in the presence of fully-developed
speckle, where an infinite number of independent backscatterers per
resolution unit is assumed, commonly referred to as homogeneous regions.

Entropy is a fundamental concept in information theory with far-reaching
applications in pattern recognition, statistical physics, image
processing, edge detection and SAR image
analysis.
@Shannon1948 introduced this concept for a random variable to measure information and uncertainty. 
Shannon entropy is a crucial descriptive parameter in statistics, especially for evaluating data
dispersion and performing tests for normality, exponentiality and
uniformity [@Wieczorkowski1999; @Zamanzade2012].

Entropy estimation is challenging, especially when the model is unknown. In these cases,
non-parametric methods are used. Spacing methods have been discussed as
a non-parametric approach in Refs. [@AlizadehNoughabi2010;@Subhash2021]. This strategy is flexible
and robust because it does not enforce a model or parametric
constraints. In this context, we introduce a novel approach: a bootstrap-improved non-parametric estimator for Shannon entropy.

When applying these statistical concepts to SAR image analysis, we aim to discern between homogeneous and non-homogeneous regions, contributing to the improved interpretation and understanding of SAR data.  In our study, we devise test statistics based on these three estimators: the classical coefficient of variation, a robust version, and the Shannon entropy estimator. We apply these test statistics to generate maps of evidence of homogeneity that reveal different types of targets in the SAR data, assuming the number of looks is known. Likewise, we show that our proposed method is superior to existing approaches with simulated data and SAR images.

## OBJETIVES

The aim of this work is the identification of roughness features in SAR intensity data by developing novel test statistics for distinguishing between homogeneous and heterogeneous domains.  
To achieve this goal, we propose the following procedure:

- Select the optimal non-parametric estimator of entropy, refined through bootstrap techniques, aiming to reduce bias and mean square error.
- Propose three test statistics: the first based on the selected non-parametric estimator of entropy; the second utilizing the classical coefficient of variation; and the third using a robust approach of CV variant.
- Conduct computational experiments to evaluate the performance of the proposed test statistics across various simulated data scenarios.
- Apply the test statistics to SAR data to assess their effectiveness in identifying roughness features.

## CONTRIBUTIONS

The main contributions of this work are the following:

The main contributions of this work are the following:

1. Development of a new non-parametric entropy-based test statistic for roughness detection in SAR images.
2. Comparative study of three different test statistics under various simulated scenarios.
3. Integration of bootstrap techniques to improve estimation accuracy and robustness.
4. Application of the proposed methods to real SAR data, demonstrating their practical relevance.

## STRUCTURE OF THE THESIS

@sec-Chapter2 provides a background on remote sensing and SAR images. @sec-Chapter3 presents methodological aspects, including statistical modeling, entropy estimation for intensity SAR data, and hypothesis testing, and @sec-Chapter4 the statistical modelling choices common to all objectives of this thesis. @sec-Chapter5 presents the main results obtained in our study with both simulated and actual data. @sec-Chapter6 concludes with a discussion on heterogeneity implications of the findings and areas for future research building on the work in this thesis.