### 2025-06-02 Text/Thesis/Chapters/Chapter3.qmd

```{r setup, include=FALSE}
source("setup.R")
```
# PROBLEM STATEMENT AND METHODOLOGY {#sec-Chapter3}

## Bootstrap Correction for  Entropy Estimators

## Hypothesis Testing

## Proposed Test Statistics for Heterogeneity Detection