### 2025-06-02 Text/Thesis/Chapters/Chapter4.qmd

```{r setup, include=FALSE}
source("setup.R")
```
# RESULTS {#sec-Chapter4}

## Dataset Description

The following steps were applied to the SAR data during pre-processing:

1. Importing the SAR image into SNAP.  
2. Creating a subset to focus on the area of interest.  
3. Performing Geometric Correction, which includes Range Doppler Terrain Correction.  
4. Exporting the processed data in ENVI format for further analysis.

## Analysis with Simulated data

## Analysis with SAR data
We compare two test statistics for detecting heterogeneity in SAR data: (i) the Rényi entropy-based test, $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, described in&nbsp;\eqref{eq-test}; and (ii) the Shannon entropy-based test, $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$, which is based on the Al-Omari estimator proposed in&nbsp;[@IbrahimAlOmari2014] and was improved via bootstrap in our previous work. The details of this test can be found in Frery et al.&nbsp;[@Frery2024].

We applied these tests to images covering urban, agricultural, and water regions. 
We analyzed images from London, United Kingdom;  the surroundings of Munich, Germany; and  Dublin Port, Ireland, as shown in Figs.&nbsp;\ref{fig:london-a},&nbsp;\ref{fig:munich-a}, and&nbsp;\ref{fig:dublin-a}. All images were acquired in HH polarization.
Table&nbsp;\ref{tab:table_param} provides the detailed acquisition parameters of these images.

\renewcommand{\arraystretch}{3}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

SAR_data <- data.frame(
  Site = c("London", "Munich", "Dublin" ),
  Mission = c("TanDEM-X", "UAVSAR", "TanDEM-X"),
  Band = c("X", "L", "X"),
  Size = c("$2000\\times2000$", "$1024\\times1024$", "$1100\\times1100$"),
  L = c(1, 12, 16),
  Resol = c("$0.99/0.99$", "$4.9/7.2$", "$1.35/1.35$"),
  Date = c("12-11-2021", "16-04-2015", "03-09-2017")
)

colnames(SAR_data) <- c(
  "\\textbf{Site}", "\\textbf{Mission}", "\\textbf{Band}", 
  "\\textbf{Size (pixels)}", "$\\bm{L}$", 
  "\\textbf{Resolution [m]}", "\\textbf{Acquisition Date}"
)

SAR_data[] <- lapply(SAR_data, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.1f$", x), sprintf("$\\phantom{-}%.1f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

kable(SAR_data, 
      format = "latex",
      booktabs = TRUE,
      align = "ccccccc",
      escape = FALSE,
      digits = 2,
      label = "table_param",
      centering = TRUE,
      caption = "Parameters of selected SAR images.",
      table.envir = "table", position = "hbt") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 21) %>% 
  kable_styling(full_width = TRUE) %>%
  column_spec(1, width = "3.0cm") %>%
  column_spec(2, width = "3.7cm") %>%
  column_spec(3, width = "1.5cm") %>%
  column_spec(4, width = "4cm") %>%
  column_spec(5, width = "1.5cm") %>%
  column_spec(6, width = "4.0cm") %>%
  column_spec(7, width = "4.5cm")

```
We used sliding windows of size $7\times 7$.
The results are presented as $p$-value maps and binary maps at a $0.05$ significance level. 
Figs.&nbsp;\ref{fig:london-b}--\ref{fig:london-c},&nbsp;\ref{fig:munich-b}--\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-b}--\ref{fig:dublin-c} correspond to $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, while Figs.&nbsp;\ref{fig:london-d}--\ref{fig:london-e},&nbsp;\ref{fig:munich-d}--\ref{fig:munich-e}, and&nbsp;\ref{fig:dublin-d}--\ref{fig:dublin-e} correspond to $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$.

The $p$-value maps use a color gradient, where darker regions indicate areas with higher roughness, and lighter regions denote smoother, less textured surfaces.
The binary maps classify these results: white areas represent $p$-values greater than $0.05$, indicating no statistical evidence to reject the null hypothesis (homogeneous regions). 
In contrast, black areas represent regions with $p$-values below $0.05$, providing statistical evidence to reject the null hypothesis and indicating heterogeneity.

The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations than the Shannon-based test due to the flexibility provided by the parameter $\lambda$, as observed in the binary maps in Figs.&nbsp;\ref{fig:london-c},&nbsp;\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-c}.

```{=latex}
\begin{figure}[H]
    \centering
    % First row: main SAR image
        \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/london_optical_.png}
        \caption*{(a) Optical image}
        \label{fig:london-a}
    \end{minipage}
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/london_2000.png}
        \caption*{(b) SAR image}
        \label{fig:london-b}
    \end{minipage}
    \vspace{1em} % Space between rows
        % Second row: p-value maps
\begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_london_Shannon_c1.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:london-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_london_shannon.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:london-d}
    \end{minipage}
        \vspace{1em} % Space between rows

    % Third row: binary maps
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_london_renyi.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:london-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_london_renyi_L1_.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:london-f}
    \end{minipage}
    \caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:london}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.85\textwidth]{Figures/munich_optical1.png}
        \caption*{(a) Optical image}
        \label{fig:munich-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/munich_1024.png}
        \caption*{(b) SAR image}
        \label{fig:munich-b}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Second row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_muni_Shan22.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:munich-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_munich_sh_AO_L12.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:munich-d}
    \end{minipage}

\vspace{1em} % Space between rows

    % Third row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_muni_renyi.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:munich-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_munich_renyi.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:munich-f}
    \end{minipage}
    \caption{Detection of heterogeneous areas in a SAR image over Munich, Germany: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:munich}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/dublin.png}
        \caption*{(a) Optical image}
        \label{fig:dublin-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/dublin_1100_hh.png}
        \caption*{(b) SAR image}
        \label{fig:dublin-b}
    \end{minipage}
    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/AO_w7_L16_b100.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:dublin-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_AO_w7_L16_b100.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:dublin-d}
    \end{minipage}
    \vspace{0.9em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/dublin_renyi_09_w7_b100.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:dublin-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_dublin_renyi_09_w7_b100.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:dublin-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over Dublin Port, Ireland: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:dublin}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/foggia-italy-optical.png}
        \caption*{(a) Optical image}
        \label{fig:Foggia-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/foggia-italy_1900.png}
        \caption*{(b) SAR image}
        \label{fig:Foggia-b}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_italy_AO_1900-H.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:Foggia-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_AO_italy_1900.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:Foggia-d}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_Renyi_Italy_1900_H.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:Foggia-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_renyi_italy_1900_L1.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:Foggia-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over Foggia, Italy: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:Foggia}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/san-francisco-optical1.png}
        \caption*{(a) Optical image}
        \label{fig:san-francisco-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{Figures/SF_3300.png}
        \caption*{(b) SAR image}
        \label{fig:san-francisco-b}
    \end{minipage}
    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_SF_AO_3300_H.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-francisco-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_AO_SF_3300_L1_3.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-francisco-d}
    \end{minipage}

    \vspace{0.9em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.3\textwidth]{Figures/H_pvalue_renyi_SF_3300_H2_L1_3.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-francisco-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{Figures/H_005_renyi_SF_3300_L1_3.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-francisco-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over San Francisco, USA: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:san-francisco}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.47\textwidth}
        \centering
        \includegraphics[width=0.8\textwidth]{Figures/new-orleans-optical.png}
        \caption*{(a) Optical image}
        \label{fig:new-orleans-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.85\textwidth]{Figures/NewOrleans_1400.png}
        \caption*{(b) SAR image}
        \label{fig:new-orleans-b}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_NewOrleans_1400_AO-H.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:new-orleans-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_NewOrleans_1400_AO.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:new-orleans-d}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_NewOrleans_1400_renyi-H.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:new-orleans-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_NewOrleans_1400_renyi.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:new-orleans-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over New Orleans city, USA: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:new-orleans}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.47\textwidth}
        \centering
        \includegraphics[width=0.8\textwidth]{Figures/san-diego-optical.png}
        \caption*{(a) Optical image}
        \label{fig:san-diego-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.85\textwidth]{Figures/sandiego_1300.png}
        \caption*{(b) SAR image}
        \label{fig:san-diego-b}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_sandiego_1300_AO-A.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-diego-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_sandiego_1300_AO.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-diego-d}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_SanDiego_1300_renyi-A.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-diego-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\linewidth]{Figures/H_005_sandiego_1300_renyi.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-diego-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over San Diego city, USA: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:san-diego}
\end{figure}

```

```{=latex}
\begin{figure}[H]
    \centering
    % First row: SAR image
    \begin{minipage}{0.47\textwidth}
        \centering
        \includegraphics[width=0.85\textwidth]{Figures/rotterdam_optical.png}
        \caption*{(a) Optical image}
        \label{fig:san-diego-a}
    \end{minipage}
    \hfill
    \begin{minipage}{0.48\textwidth}
        \centering
        \includegraphics[width=0.85\textwidth]{Figures/rotterdam.png}
        \caption*{(b) SAR image}
        \label{fig:san-diego-b}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Third row: p-value and binary for Shannon AO
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_rotterdam1_1200_AO-B.png}
        \caption*{(c) $p$-value map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-diego-c}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.85\linewidth]{Figures/H_005_rotterdam_AO_L9_B100_1.png}
        \caption*{(d) Binary map for $S_{\widetilde{H}_{\text{AO}}}$}
        \label{fig:san-diego-d}
    \end{minipage}

    \vspace{1em} % Space between rows

    % Second row: p-value and binary for Renyi
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=1.2\textwidth]{Figures/H_pvalue_rotterdam1_renyi_1200-B.png}
        \caption*{(e) $p$-value map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-diego-e}
    \end{minipage}
    \hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.85\linewidth]{Figures/H_005_renyi_rotterdam1_B100_09.png}
        \caption*{(f) Binary map for $S_{\widetilde{H}_{\lambda}}$}
        \label{fig:san-diego-f}
    \end{minipage}

    \caption{Detection of heterogeneous areas in a SAR image over Rotterdam, Netherlands: comparison with the tests $S_{\widetilde{H}_{\text{AO}}}$ and $S_{\widetilde{H}_{\lambda}}$.}
    \label{fig:san-diego}
\end{figure}

```