### 2025-06-04 Text/Articles/RENYI_ENTROPY_SAR_2025/Submissions/R0/R0-SubmissionLetter.tex
R1 has been submitted

\documentclass[12pt]{letter} % Uses 10pt

\usepackage[utf8]{inputenc}
\usepackage{merriweather}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{url}

\topmargin=-1in    % Make letterhead start about 1 inch from top of page 
\textheight=9in  % text height can be bigger for a longer letter
\oddsidemargin=0pt % leftmargin is 1 inch
\textwidth=6.5in   % textwidth of 6.5in leaves 1 inch for right margin

\begin{document}

\signature{\includegraphics[width=5cm]{assinatura_blue-no_background}\\
Alejandro C.\ Frery\\
\mbox{Professor of Statistics and Data Science}\\
alejandro.frery@vuw.ac.nz}           % name for signature 
\longindentation=0pt                       % needed to get closing flush left
\let\raggedleft\raggedright                % needed to get date flush left

\begin{letter}{Dr.\ Ronny Hänsch\\
	Editor-in-Chief\\
	IEEE Geoscience and Remote Sensing Letters}
\date{1 April 2025}

\begin{flushleft}
Alejandro C.\ Frery
\end{flushleft}
\smallskip\hrule height 2pt
\begin{flushright}
\begin{tabular}{rl}
\multirow{3}{*}[20pt]{\includegraphics[height=3.8cm]{VUW-logo}}	\\
& \small School of Mathematics and Statistics\\
	& \small Victoria University of Wellington\\
	& \small  6140 Wellington, New Zealand
\end{tabular}
\end{flushright} 
\vfill % forces letterhead to top of page

\opening{Dear Dr.\ Ronny Hänsch,}

We want to inform that we submitted the manuscript GRSL-00722-2025 (``Quantifying Roughness in SAR Imagery with the Rényi Entropy'') to the MIGARS Special Stream because it is a significantly extended version of ``Identifying departures from the fully developed speckle hypothesis in intensity SAR data with non-parametric estimation of the entropy'' (DOI \url{10.1109/MIGARS61408.2024.10544448}), presented at MIGARS~2024 and already published in IEEExplore.

\closing{Yours sincerely,}

%\encl{Dois formulários 0008/SPT-UFAL;V.2}  				% Enclosures

\end{letter}

\end{document}