### 2025-07-01 Text/Thesis/Frontmatter/abstract.tex

%\begin{otherlanguage}{spanish}
%\begin{abstract}
%\phantomsection
%\addcontentsline{toc}{chapter}{Abstract}

This work  presents a statistical approach to detect heterogeneity 
in synthetic aperture radar (SAR) intensity data using entropy-based methods.
In homogeneous regions, where fully developed speckle is present, SAR returns 
are well modeled by the Gamma distribution. 
In contrast, heterogeneous areas exhibit more complex scattering behaviors, 
typically modeled by the $\mathcal{G}_I^0$ distribution. Distinguishing between 
these regimes is fundamental for accurate interpretation in remote sensing applications, 
yet classical parametric tests often fail to address this task effectively due to analytical 
and numerical limitations.
To overcome these challenges, we propose three test statistics for detecting heterogeneity 
in SAR imagery based on Shannon, Rényi, and Tsallis entropy measures. These tests rely on 
nonparametric entropy estimators constructed from sample spacings, avoiding explicit assumptions 
about the underlying distribution. To enhance their accuracy, particularly in small samples, 
we incorporate a bootstrap bias-correction procedure that improves the estimators’ stability, 
reduces bias, and lowers mean squared error.
The proposed tests are evaluated using Monte Carlo simulations, assessing size and power 
under varying sample sizes and numbers of looks.
Results demonstrate that the Rényi and Tsallis-based tests outperform the Shannon-based version by 
detecting more subtle texture variations while maintaining greater reliability in identifying 
truly homogeneous regions.
Finally, the methodology is applied to both simulated and real SAR data. We generate $p$-value maps using a 
sliding window analysis, allowing for visual and quantitative assessment of spatial heterogeneity. 
The Rényi-based test consistently identifies fine-scale roughness patterns, 
while the Tsallis-based test shows strength in reliably detecting homogeneous areas. 
Together, these entropy-based tools provide a robust, interpretable, and unsupervised 
framework for heterogeneity detection in SAR data.

\vspace{1em}
\par
\noindent \textbf{Keywords:} SAR; Heterogeneity; Entropy; Nonparametric Estimation; Bootstrap; Statistical Test

%\end{otherlanguage}