### 2025-07-01 Text/Thesis/Frontmatter/symbols.tex

\begin{symbols}{lll} % Include a list of Symbols (a three column table)

$\mathbb{R}_+$ & Positive real numbers \\
 $\Gamma(\cdot)$  & Gamma function \\
$\Gamma_{\mathrm{SAR}}(L, \mu)$ & Gamma SAR distribution with parameters $L$ and $\mu$ \\
$\mathcal{G}^0_I$ & $\mathcal{G}^0_I$ distribution for intensity \\
$\psi^{(0)}(\cdot)$ & Digamma function \\
$H(Z)$ & Shannon entropy of random variable $Z$ \\
$R_{\lambda}(Z)$ & Rényi entropy of order $\lambda$ \\
$T_{\lambda}(Z)$ & Tsallis entropy of order $\lambda$ \\
$\widehat{H}_{\text{AO}}(\bm{Z})$ & Nonparametric Shannon entropy estimator (Al-Omari) \\
$\widehat{R}_{\lambda}(\bm{Z})$ & Nonparametric Rényi entropy estimator via sample spacings \\
$\widehat{T}_{\lambda}(\bm{Z})$ & Nonparametric Tsallis entropy estimator via sample spacings \\
$\widetilde{H}_{\text{AO}}(\bm{Z})$ & Bootstrap-corrected Shannon entropy estimator \\
$\widetilde{R}_{\lambda}(\bm{Z})$ & Bootstrap-corrected Rényi entropy estimator \\
$\widetilde{T}_{\lambda}(\bm{Z})$ & Bootstrap-corrected Tsallis entropy estimator \\
$S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$ & Test statistic based on Shannon entropy \\
$S_{\widetilde{R}_{\lambda}}(\bm{Z}; L)$ & Test statistic based on Rényi entropy \\
$S_{\widetilde{T}_{\lambda}}(\bm{Z}; L)$ & Test statistic based on Tsallis entropy \\
$\widehat{\mu}$ & Sample mean of $\bm{Z}$ \\
$Z_{(i)}$ & $i$-th order statistic in sample $\bm{Z}$ \\
$m$ & Spacing parameter (window size) used in entropy estimation \\
$c_i$ & Boundary correction factor in spacing estimators \\
\addlinespace % Gap to separate the Roman symbols from the Greek

\end{symbols}