### 2025-07-05 Text/Thesis/Frontmatter/acknowledgements.tex
thesis update

Firstly, I would like to thank my supervisors, Prof. Alejandro Frery and Prof. Abraão do Nascimento, for their guidance, dedication, and support throughout these years. 
Working with you has been an incredibly valuable learning experience. 
I truly appreciate the attention and dedication you give to your students.

I am also grateful to the Universidade Federal de Pernambuco and the 
Graduate Program in Statistics for their support during my academic journey. 
Special thanks to Prof. Pablo Rodriguez for all his help with administrative matters and for his patience.

I gratefully acknowledge the financial support provided by the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior – Brasil (CAPES) – Finance Code 001, 
the Programa de Doutorado-Sanduíche no Exterior (PDSE), and the Fundação de Amparo à Ciência e Tecnologia do Estado de Pernambuco (FACEPE).

I also thank the examination committee for their availability.

To my family, thank you for always supporting and encouraging me along the way.

Finally, thanks to all the friends I met during this time. Your support and friendship 
made this experience better.