### 2025-10-27 Text/Thesis/Chapters/Chapter4.qmd

```{r setup, include=FALSE}
source("setup.R")
```
# RESULTS {#sec-Chapter4}

## ANALYSIS WITH SIMULATED DATA {#sec-sim2}

Figure&nbsp;\ref{fig:sim_Phantom_1} presents the structural design of the synthetic phantom with dimensions of
$500\times500$ pixels, proposed by  Gomez et al.&nbsp;[-@Gomez2017] as a tool to assess the performance of speckle-reduction filters. This layout serves as a base template to generate the simulated images shown in Figures&nbsp;\ref{fig:sim_Phantom_2}–\ref{fig:sim_Phantom_3}.
Each region is filled with data simulated from the $\mathcal{G}^0_I$ distribution&nbsp;\eqref{E:gi02}, using different combinations of the roughness parameter $\alpha$ and the mean $\mu$, as indicated by the labels in each quadrant of the image. Light regions correspond to textured observations (heterogeneous), while darker regions represent textureless areas (homogeneous).
```{=latex}
\begin{figure}[H]
  \centering
  \caption{Synthetic dataset.}
  \label{fig:sim_Phantom}
  \hspace{2em}
  \begin{subfigure}[b]{0.34\textwidth}
    \centering
    \caption{Phantom.}
    \label{fig:sim_Phantom_1}
    \includegraphics[width=\linewidth]{Figures/Phantom1}
    \end{subfigure}
  \vspace{1.2em} 
  %\hspace{0.48em}
  \begin{subfigure}[b]{0.43\textwidth}
    \centering
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=5$.}
    \label{fig:sim_Phantom_2}
    \includegraphics[width=\linewidth]{Figures/Phantom_labels}
      \end{subfigure}
  \hfill
  %\hspace{0.48em}
  \begin{subfigure}[b]{0.43\textwidth}
    \centering
    \caption{Simulated image, varying $\alpha$ and $\mu$, with $L=9$.}
    \label{fig:sim_Phantom_3}
    \includegraphics[width=\linewidth]{Figures/Phantom4_L9}
      \end{subfigure}
  %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```
The first simulated image uses $L = 5$, and the second uses $L = 9$, to evaluate how the test statistics behave as the number of looks increases and speckle noise is reduced.

The $\alpha$ parameter of the $\mathcal{G}_I^0$ distribution is essential for
interpreting texture characteristics. Values near zero greater than
$-3$ suggest extremely textured targets, such as urban
zones&nbsp;[@Frery2019a]. As the value decreases, it indicates regions
with moderate texture (in the $\left[-6,-3\right]$ region), related to
forest zones, while values below $-6$ correspond to textureless
regions, such as pasture, agricultural fields, and water
bodies&nbsp;[@Neto2023].

We evaluated the three test statistics $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$ on simulated images using fixed sliding windows of size $7\times7$ and, for Tsallis only, the adaptive windowing strategy with side lengths from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$) and smoothing parameter $\eta=3$. Figures&nbsp;\ref{fig:sim1t} and \ref{fig:sim2t} present, respectively for $L=5$ and $L=9$, the $p$-value maps (top row) and the corresponding binary decision maps (bottom row).

The $p$-value maps (Figures&nbsp;\ref{fig:sim1a}–\ref{fig:sim1d} and \ref{fig:sim2a}–\ref{fig:sim2d}) are shown with a continuous color scale: dark red ($p\!\approx\!0$) indicates strong evidence of heterogeneity, while dark blue ($p\!\approx\!1$) corresponds to homogeneous areas. Intermediate colors follow a smooth progression orange ($p\!\approx\!0.2$–$0.4$) for transitional textures and yellow–green to cyan ($p\!\approx\!0.4$–$0.8$) for increasingly near-homogeneous regions.

The binary maps (Figures&nbsp;\ref{fig:sim1e}–\ref{fig:sim1h} and \ref{fig:sim2e}–\ref{fig:sim2h}) simplify these results using a threshold of $0.05$. Pixels with $p$-values below this threshold are shown in black, indicating statistically significant deviations and, therefore, heterogeneity. In contrast, pixels with $p \geq 0.05$ are displayed in white, showing no evidence to reject the null hypothesis of homogeneity.

When the number of looks increases to $L=9$ (Figure&nbsp;\ref{fig:sim2t}), the maps become cleaner due to reduced speckle. All tests successfully detect areas with different texture levels, although some differences remain. The Rényi-based statistic $S_{\widetilde{R}_{\lambda}}$ is more sensitive to texture variations, highlighting fine patterns of heterogeneity, while the Tsallis-based $S_{\widetilde{T}_{\lambda}}$ produces cleaner results in uniform areas by reducing false detections. This behavior reflects their formulations: Rényi entropy emphasizes variability, whereas Tsallis entropy is more robust to noise, helping to preserve homogeneous regions.

The adaptive Tsallis version further improves visual quality, larger windows in homogeneous areas yield smoother estimates, and smaller windows in heterogeneous regions enhance detail preservation.

```{=latex}
\begingroup
\setlength{\textfloatsep}{6pt} % menos espacio entre texto y figuras
\setlength{\floatsep}{6pt} % menos espacio entre figuras
\setlength{\intextsep}{6pt}

\begin{figure}[hbtp]
    \centering
 \caption{Results on simulated data ($L = 5$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. (a-c) $p$-value maps with a fixed $7\times7$ window.
(d) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. (e-h) Binary decision maps at 0.05 significance level.}
    \label{fig:sim1t}
    % Primera fila
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim1a}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1b}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-renyi.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1c}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_4_z_b100_w7-tsallis.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\tiny{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim1d}
        \includegraphics[width=\linewidth]{Figures/p-values_L5_Tsallis_adaptive.png}
            \end{subfigure}
    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim1e}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_b200_w7-AO.png}
            \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1f}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.9_b200_w7-renyi.png}
    \end{subfigure}
     \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim1g}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom_4_z_0.85_b200_w7-tsallis.png}
        \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary~map~$\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim1h}
        \includegraphics[width=\linewidth]{Figures/H_005_L5_Tsallis_adaptive_085.png}
    \end{subfigure}
   %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
\vspace{-0.25\baselineskip} 
\begin{figure}[hbtp]
    \centering
 \caption{Results on simulated data ($L = 9$) using $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$. (a-c) $p$-value maps with a fixed $7\times7$ window.
(d) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. (e-h) Binary decision maps at 0.05 significance level.}
    \label{fig:sim2t}
    % Primera fila
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim2a}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_1_8_b100_w7-AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2b}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_renyi_7x7.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2c}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_tsallis_7x7.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\tiny{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim2d}
        \includegraphics[width=\linewidth]{Figures/p-values_L9_SIM_tsallis_adaptative.png}
            \end{subfigure}
    \vspace{1em} % Espacio entre filas

    % Segunda fila
    \hspace{-0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:sim2e}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-AO.png}
            \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2f}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_b100_w7-09-renyi.png}
    \end{subfigure}
     \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary map  $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:sim2g}
        \includegraphics[width=\linewidth]{Figures/0.05_Phantom4_L9_1_8_0.85_b100_w7-tsallis.png}
        \end{subfigure}
    \hspace{0.5em}
    \begin{subfigure}{0.23\textwidth}
    \caption{Binary~map~$\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:sim2h}
        \includegraphics[width=\linewidth]{Figures/H_005_L9_tsallis_adaptative.png}
    \end{subfigure}
   %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
\endgroup
```

\newpage

## APPLICATIONS TO SAR DATA {#sec-sim3}

### SAR Dataset and Preprocessing {#sec:SAR-data}

We evaluated the proposed test statistics using a set of five SAR images acquired from different radar missions. These images cover diverse locations: the surroundings of London (United Kingdom), the outskirts of Munich, the city of New Orleans (USA), Dublin Port (Ireland), and the coast of Jalisco (Mexico), as shown in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The selected scenes include a variety of land types, such as urban areas, mountainous regions, agricultural zones, and water bodies. This diversity allows us to assess the performance of the proposed statistical tests under different texture conditions.

To facilitate interpretation, we also included the corresponding optical images (Figures&nbsp;\ref{fig:london-optical}–\ref{fig:ja-optical}) obtained from Sentinel-2 Level-2A products. These optical scenes were acquired on dates close to those of the SAR acquisitions to ensure visual consistency. The optical data were downloaded from the Copernicus Data Space platform\footnote{\url{https://dataspace.copernicus.eu/}}.

The SAR images come from TanDEM-X (X-band), UAVSAR (L-band), and Sentinel-1B (C-band). The two UAVSAR images were obtained from the Alaska Satellite Facility (ASF)\footnote{\url{https://search.asf.alaska.edu/}}, which provides airborne and satellite SAR data. The X-band images were downloaded from ESA’s Third Party Missions (ESA TPM)\footnote{\url{https://tpm-ds.eo.esa.int/collections/}}, and the C-band image from Copernicus Data Space. All images use single polarization (HH or VV), and their acquisition settings differ in resolution, scene size, and number of looks. Table&nbsp;\ref{tab:table_param} lists the main characteristics of each SAR image.

\renewcommand{\arraystretch}{2}   
```{r parameters_sar, echo=FALSE, message=FALSE, warning=FALSE}

SAR_data <- data.frame(
  Site = c("London",  "Munich","New Orleans", "Dublin","Coast of Jalisco" ),
  Mission = c("TanDEM-X", "UAVSAR", "UAVSAR", "TanDEM-X", "Sentinel-1B"),
  Band = c("X", "L","L", "X", "C"),
  Polarization = c("HH", "HH","HH","HH", "VV"),
  Size = c("$2000\\times2000$",  "$1024\\times1024$","$1400\\times1400$", "$1100\\times1100$","$512\\times512$"),
  L = c(1,  12, 12,  16,18),
  Resol = c("$0.99/0.99$",  "$4.9/7.2$","$4.9/7.2$", "$1.35/1.35$","$40/40$"),
  Date = c("12-11-2021", "16-04-2015","13-03-2016", "03-09-2017","29-08-2021")
)

colnames(SAR_data) <- c(
  "\\textbf{Site}", "\\textbf{Mission}", "\\textbf{Band}", "\\textbf{Polarization}",
  "\\textbf{Size (pixels)}", "$\\bm{L}$", 
  "\\textbf{Resolution [m]}", "\\textbf{Acquisition Date}"
)

SAR_data[] <- lapply(SAR_data, function(x) {
  if (is.numeric(x)) {
    if (all(x %% 1 == 0)) {
      formatted_numbers <- sprintf("$%d$", x)
    } else {
      formatted_numbers <- ifelse(x < 0, sprintf("$%.1f$", x), sprintf("$\\phantom{-}%.1f$", x))
    }
    return(formatted_numbers)
  } else {
    return(x)
  }
})

kable(SAR_data, 
      format = "latex",
      booktabs = TRUE,
      align = "cccccccc",
      escape = FALSE,
      digits = 2,
      label = "table_param",
      centering = TRUE,
      caption = "Parameters of selected SAR images.",
      table.envir = "table", position = "hbt") %>%
  row_spec(0, align = "c") %>%
  kable_styling(latex_options = "scale_down", font_size = 21) %>% 
  kable_styling(full_width = TRUE) %>%
  column_spec(1, width = "3.8cm") %>%
  column_spec(2, width = "3.7cm") %>%
  column_spec(3, width = "2.5cm") %>%
  column_spec(4, width = "3.5cm") %>%
  column_spec(5, width = "4cm") %>%
  column_spec(6, width = "2.5cm") %>%
  column_spec(7, width = "4.0cm") %>%
  column_spec(8, width = "4.5cm")

```

For each SAR image, the number of looks $L$ was obtained from the
product of azimuth and range looks provided in the metadata of SNAP. We
validated this nominal value with the equivalent number of looks (ENL)
from manually selected homogeneous areas using the classical formula
$\text{ENL} = 1/\widehat{\text{CV}}^2$, based on the sample
coefficient of variation. The results were consistent with the metadata.
We also verified that moderate deviations in $L$ did not significantly
impact the outcome of the test, confirming the robustness of the
proposed method.

The SAR images were preprocessed using the SNAP (Sentinel Application Platform) toolbox. The standard steps included terrain correction and multilooking (when necessary) to produce intensity images. UAVSAR scenes were also reprojected to align them with standard geographic coordinates. Finally, the resulting images were exported in ENVI format and used as input for statistical analysis in \texttt{R}, where the proposed tests were applied.

### Qualitative Inspection {#qualitative-inspection}

We applied the three bootstrap statistics, $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$, and $S_{\widetilde{T}_{\lambda}}$, to the five SAR intensity images in Figures&nbsp;\ref{fig:london-sar}–\ref{fig:ja-sar}. The fixed approach uses local sliding windows of size $7\times7$. The adaptive approach (used with Tsallis) varies the window side from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$), with smoothing parameter $\eta=3$, and $B=100$ bootstrap replications per window. Each figure is organized in three rows: (i) SAR scene and a matching Sentinel-2 optical image, (ii) $p$-value maps for the three tests, and (iii) binary maps at the $5\%$ level (black = heterogeneous, white = homogeneous).

Across all scenes, the three tests detect the main textured elements (urban blocks, port facilities, roads, crop boundaries, mountains, and river banks). In the $p$-value maps, dark red ($p\!\approx\!0$) concentrates on heterogeneous areas such as urban blocks, harbor structures, and field edges, while blue ($p\!\approx\!1$) dominates open water and other uniform surfaces. Intermediate colors mark transitional textures.

The adaptive Tsallis results are visually better than the fixed windows: in homogeneous regions, the maps look smoother and more stable, while in heterogeneous regions the edges are sharper and fine details are preserved. The binary maps make this clear: the adaptive strategy keeps thin structures (roads, levees, dock lines) that are partly lost with the fixed window.

The number of looks also matters. With fewer looks (stronger speckle), the adaptive scheme improves urban and boundary detection compared with the fixed window. As $L$ increases, speckle decreases and all methods improve, yet the adaptive maps still show cleaner homogeneous areas and crisper boundaries in textured zones.

Comparing statistics, the Rényi-based $S_{\widetilde{R}_{\lambda}}$ is more sensitive to texture changes and often highlights fine-scale heterogeneity. The Tsallis-based $S_{\widetilde{T}_{\lambda}}$ is more conservative in uniform areas, reducing false alarms and producing cleaner homogeneous regions. The adaptive Tsallis variant combines these benefits with data-driven window sizes, which explains its overall better visual quality across the five scenes.

<!-- The three statistical tests are applied to the SAR images using -->
<!-- $7\times 7$ local sliding windows. This window size balances detailed -->
<!-- local information and sufficient number of data points. -->

<!-- The $p$-values obtained for each test are presented in -->
<!-- Figures~\ref{fig:Rotterdam_pvalue}, \ref{fig:Mexico_pvalue}, -->
<!-- \ref{fig:lake_pvalue} and~\ref{fig:Illinois_crops_pvalue}, respectively. -->

<!-- In Figures~\ref{fig:Rotterdam_0.05}, \ref{fig:Mexico_crops_0.05}, -->
<!-- \ref{fig:lake_0.05} and~\ref{fig:Illinois_crops_0.05}, the maps of -->
<!-- $p$-values composed of a linear gradient of black and white colors, -->
<!-- represent the decisions at a \SI{5}{\percent} significance level. Dark -->
<!-- areas represent values below $0.05$, indicating evidence to reject the -->
<!-- null hypothesis and suggesting heterogeneity in these regions. In -->
<!-- contrast, values above 0.05 are represented as white areas, indicating -->
<!-- no evidence to reject the fully-developed speckle hypothesis. -->

<!-- We used sliding windows of size $7\times 7$. -->
<!-- The results are presented as $p$-value maps and binary maps at a $0.05$ significance level.  -->
<!-- Figs.&nbsp;\ref{fig:london-b}--\ref{fig:london-c},&nbsp;\ref{fig:munich-b}--\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-b}--\ref{fig:dublin-c} correspond to $S_{\widetilde{H}_{\lambda}}(\bm{Z}; L)$, while Figs.&nbsp;\ref{fig:london-d}--\ref{fig:london-e},&nbsp;\ref{fig:munich-d}--\ref{fig:munich-e}, and&nbsp;\ref{fig:dublin-d}--\ref{fig:dublin-e} correspond to $S_{\widetilde{H}_{\text{AO}}}(\bm{Z}; L)$. -->

<!-- The $p$-value maps use a color gradient, where darker regions indicate areas with higher roughness, and lighter regions denote smoother, less textured surfaces. -->
<!-- The binary maps classify these results: white areas represent $p$-values greater than $0.05$, indicating no statistical evidence to reject the null hypothesis (homogeneous regions).  -->
<!-- In contrast, black areas represent regions with $p$-values below $0.05$, providing statistical evidence to reject the null hypothesis and indicating heterogeneity. -->

<!-- The Rényi entropy-based test demonstrated greater sensitivity in detecting textural variations than the Shannon-based test due to the flexibility provided by the parameter $\lambda$, as observed in the binary maps in Figs.&nbsp;\ref{fig:london-c},&nbsp;\ref{fig:munich-c}, and&nbsp;\ref{fig:dublin-c}. -->

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over London, UK: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:london}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:london-sar}
        \includegraphics[width=\linewidth]{Figures/London-sar.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.28\textwidth}
            \caption{Optical image}
        \label{fig:london-optical}
        \includegraphics[width=\linewidth]{Figures/london-optical.jpg}
    \end{subfigure}

    \vspace{1.8em}

    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:london-shann1}
        \includegraphics[width=\linewidth]{Figures/london_AO.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:london-renyi1}
        \includegraphics[width=\linewidth]{Figures/London_renyi.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:london-tsallis1}
        \includegraphics[width=\linewidth]{Figures/london_tsallis.jpg}
    \end{subfigure}
\hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:london-tsallis1_ada}
        \includegraphics[width=\linewidth]{Figures/london_tsalis_adapt_eta3.jpg}
    \end{subfigure}
    \vspace{1.8em}

    % Tercera fila: 3 imágenes (binary maps)
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:london-005-shann2}
        \includegraphics[width=\linewidth]{Figures/H_005_london_shannon.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:london-005-renyi2}
        \includegraphics[width=\linewidth]{Figures/H_005_london_renyi_L1_.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:london-005-tsallis2}
        \includegraphics[width=\linewidth]{Figures/H_005_london1_TSALLIS.png} %
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:london-005-tsallis2_Ada}
        \includegraphics[width=\linewidth]{Figures/H_005_intensity_L1_eta_3.00_5x13_B20_lambda_12.png} %
    \end{subfigure}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over the outskirts
of Munich, Germany: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:mu}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:mun-sar}
        \includegraphics[width=\linewidth]{Figures/munich_sar.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.262\textwidth}
            \caption{Optical image}
        \label{fig:mun-optical}
        \includegraphics[width=\linewidth]{Figures/munich_optical1.jpg}
    \end{subfigure}

    \vspace{1.8em}

    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:mun-shann1}
        \includegraphics[width=\linewidth]{Figures/Munich_AO.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:mun-renyi1}
        \includegraphics[width=\linewidth]{Figures/Munich_renyi.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:mun-tsallis1}
        \includegraphics[width=\linewidth]{Figures/Munich_tsallis.jpg}
    \end{subfigure}
\hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:mun-tsallis1_ada}
        \includegraphics[width=\linewidth]{Figures/Munich_tsallis_eta3.jpg}
    \end{subfigure}
    \vspace{1.8em}

    % Tercera fila: 3 imágenes (binary maps)
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:mun-005-shann2}
        \includegraphics[width=\linewidth]{Figures/H_005_Munich_AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:mun-005-renyi2}
        \includegraphics[width=\linewidth]{Figures/H_005_Munich_renyi.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:mun-005-tsallis2}
        \includegraphics[width=\linewidth]{Figures/H_005_Munich_tsallis.png} %
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:mun-005-tsallis2_Ada}
        \includegraphics[width=\linewidth]{Figures/H_005_Munich_tsallis_eta3.png} %
    \end{subfigure}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over New Orleans city, USA: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:NO}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:no-sar}
        \includegraphics[width=\linewidth]{Figures/New-orleans1.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.245\textwidth}
            \caption{Optical image}
        \label{fig:no-optical}
        \includegraphics[width=\linewidth]{Figures/new-orleans-optical1.jpg}
    \end{subfigure}

    \vspace{1.8em}

    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:no-shann1}
        \includegraphics[width=\linewidth]{Figures/new_orleans_AO.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:no-renyi1}
        \includegraphics[width=\linewidth]{Figures/new_orleans_renyi.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:no-tsallis1}
        \includegraphics[width=\linewidth]{Figures/new_orleans_tsallis.jpg}
    \end{subfigure}
\hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:no-tsallis1_ada}
        \includegraphics[width=\linewidth]{Figures/new_orleans_tsallis_eta3.jpg}
    \end{subfigure}
    \vspace{1.8em}

    % Tercera fila: 3 imágenes (binary maps)
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:no-005-shann2}
        \includegraphics[width=\linewidth]{Figures/H_005_new_orleans_AO2.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:no-005-renyi2}
        \includegraphics[width=\linewidth]{Figures/H_005_new_orleans_renyi1.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:no-005-tsallis2}
        \includegraphics[width=\linewidth]{Figures/H_005_new_orl_tsallis1.png} %
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:no-005-tsallis2_Ada}
        \includegraphics[width=\linewidth]{Figures/H_005_new_orl_tsallis_eta3.png} %
    \end{subfigure}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over Dublin Port, Ireland: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:du}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:no-sar}
        \includegraphics[width=\linewidth]{Figures/dublin_sar_1100.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.28\textwidth}
            \caption{Optical image}
        \label{fig:du-optical}
        \includegraphics[width=\linewidth]{Figures/dublin_1100.jpg}
    \end{subfigure}

    \vspace{1.8em}

    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:no-shann1}
        \includegraphics[width=\linewidth]{Figures/Dublin_AO_w7.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:du-renyi1}
        \includegraphics[width=\linewidth]{Figures/Dublin_Renyi_w7.jpg}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:du-tsallis1}
        \includegraphics[width=\linewidth]{Figures/dublin_tsallis_w7.jpg}
    \end{subfigure}
\hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:du-tsallis1_ada}
        \includegraphics[width=\linewidth]{Figures/dublin_tsallis_adap.jpg}
    \end{subfigure}
    \vspace{1.8em}

    % Tercera fila: 3 imágenes (binary maps)
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:du-005-shann2}
        \includegraphics[width=\linewidth]{Figures/H_005_dublin_AO_1100.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:du-005-renyi2}
        \includegraphics[width=\linewidth]{Figures/H_005_dublin_renyi_1100.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:du-005-tsallis2}
        \includegraphics[width=\linewidth]{Figures/H_005_L16_envi_dublin_size_1100_tsallis_lambda085_B100.png} %
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:du-005-tsallis2_Ada}
        \includegraphics[width=\linewidth]{Figures/H_005_Dublin_tsallis_eta3.png} %
    \end{subfigure}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```

```{=latex}
\begin{figure}[H]
    \centering
\caption{Detection of heterogeneous areas in a SAR image over the Coast of Jalisco, Mexico: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window.
(f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for
$\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.}
    \label{fig:ja-sar}
    % Primera fila: 2 imágenes
    \begin{subfigure}{0.28\textwidth}
            \caption{SAR image}
        \label{fig:ja-sar}
        \includegraphics[width=\linewidth]{Figures/mexico_sar.jpg}
    \end{subfigure}
    \hspace{2em}
    \begin{subfigure}{0.27\textwidth}
    \caption{Optical image}
        \label{fig:ja-optical}
        \includegraphics[width=\linewidth]{Figures/mexico3_512.jpg}
    \end{subfigure}

    \vspace{1.8em}

    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:ja-shann1}
        \includegraphics[width=\linewidth]{Figures/p-values_Mexico_AO.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:ja-renyi1}
        \includegraphics[width=\linewidth]{Figures/p-values_Mexico_renyi.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.24\textwidth}
            \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:ja-tsallis1}
        \includegraphics[width=\linewidth]{Figures/p-values_Mexico_tsallis.png}
    \end{subfigure}
        \hfill
    \begin{subfigure}{0.24\textwidth}
    \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
            \label{fig:ja-tsallis1_ada}
        \includegraphics[width=\linewidth]{Figures/p-values_Mexico_tsallis_adapt_.png}
    \end{subfigure}

    \vspace{1.8em}

    % Tercera fila: 3 imágenes (binary maps)
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)}
        \label{fig:ja-005-shann2}
        \includegraphics[width=\linewidth]{Figures/0.05_results_Mexico_512_7_AO_300b.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)}
        \label{fig:ja-005-renyi2}
        \includegraphics[width=\linewidth]{Figures/0.05_L16_envi_mexico_512_renyi_lambda09_B200.png}
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)}
        \label{fig:ja-005-tsallis2}
        \includegraphics[width=\linewidth]{Figures/H_005_L18_envi_mexico_tsallis_lambda085_B200_w7_.png} %
    \end{subfigure}
    \hfill
    \begin{subfigure}{0.23\textwidth}
            \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)}
        \label{fig:ja-005-tsallis2_ada}
        \includegraphics[width=\linewidth]{Figures/H_005_Mexico_tsallis_Adtap.png} %
    \end{subfigure}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```

<!-- ```{=latex} -->
<!-- \begin{figure}[H] -->
<!--     \centering -->
<!-- \caption{Detection of heterogeneous areas in a SAR image over  Foggia, Italy: comparison of test statistics $\small{S_{\widetilde{H}_{\text{AO}}}}$, $\small{S_{\widetilde{R}_{\lambda}}}$, and $\small{S_{\widetilde{T}_{\lambda}}}$. Top: (a-b) input images (SAR and optical). Middle: (c-e) $p$-value maps with a fixed $7\times7$ window. -->
<!-- (f) $p$-value map with adaptive windows ($W_{\min}=5$, $W_{\max}=11$) for -->
<!-- $\eta=3.0$. Bottom:  (g-j) Binary decision maps at 0.05 significance level.} -->
<!--     \label{fig:fo} -->
<!--     % Primera fila: 2 imágenes -->
<!--     \begin{subfigure}{0.28\textwidth} -->
<!--             \caption{SAR image} -->
<!--         \label{fig:fo-sar} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia-sar1.jpg} -->
<!--     \end{subfigure} -->
<!--     \hspace{2em} -->
<!--     \begin{subfigure}{0.276\textwidth} -->
<!--             \caption{Optical image} -->
<!--         \label{fig:fo-optical} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia_optical1.jpg} -->
<!--     \end{subfigure} -->

<!--     \vspace{1.8em} -->

<!--     \begin{subfigure}{0.24\textwidth} -->
<!--             \caption{$p$-value map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-shann1} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia_AO.jpg} -->
<!--     \end{subfigure} -->
<!--     \hfill -->
<!--     \begin{subfigure}{0.24\textwidth} -->
<!--             \caption{$p$-value map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-renyi1} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia_renyi.jpg} -->
<!--     \end{subfigure} -->
<!--     \hfill -->
<!--     \begin{subfigure}{0.24\textwidth} -->
<!--             \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-tsallis1} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia_w7.jpg} -->
<!--     \end{subfigure} -->
<!-- \hfill -->
<!--     \begin{subfigure}{0.24\textwidth} -->
<!--             \caption{$p$-value map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)} -->
<!--         \label{fig:fo-tsallis1_ada} -->
<!--         \includegraphics[width=\linewidth]{Figures/Foggia_eta3.jpg} -->
<!--     \end{subfigure} -->
<!--     \vspace{1.8em} -->

<!--     % Tercera fila: 3 imágenes (binary maps) -->
<!--     \begin{subfigure}{0.23\textwidth} -->
<!--             \caption{Binary map $\small{S_{\widetilde{H}_{\text{AO}}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-005-shann2} -->
<!--         \includegraphics[width=\linewidth]{Figures/H_005_AO_italy_1900.png} -->
<!--     \end{subfigure} -->
<!--     \hfill -->
<!--     \begin{subfigure}{0.23\textwidth} -->
<!--             \caption{Binary map $\small{S_{\widetilde{R}_{\lambda}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-005-renyi2} -->
<!--         \includegraphics[width=\linewidth]{Figures/H_005_renyi_italy_1900_L1.png} -->
<!--     \end{subfigure} -->
<!--     \hfill -->
<!--     \begin{subfigure}{0.23\textwidth} -->
<!--             \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ ($7\times7$)} -->
<!--         \label{fig:fo-005-tsallis2} -->
<!--         \includegraphics[width=\linewidth]{Figures/H_005_fogia_7x7.png}  -->
<!--     \end{subfigure} -->
<!--     \hfill -->
<!--     \begin{subfigure}{0.23\textwidth} -->
<!--             \caption{Binary map $\small{S_{\widetilde{T}_{\lambda}}}$ (adapt.)} -->
<!--         \label{fig:fo-005-tsallis2_Ada} -->
<!--         \includegraphics[width=\linewidth]{Figures/H_005_FoggiaL24_eta_3.0n_5x11_B100_lambda_081.png} % -->
<!--     \end{subfigure} -->
<!-- %\caption*{\footnotesize\textbf{Source:} The author (2025)} -->
<!-- \end{figure} -->
<!-- ``` -->

In addition to the visual analysis presented above, we measured execution times to assess the computational cost of the proposed methods. Table&nbsp;\ref{tab:table_exec_times} reports the processing times for the three entropy-based tests (Shannon, Rényi, and Tsallis) with fixed $7\times7$ windows and, for comparison, the adaptive Tsallis variant with window sizes from $5\times5$ ($N_{\min}=5$) to $11\times11$ ($N_{\max}=11$). Times are reported in minutes and seconds (mm:ss) with $B=100$ bootstrap replications per window.

Overall, Shannon is the fastest across scenes, which is consistent with its simpler analytical form. Rényi and Tsallis are slightly slower due to the evaluation of additional terms. As expected, the adaptive Tsallis approach takes the longest because it evaluates multiple window sizes per pixel.

\renewcommand{\arraystretch}{1.8}   
```{r times_sar_images2, echo=FALSE, message=FALSE, warning=FALSE}

#library(kableExtra)

Times_data <- data.frame(
Scene = c("Image Simulated 1", "Image Simulated 2", "London",   "Munich", "New Orleans", "Dublin", "Coast of Jalisco"),
Shannon = c("01:12", "01:16", "12:10",   "07:02", "10:20", "8:39", "01:59"),
Renyi = c("01:17", "01:27", "13:28",   "8:54", "13:39", "9:44", "02:39"),
Tsallis = c("01:25", "01:38", "14:10",  "10:02", "14:39", "10:48", "02:54"),
Tsallis = c("01:50","02:05",   "16:22", "11:49", "15:48", "12:51", "03:20")
)

colnames(Times_data) <- c(
"\\textbf{Scene}",
"\\textbf{Shannon}",
"\\textbf{Rényi}",
"\\textbf{Tsallis}",
"\\textbf{Adaptive Tsallis}"
)

# colnames(Times_data) <- c(
#   "\\textbf{Scene}",
#   "\\textbf{Shannon (mm:ss)}",
#   "\\textbf{Rényi (mm:ss)}",
#   "\\textbf{Tsallis (mm:ss)}"
# )

kable(Times_data,
format = "latex",
booktabs = TRUE,
align = "ccccc",
escape = FALSE,
label = "table_exec_times",#adaptive windows with side lengths ranging from $5\times5$ ($N{\min}=5$) to $11\times11$ ($N_{\max}=11$)
centering = TRUE,
caption = "Execution times for the heterogeneity detection using fixed $7\\times7$ and adaptive windows ($N_{\\min}=5$, $N_{\\max}=11$), with $B=100$ bootstrap replications per scenario. Times are given in minutes and seconds (mm:ss). Rényi entropy uses $\\lambda=0.9$ and Tsallis entropy uses $\\lambda=0.85$.",
table.envir = "table", position = "hbt") %>%
row_spec(0, align = "c") %>%
kable_styling(latex_options = "scale_down", font_size = 16) %>%
kable_styling(full_width = TRUE) %>%
column_spec(1, width = "5cm") %>%
column_spec(2, width = "3.5cm") %>%
column_spec(3, width = "3.5cm") %>%
column_spec(4, width = "3.5cm") %>%
column_spec(5, width = "4cm")

```

All experiments were executed on the workstation described in @sec-repr, using 8–16 CPU cores. 
This setup allowed the entropy-based heterogeneity detection tests to process medium-sized SAR scenes within a few minutes, as reported in Table&nbsp;\ref{tab:table_exec_times}.

The total runtime also depends on the number of bootstrap replications ($B$), since each replication recomputes the test statistic to approximate its sampling distribution under the null hypothesis.  
In our simulations, we used $B=100$, which provides stable and reliable estimates with moderate computational cost. Stable results can already be achieved with $B\geq50$, whereas larger values of $B$ yield smoother $p$-value distributions at the expense of longer processing times.

<!-- Overall, the experiments on real SAR images confirm the simulated -->
<!-- findings: for highly speckled data ($L\!\approx\!1$) the Rényi -->
<!-- statistic is the most reliable detector of heterogeneity, whereas for -->
<!-- multi-look products the Tsallis statistic delivers cleaner segmentations -->
<!-- with fewer false positives.  -->

The window size also plays an important role in the heterogeneity detection process. This is particularly evident in the case of the Adaptive Tsallis test, where alternative sliding windows ranging from $5\times5$ to $11\times11$ were explored for both simulated and real SAR datasets. Although the observed differences were subtle, the adaptive approach exhibited the expected behavior: larger windows in homogeneous areas produced smoother and more stable estimates, whereas smaller windows in heterogeneous regions effectively preserved structural details and texture transitions.

For the single-look case ($L=1$), the adaptive Tsallis test using larger windows yielded more consistent detections, enhancing sensitivity to textural variations without generating excessive false alarms. This confirms that adaptive windowing improves the balance between detection accuracy and spatial smoothness, especially under strong speckle conditions.

Regarding the decision threshold, a significance level of $\alpha = 0.05$ was adopted for the binary classification of the $p$-value maps. Additional experiments with alternative levels showed the expected trade-offs: a stricter threshold ($\alpha = 0.01$) reduces false alarms but may overlook weak texture variations, while a more relaxed threshold ($\alpha = 0.10$) increases sensitivity to subtle heterogeneity at the expense of a higher false‐positive rate, particularly in highly speckled regions. Therefore, $\alpha = 0.05$ represents a balanced and robust choice for most SAR imaging scenarios.

### Quantitative Evaluation

Eight polygonal regions of interest (ROIs) were manually selected for each of the three SAR scenes used in the quantitative study—London, Munich, and Dublin (Figures&nbsp;\ref{fig:london-sar-roi1}–\ref{fig:dublin-sar-roi1}). ROIs labeled in red correspond to heterogeneous areas (class 1) and those in blue to homogeneous areas (class 0). The polygons were rasterized to the image grid to form a partial ground truth. We thresholded the $p$-value maps at $\alpha=0.05$, labeling each pixel as heterogeneous (1) or homogeneous (0), and compared the resulting decisions with the ground truth to compute the F1-score ($F_1$), the Kappa coefficient ($\kappa$), and the overall accuracy (OA). Results are summarized in Table&nbsp;\ref{tab:table_metrics1}.

```{=latex}
\begin{figure}[H]
    \centering
     \caption{Selected ROIs for the quantitative evaluation. Red polygons indicate heterogeneous areas (class 1), and blue polygons indicate homogeneous areas (class 0).}
    \label{fig:roi-scenes}
    % London
    \begin{subfigure}{0.30\textwidth}
        \includegraphics[width=\linewidth]{Figures/london_roi3.jpg}
        \caption{London SAR image with selected ROIs.}
        \label{fig:london-sar-roi1}
    \end{subfigure}
   \hspace{1em} 
    % Munich
    \begin{subfigure}{0.30\textwidth}
        \includegraphics[width=\linewidth]{Figures/munich_roi3.jpg}
        \caption{Munich SAR image with selected ROIs.}
        \label{fig:munich-sar-roi1}
    \end{subfigure}
    \hspace{1em}
    % Dublin
    \begin{subfigure}{0.30\textwidth}
        \includegraphics[width=\linewidth]{Figures/Dublin_roi3.jpg}
        \caption{Dublin SAR image with selected ROIs.}
        \label{fig:dublin-sar-roi1}
    \end{subfigure}
   %\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}

```

\renewcommand{\arraystretch}{1.3}
```{r metrics_table_k, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
suppressPackageStartupMessages({
library(dplyr)
library(knitr)
library(kableExtra)
})

load("./Data/metrics_london_shannon.RData")
load("./Data/metrics_london_renyi.RData")
load("./Data/metrics_london_tsallis.RData")
load("./Data/metrics_tsallis_london_eta3_f.RData")

load("./Data/metrics_munich_shannon.RData")
load("./Data/metrics_munich_renyi.RData")
load("./Data/metrics_munich_tsallis.RData")
load("./Data/metrics_tsallis_munich_eta3_f.RData")

load("./Data/metrics_dublin_shannon.RData")
load("./Data/metrics_dublin_renyi.RData")
load("./Data/metrics_dublin_tsallis.RData")
load("./Data/metrics_tsallis_dublin_eta3_f.RData")

all_metrics <- dplyr::bind_rows(
metrics_london_shannon,
metrics_london_renyi,
metrics_london_tsallis,
metrics_tsallis_london_eta3_f,

metrics_munich_shannon,
metrics_munich_renyi,
metrics_munich_tsallis,
metrics_tsallis_munich_eta3_f,

metrics_dublin_shannon,
metrics_dublin_renyi,
metrics_dublin_tsallis,
metrics_tsallis_dublin_eta3_f
) %>%
dplyr::select(Scene, Method, F1, Kappa, OA)

all_metrics$Scene <- factor(all_metrics$Scene, levels = c("London","Munich","Dublin"))

all_metrics$Method <- dplyr::recode(
all_metrics$Method,
"Shannon"      = "$S_{\\widetilde{H}_{\\text{AO}}}$",
"Renyi"        = "$S_{\\widetilde{R}_{\\lambda}}$",
"Tsallis"      = "$S_{\\widetilde{T}_{\\lambda}}$",
"Tsallis_eta3" = "\\quad $S_{\\widetilde{T}_{\\lambda}}$ adapt.",
.default       = as.character(all_metrics$Method)
)

fmt <- function(x) sprintf("$%.3f$", as.numeric(x))
all_metrics <- all_metrics %>%
group_by(Scene) %>%
mutate(
F1_max    = max(as.numeric(F1),    na.rm = TRUE),
Kappa_max = max(as.numeric(Kappa), na.rm = TRUE),
OA_max    = max(as.numeric(OA),    na.rm = TRUE),
F1    = ifelse(as.numeric(F1)    == F1_max,    paste0("\\bm{", fmt(F1), "}"),    fmt(F1)),
Kappa = ifelse(as.numeric(Kappa) == Kappa_max, paste0("\\bm{", fmt(Kappa), "}"), fmt(Kappa)),
OA    = ifelse(as.numeric(OA)    == OA_max,    paste0("\\bm{", fmt(OA), "}"),    fmt(OA))
) %>%
ungroup() %>%
dplyr::select(Scene, Method, F1, Kappa, OA)

colnames(all_metrics) <- c("\\textbf{Scene}", "\\textbf{Test}", "$\\bm{F_1}$", "$\\kappa$", "\\textbf{OA}")

kable(all_metrics,
caption = "Quantitative measures of heterogeneity detection",
format = "latex",
booktabs = TRUE,
align = "ccccc",      
escape = FALSE,
label = "table_metrics1",
row.names = FALSE) %>%
row_spec(0, align = "c") %>%
collapse_rows(columns = 1, latex_hline = "major", valign = "bottom") %>%
kable_styling(latex_options = c("hold_position","scale_down"), font_size = 11, full_width = T) %>%
print()

```

When comparing the three entropy-based tests, distinct patterns emerge depending on the scene characteristics.

In London, the adaptive Tsallis test achieves the highest scores across metrics ($F_1 = 0.757$, $\kappa = 0.697$, OA = 0.897), followed by the Rényi-based statistic ($F_1 = 0.695$, $\kappa = 0.626$, OA = 0.877). Both outperform Shannon ($F_1 = 0.617$, $\kappa = 0.542$, OA = 0.854) and fixed-window Tsallis ($F_1 = 0.603$, $\kappa = 0.528$, OA = 0.850). These results suggest that adaptive windowing is helpful under stronger speckle by smoothing homogeneous areas while preserving edges.

In Munich, all methods perform well, but the adaptive Tsallis variant is slightly superior ($F_1 = 0.935$, $\kappa = 0.880$, OA = 0.940), with fixed Tsallis close behind ($F_1 = 0.930$, $\kappa = 0.871$, OA = 0.936) and Rényi competitive ($F_1 = 0.924$, $\kappa = 0.861$, OA = 0.931). Shannon shows lower but still strong values ($F_1 = 0.883$, $\kappa = 0.794$, OA = 0.897).

In Dublin, the adaptive approach yields the largest gains: adaptive Tsallis leads with $F_1 = 0.901$, $\kappa = 0.844$, and OA = 0.926, followed by fixed Tsallis ($F_1 = 0.879$, $\kappa = 0.776$, OA = 0.887) and Rényi ($F_1 = 0.865$, $\kappa = 0.753$, OA = 0.875). Shannon again records the lowest figures ($F_1 = 0.645$, $\kappa = 0.464$, OA = 0.858).

Overall, the adaptive Tsallis test provides the most accurate classification across the three scenes, with consistent improvements in $F_1$, $\kappa$, and OA. Rényi remains a strong fixed-window alternative, while Shannon is the least effective under the tested conditions. These findings align with the qualitative inspection: adaptive windowing improves the balance between detail preservation in heterogeneous regions and stability in homogeneous areas.

<!-- Using the same ROIs as ground truth, we evaluated the continuous $p$-value maps by computing receiver operating characteristic (ROC) curves over all decision thresholds (Figure \ref{fig:roc}). Each curve plots the probability of detection (Pd) versus the probability of false alarm (PFA). Pd reflects the proportion of heterogeneous pixels correctly detected, while PFA represents the proportion of homogeneous pixels misclassified as heterogeneous. -->

<!-- ```{=latex} -->
<!-- \begin{figure}[H] -->
<!--   \centering -->
<!--   \caption{ROC curves comparing the tests $S_{\widetilde{H}_{\text{AO}}}$, $S_{\widetilde{R}_{\lambda}}$ and $S_{\widetilde{T}_{\lambda}}$.} -->
<!--   \label{fig:roc} -->
<!--   \includegraphics[width=0.98\textwidth]{Figures/fig-ROC6-1.pdf} -->
<!--   %\caption*{\footnotesize\textbf{Source:} The author (2025)} -->
<!-- \end{figure} -->
<!-- ``` -->
<!-- The area under each ROC curve (AUC) quantifies overall detection performance, with higher values indicating better discrimination between classes. Across all scenes, the Rényi-based test yielded the highest AUC in London ($0.963$), followed by Tsallis ($0.824$) and Shannon ($0.776$), reaffirming Rényi’s robustness under strong speckle conditions. -->

<!-- In Munich, both Rényi and Tsallis performed similarly with near-perfect AUCs ($0.996$ and $0.998$ respectively), while Shannon slightly lagged behind ($0.989$). In Dublin, a similar pattern is observed: Tsallis again leads marginally with an AUC of $0.994$, followed by Rényi ($0.990$) and Shannon ($0.986$). -->

<!-- These results confirm that Rényi and Tsallis entropies consistently outperform Shannon, with Tsallis showing slightly superior performance in higher-quality scenes. The ROC analysis aligns with the classification metrics, supporting the conclusion that Tsallis and Rényi are more effective in detecting heterogeneity across varying image conditions. -->

## INTERACTIVE SHINY APP FOR HETEROGENEITY DETECTION

Shiny is an R package (version 1.9.1) developed by RStudio (now Posit) that makes it possible to build interactive web applications directly from R code, without requiring knowledge of HTML, CSS or JavaScript.  Users can manipulate inputs (e.g. sliders, file uploads, dropdowns) and see results update instantly in the browser.

To facilitate reproducibility and allow exploration of the heterogeneity-detection pipeline on both simulated and real SAR images, we have developed a Shiny web application.

The app supports three data‐loading modes:

- **Simulated image examples**: Choose one of two simulated images (e.g. $L=5$ or $L=9$).

- **ENVI sample images**:  Select one of three small SAR samples in ENVI format (e.g. Dublin $L=16$, London $L=1$, or New Orleans $L=12$).

- **Upload ENVI**: Upload any custom \texttt{.img}/\texttt{.hdr} pair to process your own data.

Once the image is loaded, the user can select one of three entropy estimators (Shannon, Rényi, or Tsallis), set the order $\lambda$ (for Rényi and Tsallis), specify the number of looks $L$, enable or disable bootstrap resampling (and choose the number of replicates), and fix the sliding‐window size $n \times n$. By clicking **Run Detection**, the app performs the sliding‐window test and displays the resulting $p$-value heat‐map in color.
```{=latex}
\begin{figure}[H]
  \centering
    \caption{Shiny application interface. The left panel shows the data‐loading and parameter controls; the right panel displays the colorized $p$-value map and elapsed runtime.}
  \label{fig:shiny-app}
  \includegraphics[width=0.96\textwidth]{Figures/shiny1.jpg}
%\caption*{\footnotesize\textbf{Source:} The author (2025)}
\end{figure}
```
This interactive tool complements the results presented in Sections \ref{sec-sim2} and \ref{sec-sim3} by enabling real‐time parameter tuning and direct comparison across estimators.

The application is publicly accessible at:
[https://janeth-alpala.shinyapps.io/heterogeneity\_detection\_app/](https://janeth-alpala.shinyapps.io/heterogeneity\_detection\_app/)