### 2025-10-27 Text/Thesis/Frontmatter/acknowledgements.tex

Firstly, I would like to thank my supervisors, Prof. Alejandro Frery and Prof. Abraão do Nascimento, for their guidance and support throughout these years. Working with you has been an incredibly valuable learning experience. I truly appreciate the attention and dedication you give to your students.

I am also thankful to the Universidade Federal de Pernambuco and the Programa de
Pós-Graduação em Estatística for their support during my academic journey. Special thanks to Prof. Pablo Rodriguez for all his help with administrative matters and for his patience.

I acknowledge the financial support from the Coordenação de Aperfeiçoamento de Pessoal de Nível Superior – Brasil (CAPES) – Finance Code 001, the Programa de Doutorado-Sanduíche no Exterior (PDSE), and the Fundação de Amparo à Ciência e Tecnologia do Estado de Pernambuco (FACEPE), which made this work possible.

I would like to express my gratitude to the members of my examination committee, Prof. Saeid Homayouni, Prof. Paolo Gamba, Prof. Anderson de Borba, and Prof. Fátima Sombra, for agreeing to be part of the defense of my thesis and for their valuable comments and suggestions.

This journey would not have been the same without all the amazing people I met in Brazil and New Zealand. To everyone, thank you for your friendship and support, which made this experience unforgettable.

Finally, I am especially grateful to my family, who have always supported and encouraged me along the way.